\documentclass[10pt,a4paper]{article}

\usepackage[a4paper,margin=0.55in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{hyperref}

\hypersetup{hidelinks}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2pt}
\setlist[enumerate]{leftmargin=*, itemsep=1.5pt, topsep=2pt}

\newcommand{\ordic}[1]{\textbf{#1}}
\newcommand{\gloss}[1]{\hfill{\small\textsf{#1}}}

\begin{document}

{\LARGE \textbf{High Ordic: Core Phrases \small{(v0.1)}}}\\

\vspace{6pt}\hrule\vspace{8pt}

\begin{multicols}{2}

\begin{enumerate}
% --- Being / Order (axioms)
\item \ordic{nuu es.} \gloss{Being is (now/eternally).}
\item \ordic{te ord es.} \gloss{Truth: order is.}
\item \ordic{set ord es.} \gloss{Structure exists as order.}
\item \ordic{te-log es.} \gloss{Reasoning exists (proofhood).}
\item \ordic{mi es.} \gloss{Meaning exists.}
\item \ordic{te mi es.} \gloss{True meaning exists.}
\item \ordic{ord te.} \gloss{Order is true.}
\item \ordic{te vi.} \gloss{Truth is witnessed.}
\item \ordic{vi-te es.} \gloss{Witnessing-truth exists.}
\item \ordic{lu te.} \gloss{Knowledge aligns with truth.}
\item \ordic{mu es.} \gloss{Memory exists.}
\item \ordic{mu-te es.} \gloss{True memory exists (unaltered record).}
\item \ordic{ka ord es.} \gloss{Form is ordered.}
\item \ordic{il ord es.} \gloss{Essence is ordered.}
\item \ordic{un es.} \gloss{Unity exists.}
\item \ordic{du es.} \gloss{Otherness exists.}
\item \ordic{al es.} \gloss{Totality exists.}
\item \ordic{par es.} \gloss{Part exists.}
\item \ordic{eq un du.} \gloss{The one equals the other (equivalence).}
\item \ordic{dif un du.} \gloss{The one differs from the other.}

% --- Void / absence (careful negation)
\item \ordic{pra zen es-na.} \gloss{Before: void was absent (void did not obtain).}
\item \ordic{zen te-na.} \gloss{Void is without truth.}
\item \ordic{na es zen.} \gloss{Non-being is void (absence maps to null).}
\item \ordic{na mi zen.} \gloss{Without meaning: null.}
\item \ordic{na ßo zen.} \gloss{Without binding: null.}
\item \ordic{na ord zen.} \gloss{Without order: null.}
\item \ordic{na te zen.} \gloss{Without truth: null.}
\item \ordic{zen set.} \gloss{Empty set / null structure.}
\item \ordic{zen ord.} \gloss{Zero-order / no sequence.}
\item \ordic{zen mu.} \gloss{No memory / erased trace.}

% --- Time (timeless but referencable)
\item \ordic{pra ord.} \gloss{Order was (past order).}
\item \ordic{nuu ord.} \gloss{Order is (present order).}
\item \ordic{pos ord.} \gloss{Order will be (future order).}
\item \ordic{dur es.} \gloss{Being continues.}
\item \ordic{dur ord.} \gloss{Order continues.}
\item \ordic{dur vi-te.} \gloss{Truth is witnessed continuously.}
\item \ordic{pun vi-te.} \gloss{Truth is witnessed at an instant.}
\item \ordic{ret ord.} \gloss{Order returns / repeats.}
\item \ordic{ret vi-te.} \gloss{Witness returns / is reaffirmed.}
\item \ordic{vek ord.} \gloss{Order becomes / changes state.}
\item \ordic{pra vek.} \gloss{Becoming occurred (change happened).}
\item \ordic{pos vek.} \gloss{Becoming will occur (change will happen).}

% --- Space / relation
\item \ordic{in ord.} \gloss{Within order.}
\item \ordic{ex ord.} \gloss{Outside order.}
\item \ordic{lin ord.} \gloss{Toward order (path to order).}
\item \ordic{tri ord.} \gloss{Through order.}
\item \ordic{ne ord.} \gloss{Near order (adjacent alignment).}
\item \ordic{far ord.} \gloss{Far from order (distant alignment).}
\item \ordic{in te.} \gloss{Within truth.}
\item \ordic{ex te.} \gloss{Beyond truth (outside the witnessed domain).}
\item \ordic{lin te.} \gloss{Toward truth.}
\item \ordic{tri te.} \gloss{Through truth (by truth's channel).}
\item \ordic{su ord.} \gloss{Above order (higher ordering).}
\item \ordic{da ord.} \gloss{Below order (lower ordering).}

% --- Law / permission / constraint (High Ordic style)
\item \ordic{jur es.} \gloss{Law exists.}
\item \ordic{jur ord.} \gloss{Law orders.}
\item \ordic{dar ord.} \gloss{Order must.}
\item \ordic{fre ord.} \gloss{Order may (permission granted to order).}
\item \ordic{dar jur.} \gloss{Law must (mandatory law).}
\item \ordic{fre jur.} \gloss{Law may (discretionary law).}
\item \ordic{dar vi-te.} \gloss{Truth must be witnessed.}
\item \ordic{dar mu-te.} \gloss{True memory must persist (record must hold).}
\item \ordic{fre mu.} \gloss{Memory may persist (allowed to remember).}
\item \ordic{dar set-ord.} \gloss{Systems must be ordered.}
\item \ordic{jur set-ord.} \gloss{Rule: systems are ordered.}
\item \ordic{jur eq.} \gloss{Rule of equality.}
\item \ordic{jur dif.} \gloss{Rule of difference.}

% --- Witness / record formulas
\item \ordic{te vi dur.} \gloss{Truth is witnessed continuously.}
\item \ordic{te vi pun.} \gloss{Truth is witnessed at the point (instant witness).}
\item \ordic{mu vi-te.} \gloss{Memory witnesses truth (record attests).}
\item \ordic{mu-te vi.} \gloss{True memory is witnessed.}
\item \ordic{so mu-te.} \gloss{Speak the true record (recite).}
\item \ordic{so vi-te.} \gloss{Speak the witnessed truth.}
\item \ordic{so-mi-es.} \gloss{Meaning is spoken into being (naming act).}
\item \ordic{so te-mi.} \gloss{Speak true meaning (define).}
\item \ordic{so ord.} \gloss{Speak order (pronounce alignment).}
\item \ordic{so jur.} \gloss{Speak law (declare ruling).}

% --- Binding / covenant (minimal emotion; High Ordic uses it sparingly)
\item \ordic{ßo es.} \gloss{Binding exists.}
\item \ordic{ßo ord.} \gloss{Binding orders (connections create order).}
\item \ordic{sol es.} \gloss{Vow exists.}
\item \ordic{sol ßo.} \gloss{Vow binds.}
\item \ordic{ßo-sol es.} \gloss{Bound vow exists.}
\item \ordic{dar sol.} \gloss{A vow must (obligation stands).}
\item \ordic{fre sol.} \gloss{A vow may (permitted vow).}
\item \ordic{ßo dur.} \gloss{Binding continues.}
\item \ordic{ßo pun.} \gloss{Binding is enacted at an instant.}
\item \ordic{ret ßo.} \gloss{Binding is renewed.}

% --- Mercy / release (rare, controlled)
\item \ordic{gra es.} \gloss{Grace/mercy exists.}
\item \ordic{gra ša-ßo.} \gloss{Mercy releases binding.}
\item \ordic{ša ßo.} \gloss{Sever binding.}
\item \ordic{ša ord.} \gloss{Sever order (break ordering).}
\item \ordic{jur gra.} \gloss{Law contains mercy (law is tempered).}
\item \ordic{dar gra.} \gloss{Mercy must (required clemency).}
\item \ordic{fre gra.} \gloss{Mercy may (permitted clemency).}
\item \ordic{gra te.} \gloss{Mercy aligns with truth.}
\item \ordic{gra ord.} \gloss{Mercy preserves order.}
\item \ordic{gra mu.} \gloss{Mercy preserves memory (no erasure).}

% --- Calm (the sealed axis)
\item \ordic{šael es.} \gloss{Calm exists.}
\item \ordic{šael ord.} \gloss{Calm orders (stabilizes).}
\item \ordic{šael vi-te.} \gloss{Calm witnesses truth (clear seeing).}
\item \ordic{šael dur.} \gloss{Calm continues.}
\item \ordic{šael ret.} \gloss{Calm returns.}

% --- Causality / agency (impersonal)
\item \ordic{ta es.} \gloss{Causation exists.}
\item \ordic{ta ord.} \gloss{Causation orders outcomes.}
\item \ordic{ki es.} \gloss{Action exists.}
\item \ordic{ki ord.} \gloss{Action orders (acts create structure).}
\item \ordic{ta ki.} \gloss{Cause acts / causation triggers action.}
\item \ordic{se ki.} \gloss{Allow action.}
\item \ordic{vu ki.} \gloss{Resist action.}
\item \ordic{ha-ri ord.} \gloss{Giving-taking orders exchange.}
\item \ordic{ha te.} \gloss{Give truth (reveal).}
\item \ordic{ri te.} \gloss{Take truth (accept testimony).}

% --- Mathematics / structure (explicit)
\item \ordic{set es.} \gloss{A set exists.}
\item \ordic{set-ord es.} \gloss{An ordered set exists (sequence).}
\item \ordic{ord-set es.} \gloss{An ordering-of-sets exists (system).}
\item \ordic{sum set.} \gloss{Add/gather sets (union-ish).}
\item \ordic{dif set.} \gloss{Difference of sets.}
\item \ordic{mul set.} \gloss{Multiply/amplify sets (product-ish).}
\item \ordic{eq set.} \gloss{Set equality holds.}
\item \ordic{ord eq.} \gloss{Ordering equality (stable equivalence).}
\item \ordic{log set.} \gloss{Reason about a set.}
\item \ordic{te-log set.} \gloss{True reasoning about a set (proof on a set).}

% --- Perception / mind (high register)
\item \ordic{me es.} \gloss{Mind exists.}
\item \ordic{me ord.} \gloss{Mind orders (cognition structures).}
\item \ordic{vi es.} \gloss{Perception exists.}
\item \ordic{vi-te.} \gloss{Witness truth.}
\item \ordic{lu es.} \gloss{Knowing exists.}
\item \ordic{lu ret.} \gloss{Knowledge returns (relearn/reaffirm).}
\item \ordic{qe es.} \gloss{Doubt exists.}
\item \ordic{qe te.} \gloss{Doubt addresses truth (uncertainty).}
\item \ordic{qe ord.} \gloss{Doubt addresses order (questioning structure).}
\item \ordic{qe log.} \gloss{Doubt addresses reasoning (challenge).}

% --- Controlled fracture statements (canonical)
\item \ordic{qe vek.} \gloss{Doubt becomes.}
\item \ordic{ir vek.} \gloss{Anger becomes.}
\item \ordic{ord ša.} \gloss{Order breaks.}
\item \ordic{par far.} \gloss{Parts scatter far.}
\item \ordic{ex ord.} \gloss{Outside order.}

% --- Exile / boundary
\item \ordic{dar ex-ord.} \gloss{Must be placed outside order (exile).}
\item \ordic{ex-ord es.} \gloss{Exile-state exists.}
\item \ordic{ex-ord dur.} \gloss{Exile continues.}
\item \ordic{far vi-te.} \gloss{Truth is witnessed from afar.}
\item \ordic{ex mu.} \gloss{Out of memory (forgotten/erased).}
\item \ordic{ex set.} \gloss{Outside the set (excluded).}
\item \ordic{ex jur.} \gloss{Outside law (non-jurisdiction).}
\item \ordic{lin ord-ret.} \gloss{Toward return-of-order (restoration).}

% --- Artifact / access formulas (High Ordic templates)
\item \ordic{fre lin.} \gloss{Permission to proceed (may go).}
\item \ordic{dar lin.} \gloss{Must proceed (compelled path).}
\item \ordic{se lin.} \gloss{Allow passage.}
\item \ordic{vu lin.} \gloss{Resist passage.}
\item \ordic{te jur-se.} \gloss{Truth allows by law (authorized).}
\item \ordic{jur-se ord.} \gloss{Law permits order (authorization condition).}
\item \ordic{gra ßo ord.} \gloss{Mercy binds order (gentle constraint).}
\item \ordic{gra ßo set-ord.} \gloss{Mercy binds the ordered system.}
\item \ordic{gra ßo we ord.} \gloss{Mercy binds the collective order (gate formula).}
\item \ordic{te vi ša.} \gloss{Truth witnessed through severance (ša-Te core).}

% --- Case-marked short formulas (minimal, inscription-like)
\item \ordic{te-in.} \gloss{In truth.}
\item \ordic{te-ex.} \gloss{Out of truth.}
\item \ordic{ord-in.} \gloss{Within order.}
\item \ordic{ord-ex.} \gloss{Out of order / beyond order.}
\item \ordic{jur-il.} \gloss{Of law / legal essence.}
\item \ordic{mu-il.} \gloss{Of memory / memorial essence.}
\item \ordic{set-ßo.} \gloss{By means of structure.}
\item \ordic{log-ßo.} \gloss{By means of reasoning.}
\item \ordic{vi-te-ßo.} \gloss{By means of witnessing truth.}
\item \ordic{gra-ßo.} \gloss{By means of mercy.}

% --- Additional canonical micro-clauses (register pure)
\item \ordic{te set ord.} \gloss{Truth structures order.}
\item \ordic{te-log ord.} \gloss{True reasoning orders.}
\item \ordic{mi ord.} \gloss{Meaning orders (semantics structures).}
\item \ordic{so-mi ord.} \gloss{Speech-meaning orders (language aligns).}
\item \ordic{mu ord.} \gloss{Memory orders (records constrain).}
\item \ordic{jur mu-te.} \gloss{Law preserves true record.}
\item \ordic{jur vi-te.} \gloss{Law requires witnessing truth.}
\item \ordic{sol jur.} \gloss{Vow under law.}
\item \ordic{gra jur.} \gloss{Mercy within law.}
\item \ordic{ša jur.} \gloss{Sever under law (sanction).}

% --- Tight technical imperatives (still pronounless)
\item \ordic{dar so.} \gloss{Must speak (declare).}
\item \ordic{dar mu.} \gloss{Must remember (retain record).}
\item \ordic{dar vi.} \gloss{Must perceive (attend).}
\item \ordic{dar lu.} \gloss{Must know (be informed).}
\item \ordic{dar set.} \gloss{Must structure (form a set).}
\item \ordic{dar log.} \gloss{Must reason (infer).}
\item \ordic{dar eq.} \gloss{Must equate (treat as same).}
\item \ordic{dar dif.} \gloss{Must distinguish (treat as different).}
\item \ordic{dar sum.} \gloss{Must gather (aggregate).}
\item \ordic{dar ord.} \gloss{Must order (sequence/arrange).}

% --- Closing (sealed hint, not the final utterance)
\item \ordic{ret ord?} \gloss{Return of order? (sealed question; never casual).}
\item \ordic{šael\ldots} \gloss{Calm\ldots (sealed pause).}

% --- Alignment / validity
\item \ordic{te set es.} \gloss{A true structure exists.}
\item \ordic{te ord dur.} \gloss{True order continues.}
\item \ordic{te-log dur.} \gloss{True reasoning continues.}
\item \ordic{te mi dur.} \gloss{True meaning persists.}
\item \ordic{te mu dur.} \gloss{True memory persists.}
\item \ordic{te ßo dur.} \gloss{True binding persists.}
\item \ordic{te jur es.} \gloss{True law exists.}
\item \ordic{te sol es.} \gloss{A true vow exists.}
\item \ordic{te gra es.} \gloss{True mercy exists.}
\item \ordic{te šae l es.} \gloss{True calm exists.}

% --- Validity checks
\item \ordic{vi te-log.} \gloss{Witness the proof.}
\item \ordic{vi mu-te.} \gloss{Witness the true record.}
\item \ordic{vi jur-te.} \gloss{Witness lawful truth.}
\item \ordic{vi sol-te.} \gloss{Witness a true vow.}
\item \ordic{vi gra-te.} \gloss{Witness true mercy.}
\item \ordic{vi ord-te.} \gloss{Witness true order.}
\item \ordic{vi set-te.} \gloss{Witness true structure.}
\item \ordic{vi log-te.} \gloss{Witness true reasoning.}
\item \ordic{vi mi-te.} \gloss{Witness true meaning.}
\item \ordic{vi ßo-te.} \gloss{Witness true binding.}

% --- Authorization / access
\item \ordic{jur se lin.} \gloss{Law allows passage.}
\item \ordic{jur vu lin.} \gloss{Law resists passage.}
\item \ordic{te jur se.} \gloss{True law permits.}
\item \ordic{te jur vu.} \gloss{True law forbids.}
\item \ordic{fre lin ord.} \gloss{Permission toward order.}
\item \ordic{dar lin ord.} \gloss{Compelled toward order.}
\item \ordic{se lin set.} \gloss{Allow entry to structure.}
\item \ordic{vu lin set.} \gloss{Resist entry to structure.}
\item \ordic{se lin ord.} \gloss{Allow entry to order.}
\item \ordic{vu lin ord.} \gloss{Resist entry to order.}

% --- Record keeping / archives
\item \ordic{mu set es.} \gloss{A memory-set exists (archive).}
\item \ordic{mu set-ord es.} \gloss{An ordered archive exists.}
\item \ordic{mu set dur.} \gloss{Archive persists.}
\item \ordic{mu set ret.} \gloss{Archive is restored.}
\item \ordic{mu set ša.} \gloss{Archive is severed (purge).}
\item \ordic{mu-te set.} \gloss{True record set.}
\item \ordic{mu-te set-ord.} \gloss{Ordered true archive.}
\item \ordic{jur mu-te.} \gloss{Law preserves true record.}
\item \ordic{dar mu-te.} \gloss{True record must persist.}
\item \ordic{fre mu-te.} \gloss{True record may persist.}

% --- Consistency / contradiction
\item \ordic{eq mi mi.} \gloss{Meanings are equivalent.}
\item \ordic{dif mi mi.} \gloss{Meanings differ.}
\item \ordic{eq log log.} \gloss{Reasonings are equivalent.}
\item \ordic{dif log log.} \gloss{Reasonings differ.}
\item \ordic{eq ord ord.} \gloss{Orders are equivalent.}
\item \ordic{dif ord ord.} \gloss{Orders differ.}
\item \ordic{qe eq.} \gloss{Equality is questioned.}
\item \ordic{qe dif.} \gloss{Difference is questioned.}
\item \ordic{qe te-log.} \gloss{Proof is questioned.}
\item \ordic{qe mu-te.} \gloss{Record is questioned.}

% --- Sanctions / severance (controlled)
\item \ordic{jur ša-ßo.} \gloss{Law releases binding.}
\item \ordic{te jur ša.} \gloss{True law severs.}
\item \ordic{dar ša-ßo.} \gloss{Binding must be released.}
\item \ordic{fre ša-ßo.} \gloss{Binding may be released.}
\item \ordic{ša set-ord.} \gloss{Sever the ordered system.}
\item \ordic{ša log.} \gloss{Sever reasoning (invalidate).}
\item \ordic{ša mi.} \gloss{Sever meaning (nullify).}
\item \ordic{ša mu.} \gloss{Sever memory (erase).}
\item \ordic{ša ord.} \gloss{Sever order.}
\item \ordic{ša ßo.} \gloss{Sever binding.}

% --- Restoration / repair
\item \ordic{ret set-ord.} \gloss{Restore the ordered system.}
\item \ordic{ret mu-te.} \gloss{Restore true record.}
\item \ordic{ret ßo.} \gloss{Restore binding.}
\item \ordic{ret ord.} \gloss{Restore order.}
\item \ordic{ret te-log.} \gloss{Restore proof.}
\item \ordic{ret mi.} \gloss{Restore meaning.}
\item \ordic{gra ret.} \gloss{Mercy enables return.}
\item \ordic{gra ret ord.} \gloss{Mercy restores order.}
\item \ordic{gra ret ßo.} \gloss{Mercy restores binding.}
\item \ordic{gra ret mu.} \gloss{Mercy restores memory.}

% --- Measurement / comparison
\item \ordic{al set.} \gloss{The whole set.}
\item \ordic{par set.} \gloss{A subset / part-set.}
\item \ordic{al ord.} \gloss{Total order.}
\item \ordic{par ord.} \gloss{Partial order.}
\item \ordic{sum set.} \gloss{Aggregate structures.}
\item \ordic{dif set.} \gloss{Differentiate structures.}
\item \ordic{mul set.} \gloss{Amplify structures.}
\item \ordic{eq set.} \gloss{Structures are equal.}
\item \ordic{ord sum.} \gloss{Order the aggregate.}
\item \ordic{ord dif.} \gloss{Order differences.}

% --- Perception / awareness
\item \ordic{vi ord.} \gloss{Perceive order.}
\item \ordic{vi set.} \gloss{Perceive structure.}
\item \ordic{vi log.} \gloss{Perceive reasoning.}
\item \ordic{vi mi.} \gloss{Perceive meaning.}
\item \ordic{vi mu.} \gloss{Perceive memory.}
\item \ordic{lu ord.} \gloss{Know order.}
\item \ordic{lu set.} \gloss{Know structure.}
\item \ordic{lu log.} \gloss{Know reasoning.}
\item \ordic{lu mi.} \gloss{Know meaning.}
\item \ordic{lu mu.} \gloss{Know memory.}

% --- Stability / persistence
\item \ordic{dur set.} \gloss{Structure persists.}
\item \ordic{dur mi.} \gloss{Meaning persists.}
\item \ordic{dur mu.} \gloss{Memory persists.}
\item \ordic{dur ßo.} \gloss{Binding persists.}
\item \ordic{dur jur.} \gloss{Law persists.}
\item \ordic{dur sol.} \gloss{Vow persists.}
\item \ordic{dur gra.} \gloss{Mercy persists.}
\item \ordic{dur šae l.} \gloss{Calm persists.}
\item \ordic{pun ša.} \gloss{Instant severance.}
\item \ordic{pun ßo.} \gloss{Instant binding.}

% --- Boundary / scope
\item \ordic{in set.} \gloss{Within the structure.}
\item \ordic{ex set.} \gloss{Outside the structure.}
\item \ordic{in mu.} \gloss{Within memory.}
\item \ordic{ex mu.} \gloss{Outside memory.}
\item \ordic{in jur.} \gloss{Within law.}
\item \ordic{ex jur.} \gloss{Outside law.}
\item \ordic{in gra.} \gloss{Within mercy.}
\item \ordic{ex gra.} \gloss{Outside mercy.}
\item \ordic{in ord.} \gloss{Within order.}
\item \ordic{ex ord.} \gloss{Outside order.}

% --- Closure
\item \ordic{te dur.} \gloss{Truth endures.}
\item \ordic{ord dur.} \gloss{Order endures.}
\item \ordic{mi dur.} \gloss{Meaning endures.}
\item \ordic{mu dur.} \gloss{Memory endures.}
\item \ordic{ßo dur.} \gloss{Binding endures.}
\item \ordic{jur dur.} \gloss{Law endures.}
\item \ordic{gra dur.} \gloss{Mercy endures.}
\item \ordic{šael dur.} \gloss{Calm endures.}
\item \ordic{ret dur.} \gloss{Return endures.}
\item \ordic{ret ord-ret.} \gloss{Return of order remembered.}

% --- Ontological consistency
\item \ordic{es ord-te.} \gloss{Being aligns with true order.}
\item \ordic{es te-log.} \gloss{Being follows true reasoning.}
\item \ordic{es mi-te.} \gloss{Being aligns with true meaning.}
\item \ordic{es mu-te.} \gloss{Being aligns with true memory.}
\item \ordic{es ßo-te.} \gloss{Being aligns with true binding.}
\item \ordic{es jur-te.} \gloss{Being aligns with true law.}
\item \ordic{es gra-te.} \gloss{Being aligns with true mercy.}
\item \ordic{es šae l-te.} \gloss{Being aligns with true calm.}
\item \ordic{ord es-te.} \gloss{Order aligns with being-truth.}
\item \ordic{mi es-te.} \gloss{Meaning aligns with being-truth.}

% --- Verification / validation
\item \ordic{vi es-te.} \gloss{Witness true being.}
\item \ordic{vi ord-te.} \gloss{Witness true order.}
\item \ordic{vi set-te.} \gloss{Witness true structure.}
\item \ordic{vi mi-te.} \gloss{Witness true meaning.}
\item \ordic{vi mu-te.} \gloss{Witness true memory.}
\item \ordic{vi jur-te.} \gloss{Witness true law.}
\item \ordic{vi sol-te.} \gloss{Witness true vow.}
\item \ordic{vi gra-te.} \gloss{Witness true mercy.}
\item \ordic{vi šae l-te.} \gloss{Witness true calm.}
\item \ordic{vi log-te.} \gloss{Witness true reasoning.}

% --- Structural stability
\item \ordic{set ord dur.} \gloss{Structured order persists.}
\item \ordic{ord set dur.} \gloss{Ordering of structures persists.}
\item \ordic{set log dur.} \gloss{Structured reasoning persists.}
\item \ordic{log set dur.} \gloss{Reasoned structures persist.}
\item \ordic{set mi dur.} \gloss{Structured meaning persists.}
\item \ordic{mi set dur.} \gloss{Meaningful structures persist.}
\item \ordic{set mu dur.} \gloss{Structured memory persists.}
\item \ordic{mu set dur.} \gloss{Memory-structures persist.}
\item \ordic{set jur dur.} \gloss{Structured law persists.}
\item \ordic{jur set dur.} \gloss{Law-structures persist.}

% --- Logical flow / inference
\item \ordic{log ord.} \gloss{Reasoning orders.}
\item \ordic{log set.} \gloss{Reasoning structures.}
\item \ordic{log mi.} \gloss{Reasoning clarifies meaning.}
\item \ordic{log mu.} \gloss{Reasoning preserves memory.}
\item \ordic{log jur.} \gloss{Reasoning supports law.}
\item \ordic{log te.} \gloss{Reasoning aligns with truth.}
\item \ordic{te-log ord.} \gloss{True reasoning orders.}
\item \ordic{te-log set.} \gloss{True reasoning structures.}
\item \ordic{te-log mu.} \gloss{True reasoning preserves memory.}
\item \ordic{te-log jur.} \gloss{True reasoning supports law.}

% --- Constraints / permissions (formal)
\item \ordic{dar es.} \gloss{Being must (necessity holds).}
\item \ordic{fre es.} \gloss{Being may (existence permitted).}
\item \ordic{dar set.} \gloss{Structure must.}
\item \ordic{fre set.} \gloss{Structure may.}
\item \ordic{dar log.} \gloss{Reasoning must.}
\item \ordic{fre log.} \gloss{Reasoning may.}
\item \ordic{dar mu.} \gloss{Memory must.}
\item \ordic{fre mu.} \gloss{Memory may.}
\item \ordic{dar mi.} \gloss{Meaning must.}
\item \ordic{fre mi.} \gloss{Meaning may.}

% --- Continuity / recurrence
\item \ordic{ret es.} \gloss{Being returns.}
\item \ordic{ret ord.} \gloss{Order returns.}
\item \ordic{ret set.} \gloss{Structure returns.}
\item \ordic{ret mi.} \gloss{Meaning returns.}
\item \ordic{ret mu.} \gloss{Memory returns.}
\item \ordic{ret jur.} \gloss{Law returns.}
\item \ordic{ret gra.} \gloss{Mercy returns.}
\item \ordic{ret šae l.} \gloss{Calm returns.}
\item \ordic{ret ßo.} \gloss{Binding returns.}
\item \ordic{ret vi-te.} \gloss{Witnessing of truth returns.}

% --- Integrity / preservation
\item \ordic{mu ord-te.} \gloss{Memory preserves true order.}
\item \ordic{mu set-te.} \gloss{Memory preserves true structure.}
\item \ordic{mu mi-te.} \gloss{Memory preserves true meaning.}
\item \ordic{mu jur-te.} \gloss{Memory preserves true law.}
\item \ordic{mu gra-te.} \gloss{Memory preserves true mercy.}
\item \ordic{mu šae l-te.} \gloss{Memory preserves true calm.}
\item \ordic{mu ßo-te.} \gloss{Memory preserves true binding.}
\item \ordic{mu vi-te.} \gloss{Memory witnesses truth.}
\item \ordic{mu te-log.} \gloss{Memory holds proof.}
\item \ordic{mu te.} \gloss{Memory aligns with truth.}

% --- Boundaries / domains
\item \ordic{in es.} \gloss{Within being.}
\item \ordic{ex es.} \gloss{Outside being.}
\item \ordic{in set.} \gloss{Within structure.}
\item \ordic{ex set.} \gloss{Outside structure.}
\item \ordic{in log.} \gloss{Within reasoning.}
\item \ordic{ex log.} \gloss{Outside reasoning.}
\item \ordic{in mi.} \gloss{Within meaning.}
\item \ordic{ex mi.} \gloss{Outside meaning.}
\item \ordic{in gra.} \gloss{Within mercy.}
\item \ordic{ex gra.} \gloss{Outside mercy.}

% --- Enforcement / severance (high restraint)
\item \ordic{ša set.} \gloss{Sever structure.}
\item \ordic{ša log.} \gloss{Sever reasoning.}
\item \ordic{ša mi.} \gloss{Sever meaning.}
\item \ordic{ša mu.} \gloss{Sever memory.}
\item \ordic{ša jur.} \gloss{Sever law.}
\item \ordic{ša gra.} \gloss{Sever mercy (rare/forbidden).}
\item \ordic{ša ßo.} \gloss{Sever binding.}
\item \ordic{ša ord.} \gloss{Sever order.}
\item \ordic{ša es.} \gloss{Sever being (extreme boundary).}
\item \ordic{ša te-log.} \gloss{Sever proof (invalidate).}

% --- Restoration by mercy
\item \ordic{gra set.} \gloss{Mercy preserves structure.}
\item \ordic{gra ord.} \gloss{Mercy preserves order.}
\item \ordic{gra mi.} \gloss{Mercy preserves meaning.}
\item \ordic{gra mu.} \gloss{Mercy preserves memory.}
\item \ordic{gra jur.} \gloss{Mercy tempers law.}
\item \ordic{gra ßo.} \gloss{Mercy tempers binding.}
\item \ordic{gra vi-te.} \gloss{Mercy witnesses truth.}
\item \ordic{gra te-log.} \gloss{Mercy accepts proof.}
\item \ordic{gra ret ord.} \gloss{Mercy restores order.}
\item \ordic{gra ret mu.} \gloss{Mercy restores memory.}

% --- Completion / calm axis
\item \ordic{šael es.} \gloss{Calm exists.}
\item \ordic{šael ord.} \gloss{Calm orders.}
\item \ordic{šael mi.} \gloss{Calm clarifies meaning.}
\item \ordic{šael mu.} \gloss{Calm preserves memory.}
\item \ordic{šael vi-te.} \gloss{Calm witnesses truth.}
\item \ordic{šael dur.} \gloss{Calm persists.}
\item \ordic{šael ret.} \gloss{Calm returns.}
\item \ordic{šael gra.} \gloss{Calm enables mercy.}
\item \ordic{šael ord-ret.} \gloss{Calm aligns return-of-order.}
\item \ordic{šael\ldots} \gloss{Calm\ldots (sealed pause).}

% --- Closing canonical checks
\item \ordic{te es dur.} \gloss{True being endures.}
\item \ordic{te ord dur.} \gloss{True order endures.}
\item \ordic{te set dur.} \gloss{True structure endures.}
\item \ordic{te mi dur.} \gloss{True meaning endures.}
\item \ordic{te mu dur.} \gloss{True memory endures.}
\item \ordic{te jur dur.} \gloss{True law endures.}
\item \ordic{te gra dur.} \gloss{True mercy endures.}
\item \ordic{te šae l dur.} \gloss{True calm endures.}
\item \ordic{te ßo dur.} \gloss{True binding endures.}
\item \ordic{ret ord-ret.} \gloss{Order returns remembered.}

% --- Ontological boundaries
\item \ordic{es lin ord.} \gloss{Being moves toward order.}
\item \ordic{es tri ord.} \gloss{Being passes through order.}
\item \ordic{es ne ord.} \gloss{Being remains near order.}
\item \ordic{es far ord.} \gloss{Being is distant from order.}
\item \ordic{ord lin es.} \gloss{Order directs being.}
\item \ordic{ord tri es.} \gloss{Order permeates being.}
\item \ordic{ord ne es.} \gloss{Order is adjacent to being.}
\item \ordic{ord far es.} \gloss{Order is distant from being.}
\item \ordic{mi lin ord.} \gloss{Meaning moves toward order.}
\item \ordic{mi tri ord.} \gloss{Meaning passes through order.}

% --- Relational stability
\item \ordic{ßo lin ord.} \gloss{Binding draws toward order.}
\item \ordic{ßo dur ord.} \gloss{Binding sustains order.}
\item \ordic{ßo ne ord.} \gloss{Binding remains near order.}
\item \ordic{ßo far ord.} \gloss{Binding weakens from order.}
\item \ordic{ßo ret ord.} \gloss{Binding returns to order.}
\item \ordic{ßo set ord.} \gloss{Binding structures order.}
\item \ordic{ßo log ord.} \gloss{Binding reasons order.}
\item \ordic{ßo mi ord.} \gloss{Binding gives meaning to order.}
\item \ordic{ßo mu ord.} \gloss{Binding preserves order in memory.}
\item \ordic{ßo gra ord.} \gloss{Binding tempered by mercy.}

% --- Structural articulation
\item \ordic{set lin ord.} \gloss{Structure moves toward order.}
\item \ordic{set tri ord.} \gloss{Structure passes through order.}
\item \ordic{set ne ord.} \gloss{Structure rests near order.}
\item \ordic{set far ord.} \gloss{Structure drifts from order.}
\item \ordic{set ret ord.} \gloss{Structure returns to order.}
\item \ordic{set ßo ord.} \gloss{Structure binds order.}
\item \ordic{set log ord.} \gloss{Structure reasons order.}
\item \ordic{set mi ord.} \gloss{Structure conveys meaning of order.}
\item \ordic{set mu ord.} \gloss{Structure records order.}
\item \ordic{set jur ord.} \gloss{Structure enforces law.}

% --- Law refinement
\item \ordic{jur lin ord.} \gloss{Law directs toward order.}
\item \ordic{jur tri ord.} \gloss{Law permeates order.}
\item \ordic{jur ne ord.} \gloss{Law stands adjacent to order.}
\item \ordic{jur far ord.} \gloss{Law detached from order (degenerate).}
\item \ordic{jur ret ord.} \gloss{Law restored to order.}
\item \ordic{jur ßo ord.} \gloss{Law binds order.}
\item \ordic{jur log ord.} \gloss{Law reasons order.}
\item \ordic{jur mi ord.} \gloss{Law defines meaning of order.}
\item \ordic{jur mu ord.} \gloss{Law records order.}
\item \ordic{jur gra ord.} \gloss{Law tempered by mercy.}

% --- Reasoning depth
\item \ordic{log lin te.} \gloss{Reasoning moves toward truth.}
\item \ordic{log tri te.} \gloss{Reasoning passes through truth.}
\item \ordic{log ne te.} \gloss{Reasoning remains near truth.}
\item \ordic{log far te.} \gloss{Reasoning drifts from truth.}
\item \ordic{log ret te.} \gloss{Reasoning returns to truth.}
\item \ordic{log ßo te.} \gloss{Reasoning binds to truth.}
\item \ordic{log mi te.} \gloss{Reasoning clarifies truth.}
\item \ordic{log mu te.} \gloss{Reasoning preserves truth in memory.}
\item \ordic{log jur te.} \gloss{Reasoning justifies law.}
\item \ordic{log šae l te.} \gloss{Reasoning calms truth.}

% --- Mercy axis (still High Ordic, sparse)
\item \ordic{gra lin ord.} \gloss{Mercy leads toward order.}
\item \ordic{gra ne ord.} \gloss{Mercy rests near order.}
\item \ordic{gra far ord.} \gloss{Mercy withdrawn from order.}
\item \ordic{gra ret ord.} \gloss{Mercy returns order.}
\item \ordic{gra ßo ord.} \gloss{Mercy binds order gently.}
\item \ordic{gra log ord.} \gloss{Mercy reasons order.}
\item \ordic{gra mi ord.} \gloss{Mercy restores meaning to order.}
\item \ordic{gra mu ord.} \gloss{Mercy preserves order in memory.}
\item \ordic{gra jur ord.} \gloss{Mercy tempers law.}
\item \ordic{gra šae l ord.} \gloss{Mercy calms order.}

% --- Severance boundaries
\item \ordic{ša lin ord.} \gloss{Severance moves toward order (controlled cut).}
\item \ordic{ša tri ord.} \gloss{Severance passes through order.}
\item \ordic{ša ne ord.} \gloss{Severance near order (surgical).}
\item \ordic{ša far ord.} \gloss{Severance distant from order (destructive).}
\item \ordic{ša ret ord.} \gloss{Severance reversed.}
\item \ordic{ša ßo ord.} \gloss{Sever binding to order.}
\item \ordic{ša log ord.} \gloss{Sever reasoning from order.}
\item \ordic{ša mi ord.} \gloss{Sever meaning from order.}
\item \ordic{ša mu ord.} \gloss{Sever memory from order.}
\item \ordic{ša jur ord.} \gloss{Sever law from order (collapse).}

% --- Calm completion
\item \ordic{šael lin ord.} \gloss{Calm leads toward order.}
\item \ordic{šael tri ord.} \gloss{Calm permeates order.}
\item \ordic{šael ne ord.} \gloss{Calm rests near order.}
\item \ordic{šael far ord.} \gloss{Calm distant from order.}
\item \ordic{šael ret ord.} \gloss{Calm returns order.}
\item \ordic{šael ßo ord.} \gloss{Calm binds order.}
\item \ordic{šael log ord.} \gloss{Calm reasons order.}
\item \ordic{šael mi ord.} \gloss{Calm clarifies meaning of order.}
\item \ordic{šael mu ord.} \gloss{Calm preserves order in memory.}
\item \ordic{šael gra ord.} \gloss{Calm enables mercy in order.}

% --- Final stabilizers
\item \ordic{te lin ord.} \gloss{Truth leads toward order.}
\item \ordic{te ne ord.} \gloss{Truth rests near order.}
\item \ordic{te far ord.} \gloss{Truth distant from order (crisis).}
\item \ordic{te ret ord.} \gloss{Truth restores order.}
\item \ordic{te ßo ord.} \gloss{Truth binds order.}
\item \ordic{te log ord.} \gloss{Truth reasons order.}
\item \ordic{te mi ord.} \gloss{Truth gives meaning to order.}
\item \ordic{te mu ord.} \gloss{Truth preserves order in memory.}
\item \ordic{te jur ord.} \gloss{Truth grounds law.}
\item \ordic{te šae l ord.} \gloss{Truth calms order.}

% --- Ontological closure
\item \ordic{es dur ord.} \gloss{Being endures as order.}
\item \ordic{es dur te.} \gloss{Being endures as truth.}
\item \ordic{es dur mi.} \gloss{Being endures as meaning.}
\item \ordic{es dur mu.} \gloss{Being endures as memory.}
\item \ordic{es dur ßo.} \gloss{Being endures as binding.}
\item \ordic{es dur jur.} \gloss{Being endures as law.}
\item \ordic{es dur gra.} \gloss{Being endures as mercy.}
\item \ordic{es dur šae l.} \gloss{Being endures as calm.}
\item \ordic{es dur set.} \gloss{Being endures as structure.}
\item \ordic{es dur log.} \gloss{Being endures as reasoning.}

% --- Totality statements
\item \ordic{al es ord.} \gloss{All being is ordered.}
\item \ordic{al ord te.} \gloss{All order is true.}
\item \ordic{al mi te.} \gloss{All meaning is true.}
\item \ordic{al mu te.} \gloss{All memory is true.}
\item \ordic{al set ord.} \gloss{All structures are ordered.}
\item \ordic{al log te.} \gloss{All reasoning is true.}
\item \ordic{al jur ord.} \gloss{All law is ordered.}
\item \ordic{al ßo ord.} \gloss{All binding is ordered.}
\item \ordic{al gra ord.} \gloss{All mercy is ordered.}
\item \ordic{al šae l ord.} \gloss{All calm is ordered.}

% --- Identity / equivalence
\item \ordic{un es ord.} \gloss{The one is ordered being.}
\item \ordic{du es ord.} \gloss{The other is ordered being.}
\item \ordic{un du eq.} \gloss{The one and the other are equal.}
\item \ordic{un du dif.} \gloss{The one and the other differ.}
\item \ordic{eq es ord.} \gloss{Equivalence aligns being and order.}
\item \ordic{dif es ord.} \gloss{Difference aligns being and order.}
\item \ordic{eq mi.} \gloss{Meanings are equivalent.}
\item \ordic{dif mi.} \gloss{Meanings are distinct.}
\item \ordic{eq ord.} \gloss{Orders are equivalent.}
\item \ordic{dif ord.} \gloss{Orders are distinct.}

% --- Final stability lattice
\item \ordic{set ord-ret.} \gloss{Structure of return-of-order.}
\item \ordic{log ord-ret.} \gloss{Reasoning of return-of-order.}
\item \ordic{mi ord-ret.} \gloss{Meaning of return-of-order.}
\item \ordic{mu ord-ret.} \gloss{Memory of return-of-order.}
\item \ordic{ßo ord-ret.} \gloss{Binding of return-of-order.}
\item \ordic{jur ord-ret.} \gloss{Law of return-of-order.}
\item \ordic{gra ord-ret.} \gloss{Mercy of return-of-order.}
\item \ordic{šael ord-ret.} \gloss{Calm of return-of-order.}
\item \ordic{te ord-ret.} \gloss{Truth of return-of-order.}
\item \ordic{es ord-ret.} \gloss{Being of return-of-order.}

% --- Witnessed completeness
\item \ordic{vi te es.} \gloss{Truth of being is witnessed.}
\item \ordic{vi te ord.} \gloss{Truth of order is witnessed.}
\item \ordic{vi te set.} \gloss{Truth of structure is witnessed.}
\item \ordic{vi te mi.} \gloss{Truth of meaning is witnessed.}
\item \ordic{vi te mu.} \gloss{Truth of memory is witnessed.}
\item \ordic{vi te ßo.} \gloss{Truth of binding is witnessed.}
\item \ordic{vi te jur.} \gloss{Truth of law is witnessed.}
\item \ordic{vi te gra.} \gloss{Truth of mercy is witnessed.}
\item \ordic{vi te šae l.} \gloss{Truth of calm is witnessed.}
\item \ordic{vi te log.} \gloss{Truth of reasoning is witnessed.}

% --- Completion statements (not the Final Utterance)
\item \ordic{dur al es.} \gloss{All being continues.}
\item \ordic{dur al ord.} \gloss{All order continues.}
\item \ordic{dur al mi.} \gloss{All meaning continues.}
\item \ordic{dur al mu.} \gloss{All memory continues.}
\item \ordic{dur al ßo.} \gloss{All binding continues.}
\item \ordic{dur al jur.} \gloss{All law continues.}
\item \ordic{dur al gra.} \gloss{All mercy continues.}
\item \ordic{dur al šae l.} \gloss{All calm continues.}
\item \ordic{dur al set.} \gloss{All structure continues.}
\item \ordic{dur al log.} \gloss{All reasoning continues.}

% --- Sealed closing formulas
\item \ordic{ret al ord.} \gloss{All order returns.}
\item \ordic{ret al es.} \gloss{All being returns.}
\item \ordic{ret al mi.} \gloss{All meaning returns.}
\item \ordic{ret al mu.} \gloss{All memory returns.}
\item \ordic{ret al ßo.} \gloss{All binding returns.}
\item \ordic{ret al jur.} \gloss{All law returns.}
\item \ordic{ret al gra.} \gloss{All mercy returns.}
\item \ordic{ret al šae l.} \gloss{All calm returns.}
\item \ordic{ret al set.} \gloss{All structure returns.}
\item \ordic{ret al log.} \gloss{All reasoning returns.}

% --- Final seal (never spoken as completion)
\item \ordic{ord-ret.} \gloss{Order remembered (title/seal).}
\item \ordic{te\ldots} \gloss{Truth\ldots (sealed pause).}
\item \ordic{šael\ldots} \gloss{Calm\ldots (sealed pause).}
\item \ordic{dur.} \gloss{It continues.}
\item \ordic{es.} \gloss{Being.}

\end{enumerate}

\end{multicols}

\vspace{6pt}\hrule\vspace{4pt}
{\small Notes: Hyphens indicate compounds. Case suffixes appear as \texttt{-in,-ex,-lin,-ne,-far,-ßo,-we,-na,-il,-ri,-ha}. In strict High Ordic chant, maintain Mood--Tense--Aspect--Verb ordering when applicable.}

\end{document}
