\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{xparse}
\usepackage[utf8]{inputenc}

% ---------- Global script feel ----------
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\raggedright

% ---------- Counters ----------
\newcounter{gnpage}
\newcounter{gnpanel}[gnpage]
\renewcommand{\thegnpage}{\arabic{gnpage}}
\renewcommand{\thegnpanel}{\arabic{gnpanel}}

% ---------- Tunable indents ----------
\newlength{\GNDialogueLeft}
\newlength{\GNDialogueRight}
\setlength{\GNDialogueLeft}{0.22\textwidth}
\setlength{\GNDialogueRight}{0.22\textwidth}

% ---------- Title block ----------
\NewDocumentCommand{\GNTitle}{ m m O{\today} }{%
  \thispagestyle{empty}
  \vspace*{0.18\textheight}
  {\Large\bfseries #1\par}
  \vspace{0.5em}
  {\large #2\par}
  \vspace{1em}
  {\itshape #3\par}
  \vfill
  \newpage
}

% ---------- Page ----------
% Usage: \begin{GNPage}[optional page subtitle] ... \end{GNPage}
\NewDocumentEnvironment{GNPage}{ O{} }{%
  \refstepcounter{gnpage}
  \setcounter{gnpanel}{0}
  \vspace{1em}
  {\bfseries PAGE \thegnpage%
    \IfValueT{#1}{ - #1}%
  }\par
  \vspace{0.5em}
}{%
  \vspace{1em}
}

% ---------- Scene / Environment ----------
\NewDocumentCommand{\Env}{ +m }{%
  {\bfseries\MakeUppercase{#1}}\par
}

% ---------- Panel ----------
% Starred version doesn't increment (useful for splash / non-numbered beats)
% Usage: \Panel[optional label]{description}
\NewDocumentCommand{\Panel}{ s O{} +m }{%
  \par\vspace{0.5em}
  \IfBooleanTF{#1}{%
    {\bfseries PANEL%
      \IfValueT{#2}{\ \textnormal{(\MakeUppercase{#2})}}%
    }\par
  }{%
    \refstepcounter{gnpanel}
    {\bfseries PANEL \thegnpanel%
      \IfValueT{#2}{\ \textnormal{(\MakeUppercase{#2})}}%
    }\par
  }%
  #3\par
}

% ---------- Action ----------
\NewDocumentCommand{\Action}{ +m }{%
  #1\par
}

% ---------- Caption / Narration ----------
\NewDocumentCommand{\Caption}{ +m }{%
  {\itshape CAPTION: #1}\par
}
\NewDocumentCommand{\Narration}{ +m }{%
  {\itshape NARRATION: #1}\par
}

% ---------- SFX ----------
\NewDocumentCommand{\SFX}{ +m }{%
  {\bfseries SFX: \MakeUppercase{#1}}\par
}

% ---------- Transition ----------
\NewDocumentCommand{\Transition}{ +m }{%
  \hfill{\bfseries \MakeUppercase{#1}}\par
}

% ---------- Note to self (won't look like a panel element) ----------
\NewDocumentCommand{\Note}{ +m }{%
  {\small\itshape [NOTE: #1]}\par
}

% ---------- Dialogue ----------
% Usage: \Dialogue[optional parenthetical]{CHARACTER}{line(s)}
\NewDocumentCommand{\Dialogue}{ O{} m +m }{%
  {\par\noindent\centering\textsc{#2}\par}
  \begin{list}{}{%
    \setlength{\leftmargin}{\GNDialogueLeft}%
    \setlength{\rightmargin}{\GNDialogueRight}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parsep}{0pt}%
    \setlength{\topsep}{2pt}%
  }%
    \item
      \IfValueT{#1}{\emph{(#1)}\par}
      #3
  \end{list}%
}

% Optional quick beats
\NewDocumentCommand{\Beat}{}{%
  {\itshape (beat)}\par
}

\begin{document}

\GNTitle{KA\'{I}VEN | Chapter One}{noumena ashitaka hundim\"{a}gi-mei}[Draft \small{(v0.1)} - \today]

\Env{EXT. SPACE CLOSE TO BLACKHOLE}

\begin{GNPage}[THE MOUTH OF NIGHT]


\Panel[WIDE, ESTABLISHING]{\textbf{Exterior}. The ship hangs like a wounded insect in the dark. Plates peeled back. Scars of plasma burns. One engine dead, another coughing blue.\\
	Ahead: a \textbf{gargantuan black hole}, not just darkness but a presence an eye without eyelids. The accretion disk smears light into a bright, violent halo. Distant debris arcs inward like prayer beads being pulled from a snapped string.}

\Caption{We have run so long that even the stars feel tired of watching us.}

\Caption{The ship is ruined, its bones cracked, its skin torn, and there is no kindly harbor in the void.}

\Panel[INTERIOR, WIDE]{\textbf{Bridge.} The crew of six. Exhaustion is everywhere: slumped shoulders, trembling hands, bruises under eyes, dried blood at someone's hairline. Consoles flicker with dying light. A warning klaxon is either silent now or too constant to notice.\\

Only Z\'{A}ALI is upright, facing forward, hands on the console like he's holding the world in place.}

\Caption{They trusted me. That is the sharpest thing in the room.}

\Caption{Not the alarms. Not the heat bleeding from the vents. Not even the black mouth waiting outside.}

\Panel[CLOSE-UP: DAMAGE / DETAILS]{A cracked screen shows: \textbf{HULL INTEGRITY: 12\%}.
A manual patch welded over a breach. A leaking cable sparking. A tiny photo or token taped beside a gauge, something human and stubborn.}

\Caption{How many times can a man make the same choice and still call it bravery?}

\Caption{How many times can he lead others into fire and name it \textbf{the only way}?}

\Panel[CLOSE-UP: CREW FACES, ONE BY ONE]{
Quick cuts:

\begin{itemize}
  \item One crew member asleep upright, jaw clenched like resisting a nightmare.
  \item Another quietly crying without sound.
  \item Someone's hand hovering over an eject lever, then falling away.
  \item Someone's staring at the black hole on the forward display as if looking at an old god.
\end{itemize}
}

\Caption{I wanted a simpler life once. I remember that version of me like a half-forgotten song.}

\Caption{A house. A door that locks. Mornings that mean bread and not battle.}

\Caption{But I stepped onto this road, again, and again, and each time I told myself it was the last.}

\Panel[INTERIOR, LOW ANGLE ON PROTAGONIST / FORWARD WINDOW]{The forward window dominates. The black hole swells larger now, a slow certainty. Light bends wrong. The ship is being \textit{drawn}, not by engines, but by fate wearing physics as a mask. Z\'{A}ALI's face is lit by the warped glow-half saint, half criminal.}

\Caption{Why do I keep doing this?}

\Caption{Why does it feel... not like living... but like repeating?}

\Caption{As if some unseen hand has set the pieces upon the board before I ever touch them.}

\Caption{As if the world has grooves worn into it, and my feet find them no matter how I fight.}

\Panel[FULL-BLEED, COSMIC / SURREAL]{The ship tiny against the black hole's majesty. The accretion disk looks like a burning crown. The stars behind smear into thin lines, as though reality itself is being combed toward the dark. In the chaos, there is a strange calm like a theater right before the curtain falls.}

\Caption{I tell myself I am choosing.}

\Caption{I tell myself there is meaning in the struggle, and honor in the scars.}

\Caption{But some nights, like this one, I swear I can hear the machinery behind the sky.}

\Caption{A faint clicking. A silent counting.}

\Caption{And the cruelest thought of all is this:}

\Caption{If this ends the way it always ends... then what, in the name of all the lights that ever were, was I meant to learn?}

\end{GNPage}

\begin{GNPage}[THE WEIGHT OF LIVING]

\Panel[WIDE, INTERIOR / BRIDGE]{The bridge is dim, lit by failing consoles and the warped glow from the black hole. The crew are present but distant, like ghosts still operating machinery out of habit. Micro-debris ticks softly against the hull.}

\Caption{Pain is not a stranger. It is the oldest crewman aboard every ship I have ever flown.}

\Caption{It sits quietly in the corner while we laugh, and it does not laugh with us.}

\Panel[MEDIUM, MONTAGE DETAIL]{A small \textbf{ordinary} detail amid catastrophe: a dented mug floating in low gravity, a charm tied to a console; a torn sleeve stitched with crude thread.}

\Caption{It watches us eat, and sleep, and make promises we do not understand the cost of.}

\Caption{Then, when the alarms begin, it rises as if it has been waiting for its proper work.}

\Panel[WIDE, EXTERIOR / SHIP AGAINST THE HOLE]{The ship is tiny. The black hole is not \textbf{big}, it's unfair. Light arcs like frightened animals. The accretion disk burns like a crown of judgment.}

\Caption{They tell you existence is a gift.}

\Caption{If so, it is a gift wrapped in thorns, and the blood on the ribbon is usually yours.}

\Panel[CLOSE-UP, PROTAGONIST'S FACE / REFLECTION]{Close on Z\'{A}ALI's eyes, wet with exhaustion but not with surrender. In the glass of the forward viewport, the black hole distorts his reflection into something older.}

\Caption{I have crossed a hundred skies. I have tasted a hundred triumphs.}

\Caption{None of them were clean.}

\Panel[MONTAGE STRIP, THREE QUICK IMAGES IN ONE PANEL OR THREE TIGHT PANELS]{A rapid memory montage:\\

A past victory: the crew laughing under alien lanterns.\\

The same place later: scorched metal, bodies covered.\\

Z\'{A}ALI alone, staring at his hands as if they're guilty.}

\Caption{Victory is only a word we use when grief has not yet caught up.}

\end{GNPage}

\begin{GNPage}[THE LEDGER OF THE LOST]

\Panel[CLOSE-UP, HANDS / LIST]{Z\'{A}ALI's gloved hand hovers over a console. \\
On-screen: a scrolling list of names/callsigns, some blurred or corrupted, some struck through.}
\Caption{I remember faces the way sailors remember storms: not by name, but by what they took.}

\Panel[FLASHBACK, MEDIUM]{A crew member, younger, grinning mid-sentence-caught in a candid moment.}

\Caption{A smile that went out mid-sentence.}

\Panel[FLASHBACK, ACTION SNAPSHOT]{Chaos: a breach, decompression, a hand reaching through smoke and frost toward Z\'{A}ALI.}

\Caption{A hand reaching for mine and finding only emptiness.}

\Panel[FLASHBACK, QUIET HORROR]{A helmet tumbles slowly through space, turning end over end. In its visor: a frozen reflection of stars.}

\Caption{A voice cut short, like a song struck from the throat.}

\Panel[WIDE, FUNERAL THAT ISN'T A FUNERAL]{The crew eject a wrapped body into space because there is nowhere to bury anyone. No ceremony, only silence and the indifferent starlight.}

\Caption{I have buried companions in places that do not allow graves.}

\Caption{I have left them drifting where no planet will ever claim them.}

\Panel[INTERIOR, PRESENT / BRIDGE, CREW IN FRAME]{Back to the broken bridge. The crew are alive, barely, and Z\'{A}ALI looks at them as if counting blessings he doesn't deserve.}

\Caption{I have watched light itself bend around their dying, and still it did not make it holy.}

\Caption{It is a strange thing, to be called \textbf{leader}, when all your path is marked by those who followed you and did not return.}

\Caption{Some nights I wonder if I am not a captain at all, only a collector of endings.}

\Caption{And the worst of it is not that I lost them.}

\Caption{The worst of it is that I can still feel them, like phantom limbs of the soul, aching where they used to be.}

\end{GNPage}


\begin{GNPage}[THE ONES WHO COME BACK WRONG]

\Panel[WIDE, STRANGE, DREAMLIKE]{A surreal montage panel: Z\'{A}ALI in a corridor that feels half-ship, half-memory. The lighting is wrong. The perspective subtly impossible.}

\Caption{There are losses you can accept: the swift, clean kind, where the universe is merely cruel.}

\Caption{And then there are the other losses, the ones that return.}

\Panel[MONTAGE]{A bustling spaceport crowd. Z\'{A}ALI sees someone with a familiar posture, same stance as a dead friend, but in a totally different body: different age, different species, different fashion.}

\Caption{They do not return as they were.}

\Caption{A familiar laugh in a stranger's mouth.}

\Panel[MONTAGE]{A marketplace. Z\'{A}ALI catches a gaze, recognition flashes like lightning, then the person turns away, expression empty, moving on.}

\Caption{A gaze that once knew you, now sliding past as if you are smoke.}

\Panel[MONTAGE, NEAR-MISS]{A door closes. Z\'{A}ALI arrives one second too late. A departing shuttle lifts off. He stands beneath its exhaust wash like someone being erased.}

\Caption{I have found them again, sometimes years later, sometimes worlds away, and I have missed them by a single choice, a single hour, a single wrong door.}

\Panel[WIDE, SEARCH ACROSS WORLDS]{A cosmic travel montage: different planets, different moons, different ruins; Z\'{A}ALI always arriving after the fact. His silhouette repeats like a cursed refrain.}

\Caption{Do you know what it is to search for a person across the breadth of the cosmos, armed only with grief and intuition?}

\Caption{To see a silhouette in a crowd and feel your heart lunge like a dog straining at the leash only to be disappointed, again.}

\Panel[CLOSE-UP, THE NAME FALLS]{Z\'{A}ALI speaks a name (we don't show the word balloon clearly; we can crop it or blur it), and the stranger looks at him kindly, politely, without recognition.}

\Caption{To speak a name you are not sure you are allowed to speak, and watch it fall to the floor between you like a dead thing.}

\Caption{They look at me with polite eyes and do not know why my voice breaks.}

\Caption{They do not remember the promises we made under burning moons.}

\Caption{They do not remember dying for me.}

\Caption{And I... I remember too much.}

\end{GNPage}

\begin{GNPage}[THE SHAPE OF THE WAR BEYOND THIS WAR]

\Panel[PRESENT, SHIP SHUDDERS]{The ship groans. Panels peel back. A cable whips in slow-motion. The black hole's pull is now undeniable, everyone feels the tilt of reality.}

\Caption{A man can endure hardship. He can even endure horror.}

\Caption{What breaks him is the pattern.}

\Panel[VISUAL METAPHOR, THE PATTERN]{A montage of repeated scenes in different contexts: the same hand reaching, the same alarm light, the same goodbye like d\'{e}j\`{a} vu rendered as imagery.}

\Caption{The sense that suffering is not an accident, but a design.}

\Caption{That someone, somewhere, decided pain would be the price of being.}

\Panel[COSMIC POV, SOMETHING LOOKS BACK]{The black hole is framed like an eye. The starfield warps into something that resembles a staring face for a moment, just long enough to unsettle.}

\Caption{I have stared into the dark long enough to suspect it stares back.}

\Caption{As the Architect, you were not meant to have any knowledge of his existence; for his name was withheld, and his being kept from your sight.}

\Panel[THE SEAMS]{Z\'{A}ALI's gloved fingers touch the forward glass or a holo-map; the image briefly glitches, angles that don't make sense, lines repeating, star positions snapping like a grid.}

\Caption{There are days I feel the seams of the world beneath my fingertips.}

\Caption{Thin places. Wrong angles. Moments that repeat with a familiar cadence as if the cosmos has favorite lines.}

\Panel[PRESENT, THE REALIZATION RETURNS]{The crew in the background: one asleep, one praying, one shaking, one furious. Z\'{A}ALI is still, listening to something nobody else hears.}

\Caption{And I know. I know in the marrow of me, that there will be another battle after this one.}

\Panel[THE TRUE ENEMY]{A shadowy suggestion: not a person, not a monster, more like a structure behind reality. Could be geometric shapes, looming silhouettes in the distortion, or symbols hidden in the gravitational lensing.}

\Caption{Not against raiders, nor tyrants, nor the thousand petty empires that rise and rot among the stars.}

\Caption{But against the hands that set the rules of the game and call it \textbf{reality}.}

\Caption{Against the ones who made cruelty feel inevitable.}

\Caption{I do not speak their names. Names have power, and I cannot afford to give them more.}

\Caption{Yet even here, with a black hole waiting like a mouth, I feel the old summons returning.}

\Caption{The road opening beneath my feet.}

\Caption{The wheel turning.}

\Caption{Because it always turns.}

\end{GNPage}

\begin{GNPage}[THE VOW]

\Panel[QUIET, CREW IN STILLNESS]{The crew are frozen in that strange end-of-everything calm. One person's hand rests on another's shoulder. No speeches. Just breathe.}

\Caption{If I were wise, I would let the dark take us gently.}

\Caption{I would let my bones rest. I would let my mind fall silent.}

\Panel[THE TEMPTATION OF PEACE]{A panel that almost looks peaceful: stars like snow, the ship like a cradle then the edge of the black hole intrudes like a blade.}

\Caption{I would stop trying to bargain with fate as if fate had ever once been fair.}

\Caption{But I cannot.}

\Panel[MONTAGE THREAD RETURNS]{A montage of the crew across other lives: same eyes in different faces; the same gesture repeated across time. Keep it subtle, felt, not explained.}

\Caption{Not while there is even a single thread that might be pulled loose.}

\Caption{Not while the ones I love keep slipping through my hands and returning to me as strangers.}

\Panel[THE HERESY]{Z\'{A}ALI rises. The posture changes: from exhausted survivor to someone making a decision with teeth. The black hole glow crowns him in warped light.}

\Caption{Not while the cosmos pretends it is innocent.}

\Caption{I have bled for this existence. We all have.}

\Caption{And if suffering is the law, then I will become its heresy.}

\Panel[FULL-BLEED ENDING BEAT]{The ship is nearly at the threshold. In the blackness, a hairline of something impossible: a faint geometry, like an edge, like a seam, a suggestion of a door inside the mouth of night.}

\Caption{Vek zaar varm. / Let the dark come close.}

\Caption{Se senar dornun zaar l\'{i}rserm. / Let it show me the edge of the world.}

\Caption{Kiv vekra mureš bol voš ve, za seun sarnra. / If there is a door hidden in the mouth of night, I will find it.}

\Caption{Va za bolun tokra... / And I will knock...}
	
\end{GNPage}

\end{document}
