\documentclass[10pt,a4paper]{article}

\usepackage[a4paper,margin=0.6in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{multicol}
\usepackage{enumitem}

\setlength{\parindent}{0pt}
\setlength{\parskip}{2pt}
\setlist[itemize]{leftmargin=*, itemsep=2pt}

\newcommand{\ordic}[1]{\textbf{#1}}
\newcommand{\gloss}[1]{\hfill{\small\textsf{#1}}}

\begin{document}

{\LARGE \textbf{Low Ordic Lexicon — Core Words}}\\
{\small Spoken register. Erosion, emotion, immediacy, survival.}

\vspace{6pt}\hrule\vspace{8pt}

\begin{multicols}{2}

\begin{itemize}

% --- Presence / immediacy
\item \ordic{es} \gloss{I am / it is}
\item \ordic{nu} \gloss{now}
\item \ordic{nu es} \gloss{here, now}
\item \ordic{es nu} \gloss{I’m here}
\item \ordic{dur} \gloss{still / staying}
\item \ordic{dur nu} \gloss{still here}
\item \ordic{lin} \gloss{go / move}
\item \ordic{lin nu} \gloss{go now}
\item \ordic{ret} \gloss{come back}
\item \ordic{ret nu} \gloss{come back now}

% --- Feeling states (compressed)
\item \ordic{shae} \gloss{calm / okay}
\item \ordic{shae nu} \gloss{it’s okay now}
\item \ordic{qe} \gloss{unsure}
\item \ordic{qe nu} \gloss{I don’t know}
\item \ordic{gra} \gloss{kind / gentle}
\item \ordic{gra nu} \gloss{be kind}
\item \ordic{na} \gloss{no}
\item \ordic{na nu} \gloss{not now}
\item \ordic{te} \gloss{true / really}
\item \ordic{te es} \gloss{it’s real}

% --- Bonds / people
\item \ordic{xo} \gloss{us / bond}
\item \ordic{xo es} \gloss{we are}
\item \ordic{xo nu} \gloss{together now}
\item \ordic{xo dur} \gloss{we’re still together}
\item \ordic{mu} \gloss{remember}
\item \ordic{mu nu} \gloss{remember this}
\item \ordic{mu xo} \gloss{remember us}
\item \ordic{mu shae} \gloss{gentle memory}
\item \ordic{mu na} \gloss{forget}
\item \ordic{mu ret} \gloss{remember again}

% --- Speech shortcuts
\item \ordic{es-te} \gloss{yes (true being)}
\item \ordic{na-te} \gloss{no (not true)}
\item \ordic{es-es} \gloss{okay, okay}
\item \ordic{na-na} \gloss{no, stop}
\item \ordic{qe-qe} \gloss{uh\ldots}
\item \ordic{dur-dur} \gloss{hold on}
\item \ordic{ret-ret} \gloss{please come back}
\item \ordic{lin-lin} \gloss{go on}
\item \ordic{shae-shae} \gloss{calm down}
\item \ordic{xo-xo} \gloss{we’re okay}

% --- Soft imperatives
\item \ordic{lin shae} \gloss{go gently}
\item \ordic{dur shae} \gloss{stay calm}
\item \ordic{ret shae} \gloss{come back safe}
\item \ordic{gra lin} \gloss{let it go}
\item \ordic{gra dur} \gloss{stay kindly}
\item \ordic{na lin} \gloss{don’t go}
\item \ordic{na dur} \gloss{don’t stay}
\item \ordic{na qe} \gloss{don’t worry}
\item \ordic{te mu} \gloss{remember truly}
\item \ordic{te xo} \gloss{trust us}

% --- Everyday truth
\item \ordic{es shae} \gloss{I’m okay}
\item \ordic{es qe} \gloss{I don’t know}
\item \ordic{es gra} \gloss{I care}
\item \ordic{es mu} \gloss{I remember}
\item \ordic{es na} \gloss{I can’t}
\item \ordic{es lin} \gloss{I’m going}
\item \ordic{es ret} \gloss{I’m back}
\item \ordic{es dur} \gloss{I’m staying}
\item \ordic{es xo} \gloss{I’m with you}
\item \ordic{es nu\ldots} \gloss{I’m here\ldots}

% --- Child-speech fossilizations
\item \ordic{nuna} \gloss{not now}
\item \ordic{shanu} \gloss{it’s okay}
\item \ordic{mumu} \gloss{memory / story}
\item \ordic{xonu} \gloss{together}
\item \ordic{duru} \gloss{stay}
\item \ordic{lina} \gloss{go away}
\item \ordic{rete} \gloss{come back}
\item \ordic{qenu} \gloss{maybe}
\item \ordic{granu} \gloss{please}
\item \ordic{es\ldots} \gloss{being\ldots}

% --- Conflict / refusal
\item \ordic{na xo} \gloss{leave me alone}
\item \ordic{na shae} \gloss{not okay}
\item \ordic{na te} \gloss{that’s not true}
\item \ordic{na mu} \gloss{I don’t remember / don’t want to}
\item \ordic{na ret} \gloss{don’t come back}
\item \ordic{na lin} \gloss{don’t go}
\item \ordic{na dur} \gloss{don’t stay}
\item \ordic{na es} \gloss{I’m not}
\item \ordic{na al} \gloss{never / not everything}
\item \ordic{na\ldots} \gloss{no\ldots}

% --- Fear / vulnerability
\item \ordic{qe shae} \gloss{I’m scared}
\item \ordic{qe es} \gloss{what am I?}
\item \ordic{qe xo} \gloss{are we okay?}
\item \ordic{qe ret} \gloss{will you come back?}
\item \ordic{qe lin} \gloss{should I go?}
\item \ordic{qe dur} \gloss{should I stay?}
\item \ordic{qe na} \gloss{maybe not}
\item \ordic{qe nu} \gloss{I don’t know now}
\item \ordic{qe\ldots} \gloss{um\ldots}
\item \ordic{qe-qe} \gloss{hesitation}

% --- Apology / repair
\item \ordic{gra} \gloss{sorry / kindness}
\item \ordic{gra na} \gloss{I’m sorry}
\item \ordic{gra xo} \gloss{sorry for us}
\item \ordic{gra mu} \gloss{sorry, I remember}
\item \ordic{gra shae} \gloss{please calm}
\item \ordic{gra dur} \gloss{please stay}
\item \ordic{gra ret} \gloss{please come back}
\item \ordic{gra nu} \gloss{please, now}
\item \ordic{gra\ldots} \gloss{sorry\ldots}
\item \ordic{gra-gra} \gloss{very sorry}

% --- Play / teasing
\item \ordic{lin lin} \gloss{go on!}
\item \ordic{ret-ret} \gloss{come back!}
\item \ordic{xo xo} \gloss{we’re good}
\item \ordic{shae shae} \gloss{relax}
\item \ordic{na-na} \gloss{stop it}
\item \ordic{qe-qe} \gloss{maybe maybe}
\item \ordic{es-es} \gloss{yeah yeah}
\item \ordic{dur-dur} \gloss{wait wait}
\item \ordic{mu-mu} \gloss{story time}
\item \ordic{nu-nu} \gloss{right now}

% --- Trust / reassurance
\item \ordic{te xo} \gloss{trust us}
\item \ordic{te shae} \gloss{it’s okay}
\item \ordic{te dur} \gloss{it will last}
\item \ordic{te ret} \gloss{I’ll come back}
\item \ordic{te lin} \gloss{I’ll go}
\item \ordic{te mu} \gloss{I remember}
\item \ordic{te es} \gloss{I’m real}
\item \ordic{te nu} \gloss{right now}
\item \ordic{te\ldots} \gloss{really\ldots}
\item \ordic{te-te} \gloss{promise}

% --- Small togetherness
\item \ordic{xo nu} \gloss{together now}
\item \ordic{xo shae} \gloss{we’re okay}
\item \ordic{xo dur} \gloss{we’re staying}
\item \ordic{xo ret} \gloss{we’re back}
\item \ordic{xo mu} \gloss{we remember}
\item \ordic{xo na} \gloss{we can’t}
\item \ordic{xo lin} \gloss{we’re going}
\item \ordic{xo es} \gloss{we are}
\item \ordic{xo\ldots} \gloss{us\ldots}
\item \ordic{xo-xo} \gloss{we’re okay}

% --- Childlike inventions
\item \ordic{shanu} \gloss{it’s fine}
\item \ordic{nanu} \gloss{no more}
\item \ordic{qena} \gloss{maybe}
\item \ordic{duru} \gloss{stay}
\item \ordic{xona} \gloss{don’t touch}
\item \ordic{muna} \gloss{forget it}
\item \ordic{lina} \gloss{go away}
\item \ordic{rete} \gloss{come back}
\item \ordic{grana} \gloss{please}
\item \ordic{nu\ldots} \gloss{now\ldots}

% --- Love / closeness
\item \ordic{xo shae} \gloss{we’re okay together}
\item \ordic{xo gra} \gloss{I care about us}
\item \ordic{xo dur} \gloss{we’re staying}
\item \ordic{xo mu} \gloss{I remember us}
\item \ordic{xo es} \gloss{we are}
\item \ordic{xo nu} \gloss{here together}
\item \ordic{xo te} \gloss{I trust us}
\item \ordic{xo lin} \gloss{we’ll go together}
\item \ordic{xo ret} \gloss{we’ll come back}
\item \ordic{xo\ldots} \gloss{us\ldots}

% --- Loss / absence
\item \ordic{na xo} \gloss{we’re gone}
\item \ordic{xo na} \gloss{we can’t}
\item \ordic{mu na} \gloss{I forgot / it’s gone}
\item \ordic{ret na} \gloss{won’t come back}
\item \ordic{dur na} \gloss{can’t stay}
\item \ordic{shae na} \gloss{not okay}
\item \ordic{gra na} \gloss{no kindness left}
\item \ordic{es na} \gloss{I’m not anymore}
\item \ordic{al na} \gloss{nothing left}
\item \ordic{na\ldots} \gloss{no\ldots}

% --- Grief / quiet pain
\item \ordic{mu xo na} \gloss{I lost us}
\item \ordic{mu shae na} \gloss{the memory hurts}
\item \ordic{mu dur na} \gloss{the memory won’t stay}
\item \ordic{qe xo na} \gloss{why did we break?}
\item \ordic{qe es na} \gloss{why am I gone?}
\item \ordic{qe ret na} \gloss{why won’t you come back?}
\item \ordic{qe nu\ldots} \gloss{why now\ldots}
\item \ordic{dur shae na} \gloss{it hurts to stay}
\item \ordic{ret shae na} \gloss{it hurts to return}
\item \ordic{shae\ldots} \gloss{okay\ldots (not really)}

% --- Humor / teasing
\item \ordic{es-es} \gloss{yeah right}
\item \ordic{na-na} \gloss{no way}
\item \ordic{qe-qe} \gloss{maybe maybe}
\item \ordic{lin lin} \gloss{go on then}
\item \ordic{dur-dur} \gloss{hold your horses}
\item \ordic{ret-ret} \gloss{get back here}
\item \ordic{xo-xo} \gloss{we’re fine}
\item \ordic{shae-shae} \gloss{chill}
\item \ordic{mu-mu} \gloss{story time}
\item \ordic{nu-nu} \gloss{right now, geez}

% --- Soft profanity (Ordic-style)
\item \ordic{na te} \gloss{that’s nonsense}
\item \ordic{te na} \gloss{yeah, sure}
\item \ordic{qe te} \gloss{what the hell?}
\item \ordic{gra na} \gloss{damn it}
\item \ordic{shae na} \gloss{this sucks}
\item \ordic{dur na} \gloss{I can’t anymore}
\item \ordic{ret na} \gloss{forget it}
\item \ordic{lin na} \gloss{screw this}
\item \ordic{al na} \gloss{everything’s busted}
\item \ordic{na al} \gloss{absolutely not}

% --- Jokes / exaggeration
\item \ordic{al es} \gloss{everyone / everything}
\item \ordic{al nu} \gloss{all the time}
\item \ordic{al shae} \gloss{totally fine (sarcastic)}
\item \ordic{al qe} \gloss{no idea at all}
\item \ordic{al na} \gloss{nothing works}
\item \ordic{al ret} \gloss{it always comes back}
\item \ordic{al lin} \gloss{everyone’s leaving}
\item \ordic{al dur} \gloss{forever}
\item \ordic{al mu} \gloss{I remember everything}
\item \ordic{al\ldots} \gloss{everything\ldots}

% --- Intimate reassurance
\item \ordic{shae xo} \gloss{we’re okay}
\item \ordic{gra xo} \gloss{I care}
\item \ordic{te xo} \gloss{trust me}
\item \ordic{dur xo} \gloss{I’m staying}
\item \ordic{ret xo} \gloss{I’ll come back}
\item \ordic{nu xo} \gloss{right here with you}
\item \ordic{mu xo} \gloss{I remember us}
\item \ordic{es xo} \gloss{I’m with you}
\item \ordic{xo\ldots} \gloss{us\ldots}
\item \ordic{xo xo} \gloss{it’s us}

% --- Lullabies / soothing speech
\item \ordic{shae shae} \gloss{sleep, calm}
\item \ordic{dur shae} \gloss{stay calm}
\item \ordic{nu shae} \gloss{now it’s quiet}
\item \ordic{xo shae} \gloss{we’re safe}
\item \ordic{gra shae} \gloss{gentle now}
\item \ordic{mu shae} \gloss{soft memory}
\item \ordic{te shae} \gloss{it’s really okay}
\item \ordic{al shae} \gloss{everything’s fine}
\item \ordic{shae nu\ldots} \gloss{quiet now\ldots}
\item \ordic{shae\ldots} \gloss{hush\ldots}

% --- Bedtime promises
\item \ordic{ret nu} \gloss{I’ll be back}
\item \ordic{dur nu} \gloss{I’m here}
\item \ordic{xo nu} \gloss{I’m with you}
\item \ordic{te dur} \gloss{I won’t leave}
\item \ordic{gra dur} \gloss{I’ll stay gently}
\item \ordic{mu nu} \gloss{remember this}
\item \ordic{nu dur} \gloss{right here}
\item \ordic{nu ret} \gloss{back soon}
\item \ordic{nu xo} \gloss{right here, together}
\item \ordic{nu\ldots} \gloss{now\ldots}

% --- Rumors / hearsay
\item \ordic{mu qe} \gloss{I heard maybe}
\item \ordic{qe mu} \gloss{maybe remembered}
\item \ordic{mu al} \gloss{everyone says}
\item \ordic{mu nu al} \gloss{people say now}
\item \ordic{qe al} \gloss{who knows}
\item \ordic{qe te} \gloss{is that true?}
\item \ordic{te qe} \gloss{they say it’s true}
\item \ordic{mu ret qe} \gloss{I heard it came back}
\item \ordic{mu na te} \gloss{probably not true}
\item \ordic{mu\ldots} \gloss{I heard\ldots}

% --- Secrets / whispers
\item \ordic{na te mu} \gloss{don’t tell}
\item \ordic{mu na te} \gloss{forget I said it}
\item \ordic{xo na mu} \gloss{just us}
\item \ordic{xo mu na} \gloss{keep it between us}
\item \ordic{nu na te} \gloss{not now}
\item \ordic{qe na te} \gloss{don’t ask}
\item \ordic{te na mu} \gloss{you didn’t hear this}
\item \ordic{mu na} \gloss{never mind}
\item \ordic{na mu nu} \gloss{forget it now}
\item \ordic{shae na mu} \gloss{let it rest}

% --- Half-truths / comforting lies
\item \ordic{te shae} \gloss{it’s okay}
\item \ordic{te al shae} \gloss{everything’s fine}
\item \ordic{te dur} \gloss{it’ll last}
\item \ordic{te ret} \gloss{it’ll come back}
\item \ordic{te nu} \gloss{right now it’s fine}
\item \ordic{te xo} \gloss{we’re okay}
\item \ordic{te mu} \gloss{I remember right}
\item \ordic{te qe na} \gloss{don’t worry}
\item \ordic{te na qe} \gloss{no questions}
\item \ordic{te\ldots} \gloss{it’s fine\ldots}

% --- Childhood mishearings
\item \ordic{shanu} \gloss{sleep now}
\item \ordic{dunu} \gloss{stay here}
\item \ordic{xonu} \gloss{with me}
\item \ordic{muna} \gloss{story}
\item \ordic{qena} \gloss{maybe}
\item \ordic{granu} \gloss{please}
\item \ordic{nanu} \gloss{no more}
\item \ordic{lina} \gloss{go away}
\item \ordic{rete} \gloss{come back}
\item \ordic{nu\ldots} \gloss{now\ldots}

% --- Quiet fears
\item \ordic{qe shae} \gloss{I’m scared}
\item \ordic{qe dur} \gloss{will this last?}
\item \ordic{qe ret} \gloss{will you come back?}
\item \ordic{qe xo} \gloss{are we okay?}
\item \ordic{qe nu} \gloss{why now?}
\item \ordic{qe al} \gloss{what if everything…}
\item \ordic{qe na} \gloss{maybe not}
\item \ordic{qe\ldots} \gloss{um\ldots}
\item \ordic{shae qe} \gloss{it’s okay to be scared}
\item \ordic{gra qe} \gloss{it’s okay}

% --- Children rediscover structure
\item \ordic{es ord} \gloss{things make sense}
\item \ordic{ord es} \gloss{this is how it is}
\item \ordic{xo es} \gloss{we exist}
\item \ordic{es xo} \gloss{I’m with you}
\item \ordic{mu es} \gloss{I remember being}
\item \ordic{es mu} \gloss{I’m remembering}
\item \ordic{shae es} \gloss{being is calm}
\item \ordic{es shae} \gloss{I’m okay}
\item \ordic{qe es} \gloss{what am I?}
\item \ordic{es qe} \gloss{I wonder}

% --- Broken phrases that obey the Canon
\item \ordic{al es} \gloss{everything exists}
\item \ordic{es al} \gloss{I am everything (childish)}
\item \ordic{al xo} \gloss{everyone together}
\item \ordic{xo al} \gloss{we’re all}
\item \ordic{al mu} \gloss{all the stories}
\item \ordic{mu al} \gloss{I remember everything}
\item \ordic{al shae} \gloss{everything is okay}
\item \ordic{shae al} \gloss{calm everywhere}
\item \ordic{al qe} \gloss{so many questions}
\item \ordic{qe al} \gloss{what about everything?}

% --- Accidental High Ordic echoes
\item \ordic{es dur} \gloss{I keep being}
\item \ordic{dur es} \gloss{being keeps going}
\item \ordic{xo dur} \gloss{we don’t break}
\item \ordic{dur xo} \gloss{the bond stays}
\item \ordic{mu dur} \gloss{the memory stays}
\item \ordic{dur mu} \gloss{I keep remembering}
\item \ordic{ret es} \gloss{being comes back}
\item \ordic{es ret} \gloss{I come back}
\item \ordic{shae dur} \gloss{calm stays}
\item \ordic{dur shae} \gloss{stay calm}

% --- Playground metaphysics
\item \ordic{nu es} \gloss{here}
\item \ordic{es nu} \gloss{I’m here}
\item \ordic{nu xo} \gloss{with you now}
\item \ordic{xo nu} \gloss{together now}
\item \ordic{nu mu} \gloss{remember this}
\item \ordic{mu nu} \gloss{this memory}
\item \ordic{nu shae} \gloss{it’s okay now}
\item \ordic{shae nu} \gloss{now calm}
\item \ordic{nu qe} \gloss{what now?}
\item \ordic{qe nu} \gloss{why now?}

% --- The one who can speak safely
\item \ordic{vi es} \gloss{the one who sees being}
\item \ordic{vi ord} \gloss{the one who sees order}
\item \ordic{vi xo} \gloss{the one who sees bonds}
\item \ordic{vi mu} \gloss{the one who sees memory}
\item \ordic{vi shae} \gloss{the one who sees calm}
\item \ordic{vi qe} \gloss{the one who sees doubt}
\item \ordic{vi al} \gloss{the one who sees all}
\item \ordic{vi nu} \gloss{the one who is present}
\item \ordic{vi\ldots} \gloss{the witness\ldots}
\item \ordic{es vi} \gloss{being is seen}

% --- Near the Final Utterance (never complete)
\item \ordic{es al ord} \gloss{everything is ordered}
\item \ordic{al es ord} \gloss{all being aligns}
\item \ordic{al xo dur} \gloss{all bonds endure}
\item \ordic{al mu ret} \gloss{all memory returns}
\item \ordic{al shae es} \gloss{all being is calm}
\item \ordic{al qe shae} \gloss{all doubt rests}
\item \ordic{al gra xo} \gloss{all mercy binds}
\item \ordic{al\ldots} \gloss{all\ldots}
\item \ordic{\ldots} \gloss{pause}
\item \ordic{} \gloss{(silence)}

% --- Final child-speech fossils
\item \ordic{esh} \gloss{I am}
\item \ordic{xesh} \gloss{we are}
\item \ordic{mush} \gloss{story}
\item \ordic{shan} \gloss{calm}
\item \ordic{dun} \gloss{stay}
\item \ordic{ren} \gloss{return}
\item \ordic{qan} \gloss{why}
\item \ordic{grah} \gloss{please}
\item \ordic{allesh} \gloss{everything}
\item \ordic{es\ldots} \gloss{being\ldots}

\end{itemize}

\end{multicols}

\vspace{6pt}\hrule\vspace{4pt}
{\small Notes: Low Ordic allows erosion, reduplication, and sound-shift. Children often preserve ancient structures by accident.}

\end{document}
