\documentclass[10pt,a4paper]{article}

\usepackage[a4paper,margin=0.55in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{hyperref}

\hypersetup{hidelinks}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2pt}
\setlist[enumerate]{leftmargin=*, itemsep=1.5pt, topsep=2pt}

\newcommand{\ordic}[1]{\textbf{#1}}
\newcommand{\gloss}[1]{\hfill{\small\textsf{#1}}}

\begin{document}

{\LARGE \textbf{Middle Ordic: Common Phrases \small{(v0.1)}}}\\

\vspace{6pt}\hrule\vspace{8pt}

\begin{multicols}{2}

\begin{enumerate}

% --- Procedural existence
\item \ordic{ord es.} \gloss{Order exists.}
\item \ordic{jur es.} \gloss{Law exists.}
\item \ordic{set es.} \gloss{A structure exists.}
\item \ordic{mu es.} \gloss{A record exists.}
\item \ordic{log es.} \gloss{A reasoning process exists.}
\item \ordic{ord dur.} \gloss{Order remains in force.}
\item \ordic{jur dur.} \gloss{Law remains in force.}
\item \ordic{set dur.} \gloss{Structure remains in force.}
\item \ordic{mu dur.} \gloss{Record remains in force.}
\item \ordic{log dur.} \gloss{Reasoning remains valid.}

% --- Authority & mandate
\item \ordic{dar ord.} \gloss{Order must be followed.}
\item \ordic{dar jur.} \gloss{Law must be followed.}
\item \ordic{dar set.} \gloss{Structure must be applied.}
\item \ordic{dar log.} \gloss{Reasoning must be applied.}
\item \ordic{dar mu.} \gloss{Record must be kept.}
\item \ordic{fre ord.} \gloss{Order may be applied.}
\item \ordic{fre jur.} \gloss{Law may be applied.}
\item \ordic{fre set.} \gloss{Structure may be applied.}
\item \ordic{fre log.} \gloss{Reasoning may be applied.}
\item \ordic{fre mu.} \gloss{Record may be kept.}

% --- Compliance & obligation
\item \ordic{jur ord.} \gloss{Law orders.}
\item \ordic{jur set.} \gloss{Law structures.}
\item \ordic{jur mu.} \gloss{Law records.}
\item \ordic{jur log.} \gloss{Law reasons.}
\item \ordic{jur ßo.} \gloss{Law binds.}
\item \ordic{jur sol.} \gloss{Law binds by vow/contract.}
\item \ordic{dar sol.} \gloss{A contract must be honored.}
\item \ordic{sol dur.} \gloss{The contract remains valid.}
\item \ordic{sol mu.} \gloss{The contract is recorded.}
\item \ordic{sol jur.} \gloss{The contract has legal force.}

% --- Authorization & permission
\item \ordic{se lin.} \gloss{Proceed / entry allowed.}
\item \ordic{vu lin.} \gloss{Proceeding denied.}
\item \ordic{jur se lin.} \gloss{Law authorizes entry.}
\item \ordic{jur vu lin.} \gloss{Law denies entry.}
\item \ordic{fre lin.} \gloss{Passage is permitted.}
\item \ordic{dar lin.} \gloss{Passage is required.}
\item \ordic{lin ord.} \gloss{Proceed according to order.}
\item \ordic{lin jur.} \gloss{Proceed according to law.}
\item \ordic{lin set.} \gloss{Proceed according to procedure.}
\item \ordic{lin mu.} \gloss{Proceed according to record.}

% --- Records & documentation
\item \ordic{mu set.} \gloss{Record structure / file system.}
\item \ordic{mu set-ord.} \gloss{Ordered archive.}
\item \ordic{mu jur.} \gloss{Legal record.}
\item \ordic{mu sol.} \gloss{Contract record.}
\item \ordic{mu vi.} \gloss{Witness record.}
\item \ordic{mu vi-te.} \gloss{Verified record.}
\item \ordic{mu te.} \gloss{Certified true record.}
\item \ordic{mu dur.} \gloss{Record retained.}
\item \ordic{mu ša.} \gloss{Record deleted.}
\item \ordic{mu ret.} \gloss{Record restored.}

% --- Verification & review
\item \ordic{vi mu.} \gloss{Review record.}
\item \ordic{vi mu-te.} \gloss{Verify record.}
\item \ordic{vi sol.} \gloss{Review contract.}
\item \ordic{vi jur.} \gloss{Review law.}
\item \ordic{vi set.} \gloss{Review procedure.}
\item \ordic{vi log.} \gloss{Review reasoning.}
\item \ordic{vi ord.} \gloss{Review order.}
\item \ordic{te vi.} \gloss{Truth verified.}
\item \ordic{te vi dur.} \gloss{Verification remains valid.}
\item \ordic{qe vi.} \gloss{Verification questioned.}

% --- Error & exception handling
\item \ordic{qe mu.} \gloss{Record is questioned.}
\item \ordic{qe sol.} \gloss{Contract is questioned.}
\item \ordic{qe jur.} \gloss{Law is questioned.}
\item \ordic{qe set.} \gloss{Procedure is questioned.}
\item \ordic{qe log.} \gloss{Reasoning is questioned.}
\item \ordic{dif mu.} \gloss{Record discrepancy.}
\item \ordic{dif sol.} \gloss{Contract discrepancy.}
\item \ordic{dif jur.} \gloss{Legal discrepancy.}
\item \ordic{dif set.} \gloss{Procedure mismatch.}
\item \ordic{dif log.} \gloss{Reasoning mismatch.}

% --- Sanctions & enforcement
\item \ordic{dar ßo.} \gloss{Binding enforced.}
\item \ordic{jur ßo.} \gloss{Law binds.}
\item \ordic{ša ßo.} \gloss{Binding revoked.}
\item \ordic{ša sol.} \gloss{Contract terminated.}
\item \ordic{ša jur.} \gloss{Legal force removed.}
\item \ordic{ša set.} \gloss{Procedure invalidated.}
\item \ordic{ša mu.} \gloss{Record voided.}
\item \ordic{ša lin.} \gloss{Access revoked.}
\item \ordic{dar ša.} \gloss{Revocation required.}
\item \ordic{fre ša.} \gloss{Revocation permitted.}

% --- Completion & closure
\item \ordic{ret set.} \gloss{Procedure reset.}
\item \ordic{ret mu.} \gloss{Record reset.}
\item \ordic{ret sol.} \gloss{Contract renewed.}
\item \ordic{ret jur.} \gloss{Law reinstated.}
\item \ordic{ret ord.} \gloss{Order restored.}
\item \ordic{dur set.} \gloss{Procedure ongoing.}
\item \ordic{dur mu.} \gloss{Record ongoing.}
\item \ordic{dur sol.} \gloss{Contract ongoing.}
\item \ordic{dur jur.} \gloss{Law ongoing.}
\item \ordic{dur ord.} \gloss{Order ongoing.}

% --- Status & state reporting
\item \ordic{set dur.} \gloss{Procedure remains active.}
\item \ordic{set pun.} \gloss{Procedure executed (instant).}
\item \ordic{set ret.} \gloss{Procedure repeated.}
\item \ordic{set vek.} \gloss{Procedure changed.}
\item \ordic{ord vek.} \gloss{Order changed.}
\item \ordic{jur vek.} \gloss{Law amended.}
\item \ordic{sol vek.} \gloss{Contract modified.}
\item \ordic{mu vek.} \gloss{Record updated.}
\item \ordic{log vek.} \gloss{Reasoning updated.}
\item \ordic{lin vek.} \gloss{Process flow changed.}

% --- Scheduling & time handling
\item \ordic{pra set.} \gloss{Procedure completed (past).}
\item \ordic{pos set.} \gloss{Procedure scheduled.}
\item \ordic{dur set.} \gloss{Procedure ongoing.}
\item \ordic{pun set.} \gloss{Procedure executed immediately.}
\item \ordic{pra sol.} \gloss{Contract expired.}
\item \ordic{pos sol.} \gloss{Contract pending.}
\item \ordic{dur sol.} \gloss{Contract in force.}
\item \ordic{pun sol.} \gloss{Contract enacted.}
\item \ordic{pra jur.} \gloss{Law repealed.}
\item \ordic{pos jur.} \gloss{Law pending.}

% --- Delegation & responsibility
\item \ordic{ha jur.} \gloss{Authority granted.}
\item \ordic{ri jur.} \gloss{Authority received.}
\item \ordic{ha sol.} \gloss{Obligation assigned.}
\item \ordic{ri sol.} \gloss{Obligation accepted.}
\item \ordic{ha set.} \gloss{Procedure assigned.}
\item \ordic{ri set.} \gloss{Procedure accepted.}
\item \ordic{ha mu.} \gloss{Record transferred.}
\item \ordic{ri mu.} \gloss{Record received.}
\item \ordic{ha lin.} \gloss{Access granted.}
\item \ordic{ri lin.} \gloss{Access obtained.}

% --- Jurisdiction & scope
\item \ordic{in jur.} \gloss{Within jurisdiction.}
\item \ordic{ex jur.} \gloss{Outside jurisdiction.}
\item \ordic{in ord.} \gloss{Within mandate.}
\item \ordic{ex ord.} \gloss{Outside mandate.}
\item \ordic{in set.} \gloss{Within procedure.}
\item \ordic{ex set.} \gloss{Outside procedure.}
\item \ordic{in sol.} \gloss{Within contract.}
\item \ordic{ex sol.} \gloss{Outside contract.}
\item \ordic{in mu.} \gloss{Within record.}
\item \ordic{ex mu.} \gloss{Outside record.}

% --- Compliance checks
\item \ordic{vi set.} \gloss{Procedure reviewed.}
\item \ordic{vi sol.} \gloss{Contract reviewed.}
\item \ordic{vi jur.} \gloss{Law reviewed.}
\item \ordic{vi mu.} \gloss{Record reviewed.}
\item \ordic{vi ord.} \gloss{Order reviewed.}
\item \ordic{te vi set.} \gloss{Procedure verified.}
\item \ordic{te vi sol.} \gloss{Contract verified.}
\item \ordic{te vi jur.} \gloss{Law verified.}
\item \ordic{te vi mu.} \gloss{Record verified.}
\item \ordic{te vi ord.} \gloss{Order verified.}

% --- Exceptions & overrides
\item \ordic{qe dar.} \gloss{Mandate questioned.}
\item \ordic{qe fre.} \gloss{Permission questioned.}
\item \ordic{qe se.} \gloss{Authorization questioned.}
\item \ordic{qe vu.} \gloss{Denial questioned.}
\item \ordic{fre qe.} \gloss{Exception permitted.}
\item \ordic{dar qe.} \gloss{Exception required.}
\item \ordic{se qe.} \gloss{Exception authorized.}
\item \ordic{vu qe.} \gloss{Exception denied.}
\item \ordic{jur qe.} \gloss{Legal exception.}
\item \ordic{set qe.} \gloss{Procedural exception.}

% --- Enforcement outcomes
\item \ordic{ßo dur.} \gloss{Binding remains in force.}
\item \ordic{ßo pun.} \gloss{Binding enacted.}
\item \ordic{ßo ša.} \gloss{Binding released.}
\item \ordic{ßo ret.} \gloss{Binding reinstated.}
\item \ordic{ßo vek.} \gloss{Binding modified.}
\item \ordic{ša lin.} \gloss{Access revoked.}
\item \ordic{ša set.} \gloss{Procedure invalidated.}
\item \ordic{ša sol.} \gloss{Contract terminated.}
\item \ordic{ša jur.} \gloss{Legal authority removed.}
\item \ordic{ša mu.} \gloss{Record invalidated.}

% --- Reporting & notification
\item \ordic{so mu.} \gloss{Report recorded.}
\item \ordic{so jur.} \gloss{Legal notice issued.}
\item \ordic{so sol.} \gloss{Contract notice issued.}
\item \ordic{so set.} \gloss{Procedural notice issued.}
\item \ordic{so ord.} \gloss{Directive issued.}
\item \ordic{so vi.} \gloss{Observation reported.}
\item \ordic{so te.} \gloss{Statement certified.}
\item \ordic{so qe.} \gloss{Concern raised.}
\item \ordic{so dar.} \gloss{Mandate announced.}
\item \ordic{so fre.} \gloss{Permission announced.}

% --- Closure & archival
\item \ordic{mu lin.} \gloss{Record archived.}
\item \ordic{mu far.} \gloss{Record deprecated.}
\item \ordic{mu dur.} \gloss{Record retained.}
\item \ordic{mu pun.} \gloss{Record finalized.}
\item \ordic{mu ret.} \gloss{Record restored.}
\item \ordic{set lin.} \gloss{Procedure archived.}
\item \ordic{set far.} \gloss{Procedure deprecated.}
\item \ordic{set dur.} \gloss{Procedure retained.}
\item \ordic{set pun.} \gloss{Procedure finalized.}
\item \ordic{set ret.} \gloss{Procedure restored.}

% --- Institutional process flow
\item \ordic{lin set dur.} \gloss{Process flow remains active.}
\item \ordic{lin set pun.} \gloss{Process executed immediately.}
\item \ordic{lin set ret.} \gloss{Process restarted.}
\item \ordic{lin set ša.} \gloss{Process terminated.}
\item \ordic{lin ord dur.} \gloss{Mandated flow remains active.}
\item \ordic{lin ord pun.} \gloss{Mandated flow executed.}
\item \ordic{lin ord ret.} \gloss{Mandated flow restarted.}
\item \ordic{lin ord ša.} \gloss{Mandated flow revoked.}
\item \ordic{lin jur dur.} \gloss{Legal process remains active.}
\item \ordic{lin jur ša.} \gloss{Legal process terminated.}

% --- Compliance states
\item \ordic{eq set.} \gloss{Procedure compliant.}
\item \ordic{eq jur.} \gloss{Legally compliant.}
\item \ordic{eq sol.} \gloss{Contract compliant.}
\item \ordic{eq mu.} \gloss{Record compliant.}
\item \ordic{eq ord.} \gloss{Order compliant.}
\item \ordic{dif set.} \gloss{Procedure non-compliant.}
\item \ordic{dif jur.} \gloss{Legal non-compliance.}
\item \ordic{dif sol.} \gloss{Contract breach.}
\item \ordic{dif mu.} \gloss{Record discrepancy.}
\item \ordic{dif ord.} \gloss{Order deviation.}

% --- Escalation & review
\item \ordic{lin qe.} \gloss{Escalate for review.}
\item \ordic{qe lin.} \gloss{Escalation requested.}
\item \ordic{vi qe.} \gloss{Issue under review.}
\item \ordic{vi lin.} \gloss{Process under review.}
\item \ordic{vi set qe.} \gloss{Procedure under exception review.}
\item \ordic{vi jur qe.} \gloss{Legal exception review.}
\item \ordic{vi sol qe.} \gloss{Contract exception review.}
\item \ordic{vi mu qe.} \gloss{Record exception review.}
\item \ordic{vi ord qe.} \gloss{Mandate exception review.}
\item \ordic{so qe.} \gloss{Concern formally raised.}

% --- Sanction levels
\item \ordic{dar ša lin.} \gloss{Access revocation required.}
\item \ordic{fre ša lin.} \gloss{Access revocation permitted.}
\item \ordic{dar ša sol.} \gloss{Contract termination required.}
\item \ordic{fre ša sol.} \gloss{Contract termination permitted.}
\item \ordic{dar ša set.} \gloss{Procedure termination required.}
\item \ordic{fre ša set.} \gloss{Procedure termination permitted.}
\item \ordic{dar ša mu.} \gloss{Record invalidation required.}
\item \ordic{fre ša mu.} \gloss{Record invalidation permitted.}
\item \ordic{dar ša jur.} \gloss{Legal authority removal required.}
\item \ordic{fre ša jur.} \gloss{Legal authority removal permitted.}

% --- Remediation & correction
\item \ordic{ret set dur.} \gloss{Procedure correction ongoing.}
\item \ordic{ret set pun.} \gloss{Procedure corrected immediately.}
\item \ordic{ret sol dur.} \gloss{Contract remediation ongoing.}
\item \ordic{ret sol pun.} \gloss{Contract remediation executed.}
\item \ordic{ret mu dur.} \gloss{Record correction ongoing.}
\item \ordic{ret mu pun.} \gloss{Record corrected.}
\item \ordic{ret jur dur.} \gloss{Legal correction ongoing.}
\item \ordic{ret jur pun.} \gloss{Legal correction executed.}
\item \ordic{ret ord dur.} \gloss{Mandate correction ongoing.}
\item \ordic{ret ord pun.} \gloss{Mandate corrected.}

% --- Resource handling
\item \ordic{ha set.} \gloss{Resources allocated.}
\item \ordic{ri set.} \gloss{Resources received.}
\item \ordic{ha mu.} \gloss{Data transferred.}
\item \ordic{ri mu.} \gloss{Data received.}
\item \ordic{ha lin.} \gloss{Access granted.}
\item \ordic{ri lin.} \gloss{Access acquired.}
\item \ordic{ha ord.} \gloss{Mandate assigned.}
\item \ordic{ri ord.} \gloss{Mandate accepted.}
\item \ordic{ha sol.} \gloss{Obligation delegated.}
\item \ordic{ri sol.} \gloss{Obligation accepted.}

% --- Audit & traceability
\item \ordic{vi mu set.} \gloss{Archive audited.}
\item \ordic{vi mu jur.} \gloss{Legal archive audited.}
\item \ordic{vi mu sol.} \gloss{Contract archive audited.}
\item \ordic{vi mu ord.} \gloss{Mandate archive audited.}
\item \ordic{vi mu dur.} \gloss{Audit ongoing.}
\item \ordic{vi mu pun.} \gloss{Audit completed.}
\item \ordic{so mu vi.} \gloss{Audit report issued.}
\item \ordic{so mu qe.} \gloss{Audit concern logged.}
\item \ordic{mu vi-te.} \gloss{Archive verified.}
\item \ordic{mu vi-na.} \gloss{Archive unverifiable.}

% --- Termination & sunset
\item \ordic{ša set dur.} \gloss{Procedure termination in progress.}
\item \ordic{ša set pun.} \gloss{Procedure terminated.}
\item \ordic{ša sol dur.} \gloss{Contract termination in progress.}
\item \ordic{ša sol pun.} \gloss{Contract terminated.}
\item \ordic{ša mu dur.} \gloss{Record invalidation in progress.}
\item \ordic{ša mu pun.} \gloss{Record invalidated.}
\item \ordic{ša jur dur.} \gloss{Authority removal in progress.}
\item \ordic{ša jur pun.} \gloss{Authority removed.}
\item \ordic{ša ord dur.} \gloss{Mandate termination in progress.}
\item \ordic{ša ord pun.} \gloss{Mandate terminated.}

% --- Continuity assurance
\item \ordic{dur set ord.} \gloss{Procedural order maintained.}
\item \ordic{dur jur ord.} \gloss{Legal order maintained.}
\item \ordic{dur sol ord.} \gloss{Contractual order maintained.}
\item \ordic{dur mu ord.} \gloss{Record order maintained.}
\item \ordic{dur ord ord.} \gloss{Mandate order maintained.}
\item \ordic{dur lin ord.} \gloss{Operational flow maintained.}
\item \ordic{dur set set.} \gloss{Procedural integrity maintained.}
\item \ordic{dur jur set.} \gloss{Legal framework maintained.}
\item \ordic{dur mu set.} \gloss{Archive integrity maintained.}
\item \ordic{dur sol set.} \gloss{Contract framework maintained.}

% --- System states
\item \ordic{set eq.} \gloss{Procedure valid.}
\item \ordic{set na-eq.} \gloss{Procedure invalid.}
\item \ordic{ord eq.} \gloss{Mandate valid.}
\item \ordic{ord na-eq.} \gloss{Mandate invalid.}
\item \ordic{jur eq.} \gloss{Law valid.}
\item \ordic{jur na-eq.} \gloss{Law invalid.}
\item \ordic{mu eq.} \gloss{Record valid.}
\item \ordic{mu na-eq.} \gloss{Record invalid.}
\item \ordic{sol eq.} \gloss{Contract valid.}
\item \ordic{sol na-eq.} \gloss{Contract invalid.}

% --- Readiness / availability
\item \ordic{set lin.} \gloss{Procedure available.}
\item \ordic{set far.} \gloss{Procedure unavailable.}
\item \ordic{lin ord.} \gloss{Operational mandate active.}
\item \ordic{far ord.} \gloss{Operational mandate inactive.}
\item \ordic{lin jur.} \gloss{Legal pathway open.}
\item \ordic{far jur.} \gloss{Legal pathway closed.}
\item \ordic{lin mu.} \gloss{Record accessible.}
\item \ordic{far mu.} \gloss{Record inaccessible.}
\item \ordic{lin sol.} \gloss{Contract accessible.}
\item \ordic{far sol.} \gloss{Contract inaccessible.}

% --- Queueing / backlog
\item \ordic{set ne.} \gloss{Procedure pending.}
\item \ordic{set far ne.} \gloss{Procedure delayed.}
\item \ordic{mu ne.} \gloss{Record pending.}
\item \ordic{mu far ne.} \gloss{Record delayed.}
\item \ordic{sol ne.} \gloss{Contract pending.}
\item \ordic{jur ne.} \gloss{Legal review pending.}
\item \ordic{ord ne.} \gloss{Mandate pending.}
\item \ordic{lin ne.} \gloss{Access pending.}
\item \ordic{qe ne.} \gloss{Issue pending.}
\item \ordic{vi ne.} \gloss{Review pending.}

% --- Automation / repetition
\item \ordic{ret set auto.} \gloss{Procedure auto-repeated.}
\item \ordic{ret mu auto.} \gloss{Record auto-restored.}
\item \ordic{ret lin auto.} \gloss{Access auto-renewed.}
\item \ordic{ret sol auto.} \gloss{Contract auto-renewed.}
\item \ordic{ret jur auto.} \gloss{Law auto-reinstated.}
\item \ordic{dur set auto.} \gloss{Procedure runs automatically.}
\item \ordic{dur mu auto.} \gloss{Recording runs automatically.}
\item \ordic{dur lin auto.} \gloss{Access control automated.}
\item \ordic{dur sol auto.} \gloss{Contract enforcement automated.}
\item \ordic{dur jur auto.} \gloss{Legal enforcement automated.}

% --- Notifications / signals
\item \ordic{so lin.} \gloss{Access notice issued.}
\item \ordic{so far.} \gloss{Denial notice issued.}
\item \ordic{so dur.} \gloss{Status notice issued.}
\item \ordic{so pun.} \gloss{Completion notice issued.}
\item \ordic{so ret.} \gloss{Restart notice issued.}
\item \ordic{so ša.} \gloss{Termination notice issued.}
\item \ordic{so qe.} \gloss{Exception notice issued.}
\item \ordic{so vi.} \gloss{Review notice issued.}
\item \ordic{so eq.} \gloss{Compliance notice issued.}
\item \ordic{so dif.} \gloss{Non-compliance notice issued.}

% --- Capacity / limits
\item \ordic{set lim.} \gloss{Procedure at capacity.}
\item \ordic{mu lim.} \gloss{Storage at capacity.}
\item \ordic{lin lim.} \gloss{Access limit reached.}
\item \ordic{sol lim.} \gloss{Contract limit reached.}
\item \ordic{jur lim.} \gloss{Legal limit reached.}
\item \ordic{set na-lim.} \gloss{Procedure capacity available.}
\item \ordic{mu na-lim.} \gloss{Storage capacity available.}
\item \ordic{lin na-lim.} \gloss{Access capacity available.}
\item \ordic{sol na-lim.} \gloss{Contract capacity available.}
\item \ordic{jur na-lim.} \gloss{Legal capacity available.}

% --- Synchronization
\item \ordic{eq set set.} \gloss{Procedures synchronized.}
\item \ordic{eq mu mu.} \gloss{Records synchronized.}
\item \ordic{eq sol sol.} \gloss{Contracts synchronized.}
\item \ordic{eq jur jur.} \gloss{Laws synchronized.}
\item \ordic{eq ord ord.} \gloss{Mandates synchronized.}
\item \ordic{dif set set.} \gloss{Procedures desynchronized.}
\item \ordic{dif mu mu.} \gloss{Records desynchronized.}
\item \ordic{dif sol sol.} \gloss{Contracts desynchronized.}
\item \ordic{dif jur jur.} \gloss{Laws desynchronized.}
\item \ordic{dif ord ord.} \gloss{Mandates desynchronized.}

% --- Fallback / defaults
\item \ordic{ret def set.} \gloss{Procedure reverted to default.}
\item \ordic{ret def mu.} \gloss{Record reverted to default.}
\item \ordic{ret def sol.} \gloss{Contract reverted to default.}
\item \ordic{ret def jur.} \gloss{Law reverted to default.}
\item \ordic{ret def ord.} \gloss{Mandate reverted to default.}
\item \ordic{fre def.} \gloss{Default permitted.}
\item \ordic{dar def.} \gloss{Default required.}
\item \ordic{se def.} \gloss{Default authorized.}
\item \ordic{vu def.} \gloss{Default denied.}
\item \ordic{so def.} \gloss{Default declared.}

% --- End-of-cycle markers
\item \ordic{pun set end.} \gloss{Procedure cycle complete.}
\item \ordic{pun mu end.} \gloss{Record cycle complete.}
\item \ordic{pun sol end.} \gloss{Contract cycle complete.}
\item \ordic{pun jur end.} \gloss{Legal cycle complete.}
\item \ordic{pun ord end.} \gloss{Mandate cycle complete.}
\item \ordic{dur next.} \gloss{Continue to next cycle.}
\item \ordic{ret next.} \gloss{Restart next cycle.}
\item \ordic{ša next.} \gloss{Terminate next cycle.}
\item \ordic{qe next.} \gloss{Next cycle questioned.}
\item \ordic{vi next.} \gloss{Next cycle reviewed.}
 
% --- Final validations
\item \ordic{vi eq.} \gloss{Compliance confirmed.}
\item \ordic{vi na-eq.} \gloss{Compliance failed.}
\item \ordic{te eq.} \gloss{Truth validated.}
\item \ordic{te na-eq.} \gloss{Truth rejected.}
\item \ordic{mu eq.} \gloss{Record validated.}
\item \ordic{mu na-eq.} \gloss{Record rejected.}
\item \ordic{sol eq.} \gloss{Contract validated.}
\item \ordic{sol na-eq.} \gloss{Contract rejected.}
\item \ordic{jur eq.} \gloss{Law validated.}
\item \ordic{jur na-eq.} \gloss{Law rejected.}

% --- Final approvals
\item \ordic{se fin.} \gloss{Final approval granted.}
\item \ordic{vu fin.} \gloss{Final approval denied.}
\item \ordic{jur se fin.} \gloss{Final legal approval granted.}
\item \ordic{jur vu fin.} \gloss{Final legal approval denied.}
\item \ordic{so fin.} \gloss{Final decision announced.}
\item \ordic{so mu fin.} \gloss{Final record issued.}
\item \ordic{so sol fin.} \gloss{Final contract issued.}
\item \ordic{so set fin.} \gloss{Final procedure issued.}
\item \ordic{so ord fin.} \gloss{Final mandate issued.}
\item \ordic{so jur fin.} \gloss{Final ruling issued.}

% --- Terminal enforcement
\item \ordic{dar fin.} \gloss{Final enforcement required.}
\item \ordic{fre fin.} \gloss{Final enforcement permitted.}
\item \ordic{ša fin.} \gloss{Final termination executed.}
\item \ordic{ša sol fin.} \gloss{Contract terminated finally.}
\item \ordic{ša set fin.} \gloss{Procedure terminated finally.}
\item \ordic{ša mu fin.} \gloss{Record invalidated finally.}
\item \ordic{ša ord fin.} \gloss{Mandate terminated finally.}
\item \ordic{ša jur fin.} \gloss{Authority revoked finally.}
\item \ordic{ša lin fin.} \gloss{Access revoked finally.}
\item \ordic{ša all fin.} \gloss{Total system termination.}

% --- Archival closure
\item \ordic{mu lin fin.} \gloss{Record archived permanently.}
\item \ordic{mu far fin.} \gloss{Record sealed and withdrawn.}
\item \ordic{mu dur fin.} \gloss{Record retained permanently.}
\item \ordic{set lin fin.} \gloss{Procedure archived permanently.}
\item \ordic{set far fin.} \gloss{Procedure sealed permanently.}
\item \ordic{ord lin fin.} \gloss{Mandate archived permanently.}
\item \ordic{ord far fin.} \gloss{Mandate sealed permanently.}
\item \ordic{jur lin fin.} \gloss{Law archived permanently.}
\item \ordic{jur far fin.} \gloss{Law sealed permanently.}
\item \ordic{lin fin.} \gloss{System closed.}

% --- Silence markers
\item \ordic{dur fin.} \gloss{Final state persists.}
\item \ordic{ret fin-na.} \gloss{No restart permitted.}
\item \ordic{qe fin-na.} \gloss{No appeal permitted.}
\item \ordic{vi fin-na.} \gloss{No review permitted.}
\item \ordic{so fin-na.} \gloss{No statement permitted.}
\item \ordic{lin fin-na.} \gloss{No access permitted.}
\item \ordic{fre fin-na.} \gloss{No exception permitted.}
\item \ordic{dar fin-na.} \gloss{No mandate remains.}
\item \ordic{jur fin-na.} \gloss{No jurisdiction remains.}
\item \ordic{ord fin-na.} \gloss{No order remains.}

% --- Bureaucratic epilogues
\item \ordic{es fin.} \gloss{The state is final.}
\item \ordic{set fin.} \gloss{The procedure is final.}
\item \ordic{mu fin.} \gloss{The record is final.}
\item \ordic{sol fin.} \gloss{The contract is final.}
\item \ordic{jur fin.} \gloss{The law is final.}
\item \ordic{ord fin.} \gloss{The mandate is final.}
\item \ordic{ßo fin.} \gloss{The binding is final.}
\item \ordic{te fin.} \gloss{Truth is declared final.}
\item \ordic{eq fin.} \gloss{Equilibrium declared final.}
\item \ordic{fin.} \gloss{Final.}

\end{enumerate}

\end{multicols}

\vspace{6pt}\hrule\vspace{4pt}
{\small Notes: Middle Ordic favors explicit authority, procedure, and durability. Emotional particles are rare and usually absent.}

\end{document}
