
\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=0.6in]{geometry}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{microtype}
\setlength{\parindent}{0pt}
\begin{document}
{\LARGE \textbf{Veyruun --- Core 1000 Phrases \small{(v0.1)}}}\\
\vspace{6pt}
{\setlength{\tabcolsep}{3pt}
\begin{longtable}{@{}r p{0.46\linewidth} p{0.40\linewidth} p{0.10\linewidth}@{}}
\toprule
\textbf{\#} & \textbf{Veyruun} & \textbf{English} & \textbf{Tag}\\
\midrule
\endfirsthead
\toprule
\textbf{\#} & \textbf{Veyruun} & \textbf{English} & \textbf{Tag}\\
\midrule
\endhead
\bottomrule
\endfoot
1 & Relneth eh? & You been through the folds? & greeting\\
2 & Relneth. & I have. & reply\\
3 & Zan'kai eh? & Are we bonded? & bond\\
4 & Ren na. & Breathe. Easy. & calm\\
5 & Tovaš. Vul nai. & Enough. Don't speak. & conflict\\
6 & Bol sivm! & Seal the hatch! & ship ops\\
7 & Tor lirm. & Give me the light. & ship ops\\
8 & Zeth nai! & No air! & emergency\\
9 & Rask nai. & No fuel. & ship ops\\
10 & Korša. & I'm fixing it / We're fixing it. & ship ops\\
11 & Zul ker nai. Kai ker ve. & The void has no meaning. The chosen do. & theme\\
12 & Kérun ve. & Meaning is made. & theme\\
13 & Zul nai. Reln ve. Zan'kai ven. & Not the void. The seam is. We cross together. & chant\\
14 & se zethun lalgemra. & they leave air; oxygen supply. & statement\\
15 & se hílenkordun dolim . & they turn hílen-kil (plain). & log\\
16 & ti milša eh? & Does you walk? & question\\
17 & se nidethun draiša tov. & they to drift; wander food (inferred). & log\\
18 & se zešra nai. & they do not help. & negation\\
19 & tir šinša eh? & Does you (pl.) to drink; drink/water? & question\\
20 & kairelnarun donm! & break kairelnar! & command\\
21 & zan zanka škonra. & we hide with (neutral). & social\\
22 & za vethenun našteta eh? & Does I trade fear? & question\\
23 & za sizruvra nai. & I do not forget. & negation\\
24 & se zlerta, kur tir kaiumun dar. & they love, because you (pl.) to want; desire the chosen. & cause\\
25 & zan'kai zaka vomša. & we (bonded) wait with (neutral). & social\\
26 & se kethta, kur se vaskdarleun tal. & they to hold; keep, because they to rest; pause vask-darle. & cause\\
27 & zan'kai heßa nai. & we (bonded) do not to listen. & negation\\
28 & darm! & to want; desire! & command\\
29 & tir natta nai. & you (pl.) do not doubt. & negation\\
30 & ti švoniun lalgemra reln. & you leave choice (fold-dream). & log\\
31 & se mórzlenasun sok. & they to take mór-zlenas. & statement\\
32 & za mórtalun šoš. & I heal mórtal. & statement\\
33 & tir krumta, kur za zulun milra. & you (pl.) think, because I walk indifferent void; cosmic uncaring. & cause\\
34 & zan'kai atsnukra eh? & Does we (bonded) come? & question\\
35 & se glekun natra eh? & Does they doubt joy? & question\\
36 & zan hílenhešun šurra reln. & we to survive; endure hílen-heš (fold-dream). & log\\
37 & zan sálišvonun lomakra. & we find sáli-švon. & statement\\
38 & lonm! & know! & command\\
39 & se škunta. & they watch. & statement\\
40 & se seka gliekša. & they fall with (neutral). & social\\
41 & tir zanka dolimša. & you (pl.) turn with (neutral). & social\\
42 & se snonta eh? & Does they arrive? & question\\
43 & zlenasm! & build! & command\\
44 & tir kathun venša eh? & Does you (pl.) to cross; to jump a threshold oath; promise that costs something? & question\\
45 & ti zulkaíventa. & you crossing the void with the chosen. & statement\\
46 & se rukra eh? & Does they hurt? & question\\
47 & zan venša, kur tir sivarun zrad. & we to cross; to jump a threshold, because you (pl.) remember lock; seal. & cause\\
48 & zan'kai molra nai. & we (bonded) do not carry. & negation\\
49 & ti som, kur ti zethun ronzid. & you to sleep, because you believe air; oxygen supply. & cause\\
50 & ti šurša eh? & Does you to survive; endure? & question\\
51 & zan'kai vilkalethun škirra, šaen zan'kai gimkredun lirra. & we (bonded) run forgive, but we (bonded) to give doctor. & contrast\\
52 & tir stušun ruk eh? & Does you (pl.) hurt truth? & question\\
53 & ti zaka vulša. & you to speak; say with (neutral). & social\\
54 & za škunta, kur za kaizulun dunša. & I watch, because I to eat kai-zul. & cause\\
55 & za rogresarun kraš eh? & Does I rise calm? & question\\
56 & zan'kai draita nai. & we (bonded) do not to drift; wander. & negation\\
57 & se bolun ronzidša, šaen se sáliun dunša. & they believe door; hatch, but they to eat stubborn defiant hope-light. & contrast\\
58 & vulm! & to speak; say! & command\\
59 & rogresun zaknimm! & close calm! & command\\
60 & zan'kai vakrekra eh? & Does we (bonded) promise? & question\\
61 & za za'kai draita. & I to drift; wander with the chosen. & bond\\
62 & zan'kai škonra nai. & we (bonded) do not hide. & negation\\
63 & zan zaka darta. & we to want; desire with (neutral). & social\\
64 & kavgitun sairm! & to tell truth; align speech with reality air! & command\\
65 & zan'kai zlenasra, kur za sizruvumun vul. & we (bonded) build, because I to speak; say forget. & cause\\
66 & zan relnumun tovaršurra eh? & Does we endure by continuing the story-thread fold seam? & question\\
67 & zan raskun hešra. & we to listen fuel. & statement\\
68 & zan draiša, kur zan'kai drunkaiumun zimša. & we to drift; wander, because we (bonded) pay drunkaium. & cause\\
69 & bolun vakrekm! & promise door; hatch! & command\\
70 & se sokša, kur tir kaikrašun vak. & they to take, because you (pl.) to wake kai-kraš. & cause\\
71 & se bolun zim. & they pay door; hatch. & statement\\
72 & zan sair nai. & we do not to tell truth; align speech with reality. & negation\\
73 & tir za'kai šurra. & you (pl.) to survive; endure with the chosen. & bond\\
74 & relnarun kridm! & go fold-gate/place! & command\\
75 & ti somta, kur zan'kai mórkrenun darra. & you to sleep, because we (bonded) to want; desire mór-krek. & cause\\
76 & se zan'kai dun. & they to eat with the chosen. & bond\\
77 & za vaskumun mol, šaen za kathkaiun vilkalra. & I carry ship; vessel, but I forgive oath-kin. & contrast\\
78 & za dunta. & I to eat. & statement\\
79 & tir kathvakrekun molta eh? & Does you (pl.) carry kath-vakrek? & question\\
80 & zan'kai zulrukun sizruvša tov. & we (bonded) forget zul-ruk (inferred). & log\\
81 & sairm! & to tell truth; align speech with reality! & command\\
82 & za šurra. & I to survive; endure. & statement\\
83 & zan lon eh? & Does we know? & question\\
84 & za vrenra, kur za klutenun zradša. & I to lie; a lie, because I remember strong. & cause\\
85 & zan sairta eh? & Does we to tell truth; align speech with reality? & question\\
86 & zan sálidolimun dar. & we to want; desire sáli-dolim. & statement\\
87 & zan hílensuvun snon eh? & Does we arrive hílen-suv? & question\\
88 & zan mekun šurša. & we to survive; endure food. & statement\\
89 & kathun trem! & fly oath; promise that costs something! & command\\
90 & tir torun škunta eh? & Does you (pl.) watch light? & question\\
91 & tir ronzidra eh? & Does you (pl.) believe? & question\\
92 & za zan'kai zlus. & I buy with the chosen. & bond\\
93 & bolethun korm! & to work; fix door; hatch! & command\\
94 & za dolim nai. & I do not turn. & negation\\
95 & tir venta, kur za kavgitun zlenasša. & you (pl.) to cross; to jump a threshold, because I build air. & cause\\
96 & zan'kai ašun venša eh? & Does we (bonded) to cross; to jump a threshold heat? & question\\
97 & za vaskrorun zlusša. & I buy vask-ror. & statement\\
98 & se zeßa nai. & they do not help. & negation\\
99 & za zanka dar. & I to want; desire with (neutral). & social\\
100 & ti venša nai. & you do not to cross; to jump a threshold. & negation\\
101 & se sivun vul. & they to speak; say lock; seal. & statement\\
102 & se se'kai lalgem. & they leave with the chosen. & bond\\
103 & za šurša nai. & I do not to survive; endure. & negation\\
104 & se sáliun dolimra . & they turn stubborn defiant hope-light (plain). & log\\
105 & kathstilun krumm! & think kath-stil! & command\\
106 & se dunta eh? & Does they to eat? & question\\
107 & tir zanka kridša. & you (pl.) go with (neutral). & social\\
108 & za vaskkrekun somra, šaen za drunbolun vrenra. & I to sleep vask-krek, but I to lie; a lie drun-bol. & contrast\\
109 & se zulklutun hešra. & they to listen zul-klut. & statement\\
110 & zan'kai korta, kur se vekun draita. & we (bonded) to work; fix, because they to drift; wander dark. & cause\\
111 & za vevarun lirta eh? & Does I to give problem? & question\\
112 & tir zanka zradra. & you (pl.) remember with (neutral). & social\\
113 & za rukra nai. & I do not hurt. & negation\\
114 & se kethra nai. & they do not to hold; keep. & negation\\
115 & se gliek nai. & they do not fall. & negation\\
116 & ti vyrun somta tov. & you to sleep cold (inferred). & log\\
117 & šurm! & to survive; endure! & command\\
118 & ti ti'kai treša. & you fly with the chosen. & bond\\
119 & se zethun molša vem. & they carry air; oxygen supply (reported). & log\\
120 & ti hílenun šinra, šaen ti ašun škirša. & you to drink; drink/water homesickmemory, but you run heat. & contrast\\
121 & se kavgitun dar reln. & they to want; desire air (fold-dream). & log\\
122 & zan darta, kur tir švularun treta. & we to want; desire, because you (pl.) fly empty. & cause\\
123 & se gorir, kur ti sivun dun. & they learn, because you to eat lock; seal. & cause\\
124 & zan zulvrezomun vakta, šaen zan krašarun lalgem. & we to wake zul-vrezom, but we leave rise. & contrast\\
125 & nérenun trem! & fly star! & command\\
126 & ti kathzešun rlasša eh? & Does you lose kath-zeš? & question\\
127 & sálisnonun škunm! & watch sáli-snon! & command\\
128 & za kordun lon vem. & I know tool (reported). & log\\
129 & se seka krid. & they go with (neutral). & social\\
130 & škirm! & run! & command\\
131 & tir tovarvrunun rlasra. & you (pl.) lose tovar-vrun. & statement\\
132 & za zulkaívenša. & I crossing the void with the chosen. & statement\\
133 & zan'kai za'kai šur. & we (bonded) to survive; endure with the chosen. & bond\\
134 & kaiumiun sairm! & to tell truth; align speech with reality the chosen! & command\\
135 & tir kathkordun našte tov. & you (pl.) trade kath-kil (inferred). & log\\
136 & zan drisun šurta . & we to survive; endure home (plain). & log\\
137 & lesm! & hate! & command\\
138 & za seka dunra. & I to eat with (neutral). & social\\
139 & ti treenun gaktadša . & you fail fly (plain). & log\\
140 & tir drunvnunvutun sizruvša . & you (pl.) forget drun-vnunvut (plain). & log\\
141 & se švulenun šurta ser. & they to survive; endure empty (seen/verified). & log\\
142 & ti donra eh? & Does you break? & question\\
143 & ti krumša nai. & you do not think. & negation\\
144 & zan zrismidun sokra, šaen zan mórtrerun kethša. & we to take map, but we to hold; keep mór-trer. & contrast\\
145 & se relnarun venša tov. & they to cross; to jump a threshold fold-gate/place (inferred). & log\\
146 & za relnlokun tatša eh? & Does I choose reln-lok? & question\\
147 & vaskrorun trem! & fly vask-ror! & command\\
148 & tir nérzešun somša. & you (pl.) to sleep nér-zeš. & statement\\
149 & somm! & to sleep! & command\\
150 & hešm! & to listen! & command\\
151 & tir zan'kai hešta. & you (pl.) to listen with the chosen. & bond\\
152 & zan nidun vretra, šaen zan hílensuvun keth. & we push food, but we to hold; keep hílen-suv. & contrast\\
153 & se mil. & they walk. & statement\\
154 & ti tovarsáliun šinša tov. & you to drink; drink/water tovar-sáli (inferred). & log\\
155 & rerlerenun sokm! & to take water! & command\\
156 & nekun somm! & to sleep pilot! & command\\
157 & za darra. & I to want; desire. & statement\\
158 & tir se'kai dolimta. & you (pl.) turn with the chosen. & bond\\
159 & za vakrekša, kur tir tešarun šin. & I promise, because you (pl.) to drink; drink/water afraid. & cause\\
160 & se vruniun ronzidra, šaen se sálihílenun zlenasra. & they believe fear, but they build sáli-hílen. & contrast\\
161 & se grimarun lesra, šaen se kešarun drai. & they hate room, but they to drift; wander wire; line; cable. & contrast\\
162 & tir heßa eh? & Does you (pl.) to listen? & question\\
163 & tir lalgem, kur ti draiethun kor. & you (pl.) leave, because you to work; fix to drift; wander. & cause\\
164 & tir zanka sokša. & you (pl.) to take with (neutral). & social\\
165 & zan kathun venra. & we to cross; to jump a threshold oath; promise that costs something. & statement\\
166 & zlusm! & buy! & command\\
167 & venm! & to cross; to jump a threshold! & command\\
168 & zan'kai zanka vakra. & we (bonded) to wake with (neutral). & social\\
169 & zan se'kai zulkaíventa. & we crossing the void with the chosen with the chosen. & bond\\
170 & ti zaka milra. & you walk with (neutral). & social\\
171 & se zan'kai zlus. & they buy with the chosen. & bond\\
172 & za tika krumra. & I think with (neutral). & social\\
173 & tir mórvretun šinta, šaen tir mórzošonun tikta. & you (pl.) to drink; drink/water mór-vret, but you (pl.) teach mór-zošon. & contrast\\
174 & ti tika vren. & you to lie; a lie with (neutral). & social\\
175 & ti sokarun heš. & you to listen to take. & statement\\
176 & se za'kai krid. & they go with the chosen. & bond\\
177 & za kaiiun dolimša eh? & Does I turn chosen companion; bonded company? & question\\
178 & ti seka zler. & you love with (neutral). & social\\
179 & ti za'kai kethta. & you to hold; keep with the chosen. & bond\\
180 & za vulra eh? & Does I to speak; say? & question\\
181 & sokm! & to take! & command\\
182 & ti šurta, kur tir sálitrelun lonša. & you to survive; endure, because you (pl.) know sáli-trel. & cause\\
183 & se tika zlerra. & they love with (neutral). & social\\
184 & se kešun atsnukša, šaen se drisenun ven. & they come wire; line; cable, but they to cross; to jump a threshold home. & contrast\\
185 & zan'kai vretta, kur tir kaisáliun don. & we (bonded) push, because you (pl.) break kai-sáli. & cause\\
186 & se raskun darša . & they to want; desire fuel (plain). & log\\
187 & ti vyrun vakta reln. & you to wake cold (fold-dream). & log\\
188 & se vnunvutun vulša, šaen se nérkrenun škunša. & they to speak; say captain, but they watch nér-krek. & contrast\\
189 & zan'kai zler nai. & we (bonded) do not love. & negation\\
190 & zan'kai nérledzrasun lirša, šaen zan'kai zulrelnarun mis. & we (bonded) to give nér-ledzras, but we (bonded) open zulrelnar. & contrast\\
191 & se sokša, kur zan'kai zulenun venta. & they to take, because we (bonded) to cross; to jump a threshold indifferent void; cosmic uncaring. & cause\\
192 & kridenun venm! & to cross; to jump a threshold go! & command\\
193 & zan'kai vilkalenun lirša eh? & Does we (bonded) to give forgive? & question\\
194 & zan draira nai. & we do not to drift; wander. & negation\\
195 & se škirta. & they run. & statement\\
196 & se vretta, kur zan hílenun darra. & they push, because we to want; desire homesickmemory. & cause\\
197 & zan zan'kai somta. & we to sleep with the chosen. & bond\\
198 & tir torun šoš. & you (pl.) heal light. & statement\\
199 & tir golarun vradgota eh? & Does you (pl.) start door? & question\\
200 & zan kathvakrekun zradra eh? & Does we remember kath-vakrek? & question\\
201 & zulun donm! & break indifferent void; cosmic uncaring! & command\\
202 & snonm! & arrive! & command\\
203 & se lomakša eh? & Does they find? & question\\
204 & zan'kai zulstokzorun šoßa. & we (bonded) heal zul-stokzor. & statement\\
205 & zan snonša. & we arrive. & statement\\
206 & za somša eh? & Does I to sleep? & question\\
207 & za zeßa. & I help. & statement\\
208 & zan'kai se'kai sairša. & we (bonded) to tell truth; align speech with reality with the chosen. & bond\\
209 & mórškikun donm! & break mór-škik! & command\\
210 & suvun gliekm! & fall contract! & command\\
211 & se mitumun atsnukša. & they come touch. & statement\\
212 & zan zan'kai vren. & we to lie; a lie with the chosen. & bond\\
213 & zan'kai zethun kor eh? & Does we (bonded) to work; fix air; oxygen supply? & question\\
214 & ašun zlenasm! & build heat! & command\\
215 & atsnukiun zaknimm! & close come! & command\\
216 & ti tešarun kethta, šaen ti hílenskultiun vakra. & you to hold; keep afraid, but you to wake hílen-skulti. & contrast\\
217 & se seka mitra. & they touch with (neutral). & social\\
218 & ti vakša nai. & you do not to wake. & negation\\
219 & ti senethun darleta, šaen ti zulzradun zlenas. & you live skin; surface; hull, but you build zul-zrad. & contrast\\
220 & za lirra nai. & I do not to give. & negation\\
221 & tir gliek. & you (pl.) fall. & statement\\
222 & zan'kai škirra nai. & we (bonded) do not run. & negation\\
223 & vradgom! & start! & command\\
224 & ti zan'kai snon. & you arrive with the chosen. & bond\\
225 & dolimm! & turn! & command\\
226 & ti zan'kai darta. & you to want; desire with the chosen. & bond\\
227 & tir kor. & you (pl.) to work; fix. & statement\\
228 & se seka korra. & they to work; fix with (neutral). & social\\
229 & zan'kai se'kai krumra. & we (bonded) think with the chosen. & bond\\
230 & zan vnunvutarun zradša, šaen zan golumun dolimša. & we remember captain, but we turn door. & contrast\\
231 & za veniun zleramša . & I escape to cross; to jump a threshold (plain). & log\\
232 & tir drunrelnethun zleramra eh? & Does you (pl.) escape drunrelneth? & question\\
233 & ti zlerethun rlas. & you lose love. & statement\\
234 & se šurta, kur zan sálišvonun šur. & they to survive; endure, because we to survive; endure sáli-švon. & cause\\
235 & se zulun hešra eh? & Does they to listen indifferent void; cosmic uncaring? & question\\
236 & ti se'kai tikra. & you teach with the chosen. & bond\\
237 & tir vyrun vom ser. & you (pl.) wait cold (seen/verified). & log\\
238 & zan relnarumun vakta eh? & Does we to wake fold-gate/place? & question\\
239 & kathvakun sokm! & to take kath-vak! & command\\
240 & hílenzlusun vulm! & to speak; say hílen-zlus! & command\\
241 & se ti'kai kethra. & they to hold; keep with the chosen. & bond\\
242 & ti zaka korša. & you to work; fix with (neutral). & social\\
243 & ti se'kai zlenasra. & you build with the chosen. & bond\\
244 & híleniun kethm! & to hold; keep homesickmemory! & command\\
245 & za kethta. & I to hold; keep. & statement\\
246 & ti kilun som eh? & Does you to sleep tool? & question\\
247 & za draiethun draira, šaen za torun zimša. & I to drift; wander to drift; wander, but I pay light. & contrast\\
248 & zan'kai zanka snašta. & we (bonded) receive with (neutral). & social\\
249 & za kešarun gliekta ser. & I fall wire; line; cable (seen/verified). & log\\
250 & tir tika gliek. & you (pl.) fall with (neutral). & social\\
251 & nérzrismidun sairm! & to tell truth; align speech with reality nér-zrismid! & command\\
252 & ti susdisra. & you feel. & statement\\
253 & za se'kai lir. & I to give with the chosen. & bond\\
254 & zan'kai ti'kai dun. & we (bonded) to eat with the chosen. & bond\\
255 & lirm! & to give! & command\\
256 & zaknimm! & close! & command\\
257 & rimenun vradgom! & start angry! & command\\
258 & ti glaethun kethta. & you to hold; keep plan. & statement\\
259 & zan ti'kai vakrek. & we promise with the chosen. & bond\\
260 & drunethun darm! & to want; desire dread; the weight of meaninglessness! & command\\
261 & tir sairta. & you (pl.) to tell truth; align speech with reality. & statement\\
262 & ti vakrekša eh? & Does you promise? & question\\
263 & gaktadm! & fail! & command\\
264 & zan vasktreun dolimša vem. & we turn vask-tre (reported). & log\\
265 & druniarun vomm! & wait dread! & command\\
266 & za hešra. & I to listen. & statement\\
267 & ti lumlavarun snon. & you arrive refuse. & statement\\
268 & nérvnunvutun zlusm! & buy nér-vnunvut! & command\\
269 & zan ti'kai zulkaívenša. & we crossing the void with the chosen with the chosen. & bond\\
270 & zan ti'kai sokta. & we to take with the chosen. & bond\\
271 & zan'kai zaka zulkaíventa. & we (bonded) crossing the void with the chosen with (neutral). & social\\
272 & tir tika vakta. & you (pl.) to wake with (neutral). & social\\
273 & ti vulša nai. & you do not to speak; say. & negation\\
274 & vakrekm! & promise! & command\\
275 & se švulenun sairša, šaen se bolenun tre. & they to tell truth; align speech with reality empty, but they fly door; hatch. & contrast\\
276 & ti mitumun vradgora. & you start touch. & statement\\
277 & zan tika vakra. & we to wake with (neutral). & social\\
278 & za donra, kur ti tovarnenun sairra. & I break, because you to tell truth; align speech with reality tovar-nen. & cause\\
279 & zan'kai duniun heš eh? & Does we (bonded) to listen to eat? & question\\
280 & ti zlerenun zešta, šaen ti golumun zulkaíventa. & you help love, but you crossing the void with the chosen door. & contrast\\
281 & tir donta eh? & Does you (pl.) break? & question\\
282 & zan zaknimenun natša, šaen zan zrismidenun gorirta. & we doubt close, but we learn map. & contrast\\
283 & zan'kai tovarkrekun venta ser. & we (bonded) to cross; to jump a threshold tovar-krek (seen/verified). & log\\
284 & tir seka lonta. & you (pl.) know with (neutral). & social\\
285 & zan vilkal eh? & Does we forgive? & question\\
286 & tir hílenskultiun lalgemra eh? & Does you (pl.) leave hílen-skulti? & question\\
287 & za tovarrukun dolim eh? & Does I turn tovar-ruk? & question\\
288 & se vakra. & they to wake. & statement\\
289 & za raskun draiša reln. & I to drift; wander fuel (fold-dream). & log\\
290 & zan'kai kathlesun vradgo eh? & Does we (bonded) start kath-les? & question\\
291 & ti zulkaívenša. & you crossing the void with the chosen. & statement\\
292 & se kethša. & they to hold; keep. & statement\\
293 & zan'kai drai nai. & we (bonded) do not to drift; wander. & negation\\
294 & ti zlenas, kur za zethun somša. & you build, because I to sleep air; oxygen supply. & cause\\
295 & tir mil eh? & Does you (pl.) walk? & question\\
296 & ti relnarun venta reln. & you to cross; to jump a threshold fold-gate/place (fold-dream). & log\\
297 & zan'kai se'kai heš. & we (bonded) to listen with the chosen. & bond\\
298 & zan seka šošra. & we heal with (neutral). & social\\
299 & tir mórrlotun kraßa, šaen tir zradethun vradgoša. & you (pl.) rise mór-rlot, but you (pl.) start remember. & contrast\\
300 & ti krekethun lir, šaen ti sniumun darta. & you to give anger, but you to want; desire wrong. & contrast\\
301 & šinm! & to drink; drink/water! & command\\
302 & zan'kai kavgitenun dolimta eh? & Does we (bonded) turn air? & question\\
303 & tir milta nai. & you (pl.) do not walk. & negation\\
304 & ti seka zleram. & you escape with (neutral). & social\\
305 & ti donta. & you break. & statement\\
306 & vekun zešm! & help dark! & command\\
307 & tir krid eh? & Does you (pl.) go? & question\\
308 & zan'kai zanka šošra. & we (bonded) heal with (neutral). & social\\
309 & za zan'kai talta. & I to rest; pause with the chosen. & bond\\
310 & talm! & to rest; pause! & command\\
311 & ti zan'kai ronzidša. & you believe with the chosen. & bond\\
312 & zan susdista. & we feel. & statement\\
313 & tir druntarnun sairša, šaen tir nekun mitra. & you (pl.) to tell truth; align speech with reality druntarn, but you (pl.) touch pilot. & contrast\\
314 & zan zanka rukra. & we hurt with (neutral). & social\\
315 & zan zradra nai. & we do not remember. & negation\\
316 & zan seka trera. & we fly with (neutral). & social\\
317 & vomm! & wait! & command\\
318 & se vradgo, kur tir kathun škonša. & they start, because you (pl.) hide oath; promise that costs something. & cause\\
319 & zan heš nai. & we do not to listen. & negation\\
320 & se zuliun milša eh? & Does they walk indifferent void; cosmic uncaring? & question\\
321 & zan kakumun susdisra, šaen zan grimun rukra. & we feel sadness, but we hurt room. & contrast\\
322 & renumun susdism! & feel breath! & command\\
323 & zan dun, kur zan'kai mekun molta. & we to eat, because we (bonded) carry food. & cause\\
324 & zan gorirethun serta eh? & Does we to see; verify learn? & question\\
325 & ti kaiumun tovaršurta, šaen ti hílenhešun zleramra. & you endure by continuing the story-thread the chosen, but you escape hílen-heš. & contrast\\
326 & za hílenzulun vradgoša, šaen za vnitkeenun sokša. & I start hílen-zul, but I to take right. & contrast\\
327 & ti zan'kai šoßa. & you heal with the chosen. & bond\\
328 & ti gren. & you fight. & statement\\
329 & relnhílenun somm! & to sleep reln-hílen! & command\\
330 & tir vomra eh? & Does you (pl.) wait? & question\\
331 & se molenun venra, šaen se mórvretun šurra. & they to cross; to jump a threshold carry, but they to survive; endure mór-vret. & contrast\\
332 & ti tika zešta. & you help with (neutral). & social\\
333 & se vakiun hešta vem. & they to listen to wake (reported). & log\\
334 & rogresun susdism! & feel calm! & command\\
335 & zan'kai kordarun mitša. & we (bonded) touch tool. & statement\\
336 & zan tika som. & we to sleep with (neutral). & social\\
337 & se kailuvaun sairra, šaen se zulvrezomun kethta. & they to tell truth; align speech with reality kai-luva, but they to hold; keep zul-vrezom. & contrast\\
338 & se sálivenun rlasša eh? & Does they lose sáli-ven? & question\\
339 & zan seka vrenša. & we to lie; a lie with (neutral). & social\\
340 & zan'kai tovarun dunta . & we (bonded) to eat story-thread; chosen route (plain). & log\\
341 & tir nérun korra, šaen tir sinmodethun zleramša. & you (pl.) to work; fix star, but you (pl.) escape broken. & contrast\\
342 & tir sálikriun snonša. & you (pl.) arrive sáli-kri. & statement\\
343 & zan rogresenun grenša. & we fight calm. & statement\\
344 & tir vaskeniun darleša, šaen tir kathsusdisun gren. & you (pl.) live ship-keeper/pilot, but you (pl.) fight kath-susdis. & contrast\\
345 & ti zulkaíven eh? & Does you crossing the void with the chosen? & question\\
346 & se relnarun vakrekra vem. & they promise fold-gate/place (reported). & log\\
347 & zan'kai zaka vulra. & we (bonded) to speak; say with (neutral). & social\\
348 & zan'kai zan'kai molta. & we (bonded) carry with the chosen. & bond\\
349 & tir lumlavta, kur se zošoniun zlenasša. & you (pl.) refuse, because they build full. & cause\\
350 & za klitrinun rukta, šaen za sáliun drai. & I hurt port, but I to drift; wander stubborn defiant hope-light. & contrast\\
351 & zan keth nai. & we do not to hold; keep. & negation\\
352 & zan'kai dunša. & we (bonded) to eat. & statement\\
353 & se tika sokta. & they to take with (neutral). & social\\
354 & se šoš. & they heal. & statement\\
355 & se vyrun darleša. & they live cold. & statement\\
356 & ti naštera nai. & you do not trade. & negation\\
357 & za tika lon. & I know with (neutral). & social\\
358 & zan'kai mekarun ronzidra eh? & Does we (bonded) believe food? & question\\
359 & ti venumun heš, šaen ti nérvaskkaiun šur. & you to listen to cross; to jump a threshold, but you to survive; endure nér-vaskkai. & contrast\\
360 & zan škon, kur se hílenun sairta. & we hide, because they to tell truth; align speech with reality homesickmemory. & cause\\
361 & se sadra. & they send. & statement\\
362 & zan'kai šuriun tikša, šaen zan'kai sálitrelun kraßa. & we (bonded) teach to survive; endure, but we (bonded) rise sáli-trel. & contrast\\
363 & zan'kai tik nai. & we (bonded) do not teach. & negation\\
364 & zan'kai rukiun venra tov. & we (bonded) to cross; to jump a threshold hurt (inferred). & log\\
365 & zan se'kai lumlavta. & we refuse with the chosen. & bond\\
366 & zan'kai za'kai tovaršurra. & we (bonded) endure by continuing the story-thread with the chosen. & bond\\
367 & šošm! & heal! & command\\
368 & tir relnarun som, šaen tir trelenun lomakta. & you (pl.) to sleep fold-gate/place, but you (pl.) find tired. & contrast\\
369 & ti milra eh? & Does you walk? & question\\
370 & se mórsokun venta eh? & Does they to cross; to jump a threshold mór-sok? & question\\
371 & se vaskun susdisša. & they feel ship; vessel. & statement\\
372 & ti seka dunta. & you to eat with (neutral). & social\\
373 & se zan'kai grenša. & they fight with the chosen. & bond\\
374 & ti šoš nai. & you do not heal. & negation\\
375 & ti kathzlerun vul eh? & Does you to speak; say kath-zler? & question\\
376 & se se'kai lon. & they know with the chosen. & bond\\
377 & zan'kai se'kai tikta. & we (bonded) teach with the chosen. & bond\\
378 & tir tarnun rlasra. & you (pl.) lose money; credits; tokens. & statement\\
379 & ti zan'kai tik. & you teach with the chosen. & bond\\
380 & tir zan'kai krumša. & you (pl.) think with the chosen. & bond\\
381 & se mórzošonun lomakta, šaen se vakiun ronzid. & they find mór-zošon, but they believe to wake. & contrast\\
382 & ti zulkaíven, kur ti rervaumun šin. & you crossing the void with the chosen, because you to drink; drink/water metal. & cause\\
383 & zan'kai relnarun mista. & we (bonded) open fold-gate/place. & statement\\
384 & zlerm! & love! & command\\
385 & zan relnsadun vul . & we to speak; say reln-sad (plain). & log\\
386 & zan škiriun drai ser. & we to drift; wander run (seen/verified). & log\\
387 & ti gaktadta, kur ti mórsurenun zulkaívenša. & you fail, because you crossing the void with the chosen mór-suren. & cause\\
388 & se venša. & they to cross; to jump a threshold. & statement\\
389 & zan'kai zlerta nai. & we (bonded) do not love. & negation\\
390 & se drai, kur tir zuliun mit. & they to drift; wander, because you (pl.) touch indifferent void; cosmic uncaring. & cause\\
391 & ti ti'kai sairra. & you to tell truth; align speech with reality with the chosen. & bond\\
392 & zan'kai raskun draira eh? & Does we (bonded) to drift; wander fuel? & question\\
393 & zan'kai sairta nai. & we (bonded) do not to tell truth; align speech with reality. & negation\\
394 & vyrun krašm! & rise cold! & command\\
395 & torun sokm! & to take light! & command\\
396 & zan vradgoša. & we start. & statement\\
397 & za darta. & I to want; desire. & statement\\
398 & za vekun draiša eh? & Does I to drift; wander dark? & question\\
399 & zleramm! & escape! & command\\
400 & za zleram, kur se kaiumun vradgoša. & I escape, because they start the chosen. & cause\\
401 & zan'kai vekun milša. & we (bonded) walk dark. & statement\\
402 & za snonra. & I arrive. & statement\\
403 & ti ronzidša, kur zan rukiun venra. & you believe, because we to cross; to jump a threshold hurt. & cause\\
404 & ti ti'kai dolim. & you turn with the chosen. & bond\\
405 & se dar eh? & Does they to want; desire? & question\\
406 & tir lumlav. & you (pl.) refuse. & statement\\
407 & zan kaigrimun keth eh? & Does we to hold; keep kai-grim? & question\\
408 & susdism! & feel! & command\\
409 & zan tika snonta. & we arrive with (neutral). & social\\
410 & zan darra eh? & Does we to want; desire? & question\\
411 & tir ti'kai hešra. & you (pl.) to listen with the chosen. & bond\\
412 & zan sizruvša, kur zan'kai dinumun lonša. & we forget, because we (bonded) know signal. & cause\\
413 & zulstokzorun zlenasm! & build zul-stokzor! & command\\
414 & za tika sairra. & I to tell truth; align speech with reality with (neutral). & social\\
415 & ti vavun šinta eh? & Does you to drink; drink/water day? & question\\
416 & zethun dolimm! & turn air; oxygen supply! & command\\
417 & se korra, kur tir vaskun zlerra. & they to work; fix, because you (pl.) love ship; vessel. & cause\\
418 & zan'kai dinenun rukta. & we (bonded) hurt signal. & statement\\
419 & zan seka gliekta. & we fall with (neutral). & social\\
420 & misumun zimm! & pay open! & command\\
421 & ti móriun venta eh? & Does you to cross; to jump a threshold laughter in the dark? & question\\
422 & korm! & to work; fix! & command\\
423 & ti mórvulun sairša eh? & Does you to tell truth; align speech with reality mór-vul? & question\\
424 & zan tovarenun šur reln. & we to survive; endure story-thread; chosen route (fold-dream). & log\\
425 & tir zan'kai susdista. & you (pl.) feel with the chosen. & bond\\
426 & zan škir, kur se stušun škon. & we run, because they hide truth. & cause\\
427 & zan škonra. & we hide. & statement\\
428 & naštearun šinm! & to drink; drink/water trade! & command\\
429 & ti tika sairra. & you to tell truth; align speech with reality with (neutral). & social\\
430 & zan'kai nérun darša eh? & Does we (bonded) to want; desire star? & question\\
431 & zan vak. & we to wake. & statement\\
432 & se raskun krašta tov. & they rise fuel (inferred). & log\\
433 & za tikenun lirta eh? & Does I to give teach? & question\\
434 & zan'kai zulserun vilkalša. & we (bonded) forgive zulser. & statement\\
435 & tir škirumun darša. & you (pl.) to want; desire run. & statement\\
436 & zan'kai se'kai vrenta. & we (bonded) to lie; a lie with the chosen. & bond\\
437 & rerlerenun zimm! & pay water! & command\\
438 & zan'kai rukta eh? & Does we (bonded) hurt? & question\\
439 & zan vasksadun gliekra. & we fall vask-sad. & statement\\
440 & darlem! & live! & command\\
441 & za vrunumun hešta . & I to listen fear (plain). & log\\
442 & krašumun zlenasm! & build rise! & command\\
443 & ti šur nai. & you do not to survive; endure. & negation\\
444 & za nérun lalgemra. & I leave star. & statement\\
445 & ti zethun vrenra. & you to lie; a lie air; oxygen supply. & statement\\
446 & za seka ronzidta. & I believe with (neutral). & social\\
447 & zan venta, kur zan našethun lumlavta. & we to cross; to jump a threshold, because we refuse hand; grip. & cause\\
448 & za zularun vretta eh? & Does I push indifferent void; cosmic uncaring? & question\\
449 & ti seka lesša. & you hate with (neutral). & social\\
450 & za zleramenun zleramra eh? & Does I escape escape? & question\\
451 & zan'kai milša eh? & Does we (bonded) walk? & question\\
452 & tir zaka nat. & you (pl.) doubt with (neutral). & social\\
453 & se talenun zlusša. & they buy to rest; pause. & statement\\
454 & za vrusenun ser . & I to see; verify bed (plain). & log\\
455 & zan'kai siddreumun škirša. & we (bonded) run cold. & statement\\
456 & zan'kai darša nai. & we (bonded) do not to want; desire. & negation\\
457 & zan vulta. & we to speak; say. & statement\\
458 & ti vakrekra, kur tir sáliun zlenas. & you promise, because you (pl.) build stubborn defiant hope-light. & cause\\
459 & zrosethun sizruvm! & forget false! & command\\
460 & zan'kai draira, kur tir kathgrimun milša. & we (bonded) to drift; wander, because you (pl.) walk kath-grim. & cause\\
461 & ti se'kai trera. & you fly with the chosen. & bond\\
462 & ti ronzidra eh? & Does you believe? & question\\
463 & zan lomak, kur se zulkrašun mil. & we find, because they walk zul-kraš. & cause\\
464 & ti susdisša nai. & you do not feel. & negation\\
465 & se kaizurun dolim vem. & they turn kai-zur (reported). & log\\
466 & zan krumenun sairra. & we to tell truth; align speech with reality think. & statement\\
467 & ti zlusta eh? & Does you buy? & question\\
468 & zan mórvrusun venra ser. & we to cross; to jump a threshold mór-vrus (seen/verified). & log\\
469 & ti škonra eh? & Does you hide? & question\\
470 & ti vakta eh? & Does you to wake? & question\\
471 & kešarun venm! & to cross; to jump a threshold wire; line; cable! & command\\
472 & se darle, kur se murenun draira. & they live, because they to drift; wander mouth. & cause\\
473 & zan'kai mórtrerun draira. & we (bonded) to drift; wander mór-trer. & statement\\
474 & tir snaš eh? & Does you (pl.) receive? & question\\
475 & za vaskun snaßa vem. & I receive ship; vessel (reported). & log\\
476 & zan vilkal nai. & we do not forgive. & negation\\
477 & zan'kai za'kai dar. & we (bonded) to want; desire with the chosen. & bond\\
478 & zan'kai kridiun sairša. & we (bonded) to tell truth; align speech with reality go. & statement\\
479 & zan'kai zulun gaktadra eh? & Does we (bonded) fail indifferent void; cosmic uncaring? & question\\
480 & ti mórsadun venta . & you to cross; to jump a threshold mór-sad (plain). & log\\
481 & ti tika zradša. & you remember with (neutral). & social\\
482 & zan heš eh? & Does we to listen? & question\\
483 & se snuvenun ven . & they to cross; to jump a threshold safe (plain). & log\\
484 & zan'kai nérun vilkalra, šaen zan'kai nérun zleramra. & we (bonded) forgive star, but we (bonded) escape star. & contrast\\
485 & draiiun škonm! & hide to drift; wander! & command\\
486 & zan'kai se'kai lomakta. & we (bonded) find with the chosen. & bond\\
487 & natethun šinm! & to drink; drink/water doubt! & command\\
488 & se mórzošonun darša, šaen se vaskstilun kethša. & they to want; desire mór-zošon, but they to hold; keep vask-stil. & contrast\\
489 & zan'kai kathun sok . & we (bonded) to take oath; promise that costs something (plain). & log\\
490 & zleramenun tikm! & teach escape! & command\\
491 & za mórzlenasun sizruvra. & I forget mór-zlenas. & statement\\
492 & zan'kai kaiumun korta ser. & we (bonded) to work; fix the chosen (seen/verified). & log\\
493 & za zanka zlusša. & I buy with (neutral). & social\\
494 & zan'kai mitethun hešta eh? & Does we (bonded) to listen touch? & question\\
495 & zan'kai vrenta nai. & we (bonded) do not to lie; a lie. & negation\\
496 & za tika kethta. & I to hold; keep with (neutral). & social\\
497 & tir tovaršurša, kur za mórzrosun gliekra. & you (pl.) endure by continuing the story-thread, because I fall mór-zros. & cause\\
498 & zan'kai darta, kur zan zulrelnarun zler. & we (bonded) to want; desire, because we love zulrelnar. & cause\\
499 & za vasksokun kor. & I to work; fix vask-sok. & statement\\
500 & zan'kai relndonun venša reln. & we (bonded) to cross; to jump a threshold reln-don (fold-dream). & log\\
501 & za se'kai kridša. & I go with the chosen. & bond\\
502 & tir snašra, kur zan'kai stokzorethun vak. & you (pl.) receive, because we (bonded) to wake key. & cause\\
503 & za zanka gliek. & I fall with (neutral). & social\\
504 & se hílenzrismidun škunta tov. & they watch hílen-zrismid (inferred). & log\\
505 & ti nérsnustatun tovaršurša, šaen ti nérun susdis. & you endure by continuing the story-thread nér-snustat, but you feel star. & contrast\\
506 & zulmolun krumm! & think zulmol! & command\\
507 & se krumša eh? & Does they think? & question\\
508 & ti zulkaíven. & you crossing the void with the chosen. & statement\\
509 & zan'kai zaka trera. & we (bonded) fly with (neutral). & social\\
510 & drunrelnkaiun vomm! & wait drunrelnkai! & command\\
511 & zan relndonun lonra tov. & we know reln-don (inferred). & log\\
512 & za draiethun vilkalta tov. & I forgive to drift; wander (inferred). & log\\
513 & ti raskiun darra eh? & Does you to want; desire fuel? & question\\
514 & ti vretra. & you push. & statement\\
515 & se korša nai. & they do not to work; fix. & negation\\
516 & tir tarnun šošta eh? & Does you (pl.) heal money; credits; tokens? & question\\
517 & mism! & open! & command\\
518 & za draita nai. & I do not to drift; wander. & negation\\
519 & gorirm! & learn! & command\\
520 & se se'kai draita. & they to drift; wander with the chosen. & bond\\
521 & se kilun vren tov. & they to lie; a lie tool (inferred). & log\\
522 & ti vrunethun heßa, šaen ti vasksairun vilkalša. & you to listen fear, but you forgive vask-sair. & contrast\\
523 & ti hílenun hešta eh? & Does you to listen homesickmemory? & question\\
524 & zan zan'kai serša. & we to see; verify with the chosen. & bond\\
525 & ti druniiun škunša, šaen ti zethun rlasša. & you watch dread, but you lose air; oxygen supply. & contrast\\
526 & za relnarun lalgemra. & I leave fold-gate/place. & statement\\
527 & za škiriun vakta vem. & I to wake run (reported). & log\\
528 & ti tika ruk. & you hurt with (neutral). & social\\
529 & ti zaka zaknimra. & you close with (neutral). & social\\
530 & zan'kai seka venša. & we (bonded) to cross; to jump a threshold with (neutral). & social\\
531 & ti tika sadra. & you send with (neutral). & social\\
532 & se luvaiun škonta tov. & they hide name (inferred). & log\\
533 & tir za'kai vret. & you (pl.) push with the chosen. & bond\\
534 & zan'kai snašta nai. & we (bonded) do not receive. & negation\\
535 & se zlusta nai. & they do not buy. & negation\\
536 & zan škunethun zleramta reln. & we escape watch (fold-dream). & log\\
537 & za talenun zlusra, šaen za mórumun vakša. & I buy to rest; pause, but I to wake laughter in the dark. & contrast\\
538 & ti mórenun tal eh? & Does you to rest; pause laughter in the dark? & question\\
539 & zan kathun zulkaívenra reln. & we crossing the void with the chosen oath; promise that costs something (fold-dream). & log\\
540 & tir nat, kur zan drunkairenun vilkal. & you (pl.) doubt, because we forgive drunkairen. & cause\\
541 & zan kaiumun hešta . & we to listen the chosen (plain). & log\\
542 & tir zaka rlasša. & you (pl.) lose with (neutral). & social\\
543 & tir ti'kai zradra. & you (pl.) remember with the chosen. & bond\\
544 & nérun donm! & break star! & command\\
545 & tir zleramša eh? & Does you (pl.) escape? & question\\
546 & ti vomethun zaknimša eh? & Does you close wait? & question\\
547 & tir zaka darra. & you (pl.) to want; desire with (neutral). & social\\
548 & zan donra eh? & Does we break? & question\\
549 & zan'kai dunša, kur se kairelnkaiun šoßa. & we (bonded) to eat, because they heal kairelnkai. & cause\\
550 & zan'kai roriun somša. & we (bonded) to sleep station. & statement\\
551 & zan tika darra. & we to want; desire with (neutral). & social\\
552 & se relnarun zlus . & they buy fold-gate/place (plain). & log\\
553 & ti tulmukiun korta ser. & you to work; fix wound (seen/verified). & log\\
554 & kordun dunm! & to eat tool! & command\\
555 & zan'kai nérun susdis, šaen zan'kai vekun ven. & we (bonded) feel star, but we (bonded) to cross; to jump a threshold dark. & contrast\\
556 & zan tika trera. & we fly with (neutral). & social\\
557 & torun darm! & to want; desire light! & command\\
558 & zan'kai talša. & we (bonded) to rest; pause. & statement\\
559 & se gliekra nai. & they do not fall. & negation\\
560 & se torun zaknim. & they close light. & statement\\
561 & za kethra, kur tir vyrun šin. & I to hold; keep, because you (pl.) to drink; drink/water cold. & cause\\
562 & za vaskniritun nat eh? & Does I doubt vask-nirit? & question\\
563 & se hílenskultiun zlerša tov. & they love hílen-skulti (inferred). & log\\
564 & se renarun tat, šaen se rerlerun sokta. & they choose breath, but they to take water. & contrast\\
565 & raskumun vilkalm! & forgive fuel! & command\\
566 & tir sinmodethun vakra reln. & you (pl.) to wake broken (fold-dream). & log\\
567 & ašun darlem! & live heat! & command\\
568 & za tovarrogresun heš, šaen za stokzorethun dunta. & I to listen tovar-rogres, but I to eat key. & contrast\\
569 & se vaskškonun škunša . & they watch vask-škon (plain). & log\\
570 & zan'kai kathsiddreun ronzidta . & we (bonded) believe kath-siddre (plain). & log\\
571 & ti šurta, kur zan vnunvutarun heš. & you to survive; endure, because we to listen captain. & cause\\
572 & tir se'kai vul. & you (pl.) to speak; say with the chosen. & bond\\
573 & zan lonenun zim eh? & Does we pay know? & question\\
574 & tir kathun vomra, šaen tir golarun vakta. & you (pl.) wait oath; promise that costs something, but you (pl.) to wake door. & contrast\\
575 & za vul. & I to speak; say. & statement\\
576 & ti dar. & you to want; desire. & statement\\
577 & zan darša nai. & we do not to want; desire. & negation\\
578 & kathlesun snonm! & arrive kath-les! & command\\
579 & kathkaiun tikm! & teach oath-kin! & command\\
580 & tir skultiun trera. & you (pl.) fly mechanic. & statement\\
581 & nekumun zradm! & remember pilot! & command\\
582 & ti krumta, kur se zaknimiun sad. & you think, because they send close. & cause\\
583 & ti darša eh? & Does you to want; desire? & question\\
584 & zan'kai vradgota eh? & Does we (bonded) start? & question\\
585 & vyrarun korm! & to work; fix cold! & command\\
586 & zan druniiun dar reln. & we to want; desire dread (fold-dream). & log\\
587 & za vradgo eh? & Does I start? & question\\
588 & tir kaiumun vakrekta, šaen tir tovaršvonun heßa. & you (pl.) promise the chosen, but you (pl.) to listen tovar-švon. & contrast\\
589 & ti rogresumun nat, šaen ti skultiumun somra. & you doubt calm, but you to sleep mechanic. & contrast\\
590 & za tovarrelnarun atsnukra eh? & Does I come tovarrelnar? & question\\
591 & ti vrunumun zaknimra ser. & you close fear (seen/verified). & log\\
592 & ti vilkalša, kur zan krarkesenun darra. & you forgive, because we to want; desire night. & cause\\
593 & tir zaka somta. & you (pl.) to sleep with (neutral). & social\\
594 & ti za'kai grenra. & you fight with the chosen. & bond\\
595 & hílenkešun zulkaívenm! & crossing the void with the chosen hílen-keš! & command\\
596 & ti našteumun talša, šaen ti treethun lon. & you to rest; pause trade, but you know fly. & contrast\\
597 & se vakrek nai. & they do not promise. & negation\\
598 & mitm! & touch! & command\\
599 & tir hílenkrotun sairta. & you (pl.) to tell truth; align speech with reality hílen-krot. & statement\\
600 & se kérunarun snaßa, šaen se kordun škon. & they receive made-meaning, but they hide tool. & contrast\\
601 & za tika šoš. & I heal with (neutral). & social\\
602 & zan vrenša. & we to lie; a lie. & statement\\
603 & tir kordun korša, šaen tir krenarun zešta. & you (pl.) to work; fix tool, but you (pl.) help anger. & contrast\\
604 & zan kairelnarun škirta. & we run kairelnar. & statement\\
605 & zan venša eh? & Does we to cross; to jump a threshold? & question\\
606 & se se'kai škun. & they watch with the chosen. & bond\\
607 & zan'kai talra nai. & we (bonded) do not to rest; pause. & negation\\
608 & tir rimenun zaknimša vem. & you (pl.) close angry (reported). & log\\
609 & tir dun eh? & Does you (pl.) to eat? & question\\
610 & se zan'kai tovaršurša. & they endure by continuing the story-thread with the chosen. & bond\\
611 & ti zulun zlenasša vem. & you build indifferent void; cosmic uncaring (reported). & log\\
612 & tir sokra nai. & you (pl.) do not to take. & negation\\
613 & zan gorirra eh? & Does we learn? & question\\
614 & tir zulkaíven. & you (pl.) crossing the void with the chosen. & statement\\
615 & zan za'kai korša. & we to work; fix with the chosen. & bond\\
616 & zan rerlerumun škirta. & we run water. & statement\\
617 & zan dunra nai. & we do not to eat. & negation\\
618 & zan zanka gorirra. & we learn with (neutral). & social\\
619 & se drunun zlusra eh? & Does they buy dread; the weight of meaninglessness? & question\\
620 & se zaka gorirta. & they learn with (neutral). & social\\
621 & ti šurta eh? & Does you to survive; endure? & question\\
622 & tir dunra eh? & Does you (pl.) to eat? & question\\
623 & zan'kai sairra eh? & Does we (bonded) to tell truth; align speech with reality? & question\\
624 & zan tika draiša. & we to drift; wander with (neutral). & social\\
625 & se ašun vradgota. & they start heat. & statement\\
626 & za somta. & I to sleep. & statement\\
627 & se renumun som. & they to sleep breath. & statement\\
628 & ti kordun snaßa eh? & Does you receive tool? & question\\
629 & tir raskarun vul eh? & Does you (pl.) to speak; say fuel? & question\\
630 & zan ti'kai zlusra. & we buy with the chosen. & bond\\
631 & tir dinenun sadša, šaen tir vaskun tal. & you (pl.) send signal, but you (pl.) to rest; pause ship; vessel. & contrast\\
632 & ti drunkevralun vulša, šaen ti kisiun sair. & you to speak; say drun-kevral, but you to tell truth; align speech with reality engine. & contrast\\
633 & zan'kai gorirra. & we (bonded) learn. & statement\\
634 & za korša. & I to work; fix. & statement\\
635 & atsnukm! & come! & command\\
636 & se dunta nai. & they do not to eat. & negation\\
637 & se vakrekša eh? & Does they promise? & question\\
638 & se tikta. & they teach. & statement\\
639 & ti hílenkrotun lirša, šaen ti tovarrlotun zlerta. & you to give hílen-krot, but you love tovar-rlot. & contrast\\
640 & zan'kai hílenzlusun dunta. & we (bonded) to eat hílen-zlus. & statement\\
641 & ti hešta, kur za mórsizruvun venša. & you to listen, because I to cross; to jump a threshold mór-sizruv. & cause\\
642 & zan'kai drunvnunvutun našteša. & we (bonded) trade drun-vnunvut. & statement\\
643 & se nérronzidun zulkaíventa eh? & Does they crossing the void with the chosen nér-ronzid? & question\\
644 & kailuvaun tovaršurm! & endure by continuing the story-thread kai-luva! & command\\
645 & tir stušun lumlavra ser. & you (pl.) refuse truth (seen/verified). & log\\
646 & ti kathtreun les eh? & Does you hate kath-tre? & question\\
647 & zan'kai vilkalta eh? & Does we (bonded) forgive? & question\\
648 & zan zulvaskenun sokta, šaen zan vekumun šur. & we to take zul-vasken, but we to survive; endure dark. & contrast\\
649 & tir sadša nai. & you (pl.) do not send. & negation\\
650 & tir kathun gaktadra. & you (pl.) fail oath; promise that costs something. & statement\\
651 & natm! & doubt! & command\\
652 & tir mórtalun krumra, šaen tir kainérun somša. & you (pl.) think mórtal, but you (pl.) to sleep kai-nér. & contrast\\
653 & zan'kai zlerra. & we (bonded) love. & statement\\
654 & tir somta. & you (pl.) to sleep. & statement\\
655 & zan gorirra, kur ti kathun snaš. & we learn, because you receive oath; promise that costs something. & cause\\
656 & tir snonša eh? & Does you (pl.) arrive? & question\\
657 & glekumun somm! & to sleep joy! & command\\
658 & ti raskarun ronzid vem. & you believe fuel (reported). & log\\
659 & tir kethenun darša. & you (pl.) to want; desire to hold; keep. & statement\\
660 & kathzešun zleramm! & escape kath-zeš! & command\\
661 & zan kathtreun sokša eh? & Does we to take kath-tre? & question\\
662 & zan nekun draiša tov. & we to drift; wander pilot (inferred). & log\\
663 & zan zan'kai kridra. & we go with the chosen. & bond\\
664 & zan ašun vomta. & we wait heat. & statement\\
665 & dunm! & to eat! & command\\
666 & tir ti'kai kethra. & you (pl.) to hold; keep with the chosen. & bond\\
667 & zan'kai vrunarun krašta eh? & Does we (bonded) rise fear? & question\\
668 & zan darta nai. & we do not to want; desire. & negation\\
669 & snašm! & receive! & command\\
670 & za kathun mitta tov. & I touch oath; promise that costs something (inferred). & log\\
671 & zan'kai kordun kethra, šaen zan'kai vyrun tovaršurra. & we (bonded) to hold; keep tool, but we (bonded) endure by continuing the story-thread cold. & contrast\\
672 & tir zanka misša. & you (pl.) open with (neutral). & social\\
673 & tir škirta, kur zan relntrerun darta. & you (pl.) run, because we to want; desire reln-trer. & cause\\
674 & tir vomša eh? & Does you (pl.) wait? & question\\
675 & tir vul. & you (pl.) to speak; say. & statement\\
676 & zan'kai zan'kai dun. & we (bonded) to eat with the chosen. & bond\\
677 & vakm! & to wake! & command\\
678 & se lirša. & they to give. & statement\\
679 & za raskun kraš, šaen za torun heš. & I rise fuel, but I to listen light. & contrast\\
680 & tir zešra, kur za venethun lirta. & you (pl.) help, because I to give to cross; to jump a threshold. & cause\\
681 & kriarun snonm! & arrive true! & command\\
682 & tir zler eh? & Does you (pl.) love? & question\\
683 & za zlenasta, kur tir zlukraarun venta. & I build, because you (pl.) to cross; to jump a threshold friend. & cause\\
684 & ti ti'kai zlenas. & you build with the chosen. & bond\\
685 & tir se'kai dun. & you (pl.) to eat with the chosen. & bond\\
686 & zan'kai darta eh? & Does we (bonded) to want; desire? & question\\
687 & tir kethra nai. & you (pl.) do not to hold; keep. & negation\\
688 & zan relntrerun somra tov. & we to sleep reln-trer (inferred). & log\\
689 & ti zethun šošra. & you heal air; oxygen supply. & statement\\
690 & za se'kai zešta. & I help with the chosen. & bond\\
691 & za mórlokun kraßa, šaen za zleramenun dunta. & I rise mór-lok, but I to eat escape. & contrast\\
692 & tir šošiun dunša eh? & Does you (pl.) to eat heal? & question\\
693 & se luvaun darša. & they to want; desire name. & statement\\
694 & zan gaktadra. & we fail. & statement\\
695 & se vrenša, kur tir vaskrukun vrenra. & they to lie; a lie, because you (pl.) to lie; a lie vask-ruk. & cause\\
696 & zan'kai snaßa nai. & we (bonded) do not receive. & negation\\
697 & ti vaskkaiun sairša. & you to tell truth; align speech with reality ship-kin; crew-as-family. & statement\\
698 & tir draiša eh? & Does you (pl.) to drift; wander? & question\\
699 & se dar. & they to want; desire. & statement\\
700 & draim! & to drift; wander! & command\\
701 & zan škonra nai. & we do not hide. & negation\\
702 & za tika treta. & I fly with (neutral). & social\\
703 & ti rlasethun zulkaívenra tov. & you crossing the void with the chosen lose (inferred). & log\\
704 & tir zan'kai zlus. & you (pl.) buy with the chosen. & bond\\
705 & zan sáliun sairša eh? & Does we to tell truth; align speech with reality stubborn defiant hope-light? & question\\
706 & ti šin nai. & you do not to drink; drink/water. & negation\\
707 & snonumun dolimm! & turn arrive! & command\\
708 & za kathtovarun som tov. & I to sleep kathtovar (inferred). & log\\
709 & tir darleta nai. & you (pl.) do not live. & negation\\
710 & za don, kur se mekun sadša. & I break, because they send food. & cause\\
711 & zan'kai draienun snašra eh? & Does we (bonded) receive to drift; wander? & question\\
712 & zan'kai zulun vilkalta. & we (bonded) forgive indifferent void; cosmic uncaring. & statement\\
713 & za zuldimteun natta eh? & Does I doubt zul-dimte? & question\\
714 & zan relnethun zaknim. & we close foldcraft; seam-sense. & statement\\
715 & ti zaka sokra. & you to take with (neutral). & social\\
716 & za ti'kai ven. & I to cross; to jump a threshold with the chosen. & bond\\
717 & zan vomra. & we wait. & statement\\
718 & zan gaktadta, kur zan vilkalenun somta. & we fail, because we to sleep forgive. & cause\\
719 & se sivarun dar. & they to want; desire lock; seal. & statement\\
720 & se ti'kai venša. & they to cross; to jump a threshold with the chosen. & bond\\
721 & tir vulša, kur zan relndonun zulkaíven. & you (pl.) to speak; say, because we crossing the void with the chosen reln-don. & cause\\
722 & kathun milm! & walk oath; promise that costs something! & command\\
723 & tir vulra. & you (pl.) to speak; say. & statement\\
724 & zan'kai nérgaktadun treša vem. & we (bonded) fly nér-gaktad (reported). & log\\
725 & se tal eh? & Does they to rest; pause? & question\\
726 & se zethun korra eh? & Does they to work; fix air; oxygen supply? & question\\
727 & zan'kai drai. & we (bonded) to drift; wander. & statement\\
728 & se rogresiun zleramta. & they escape calm. & statement\\
729 & rerlerenun lirm! & to give water! & command\\
730 & zan tovarvulun gliek, šaen zan vasksadun molra. & we fall tovar-vul, but we carry vask-sad. & contrast\\
731 & tovarun zaknimm! & close story-thread; chosen route! & command\\
732 & zan'kai snašra nai. & we (bonded) do not receive. & negation\\
733 & ti mórumun vilkalša eh? & Does you forgive laughter in the dark? & question\\
734 & se seka som. & they to sleep with (neutral). & social\\
735 & zan sairta nai. & we do not to tell truth; align speech with reality. & negation\\
736 & nérun venm! & to cross; to jump a threshold star! & command\\
737 & vomenun sairm! & to tell truth; align speech with reality wait! & command\\
738 & tir dun. & you (pl.) to eat. & statement\\
739 & za ti'kai sokša. & I to take with the chosen. & bond\\
740 & se naštearun talta tov. & they to rest; pause trade (inferred). & log\\
741 & bolun atsnukm! & come door; hatch! & command\\
742 & ti tovarmilun vul eh? & Does you to speak; say tovarmil? & question\\
743 & kordun snašm! & receive tool! & command\\
744 & ti zlerra eh? & Does you love? & question\\
745 & tir tatša. & you (pl.) choose. & statement\\
746 & se vomra eh? & Does they wait? & question\\
747 & za tatša. & I choose. & statement\\
748 & vakarun hešm! & to listen to wake! & command\\
749 & se hílenun draita eh? & Does they to drift; wander homesickmemory? & question\\
750 & tir grenarun donra eh? & Does you (pl.) break fight? & question\\
751 & krašm! & rise! & command\\
752 & tir vavumun milša. & you (pl.) walk day. & statement\\
753 & rimenun mitm! & touch angry! & command\\
754 & se kaizlenasun sadra vem. & they send kai-zlenas (reported). & log\\
755 & ti našethun škirša, šaen ti vaskun zleramra. & you run hand; grip, but you escape ship; vessel. & contrast\\
756 & za dar nai. & I do not to want; desire. & negation\\
757 & zan golumun somta ser. & we to sleep door (seen/verified). & log\\
758 & zan'kai krumra, kur zan vaskkaiun venša. & we (bonded) think, because we to cross; to jump a threshold ship-kin; crew-as-family. & cause\\
759 & ti zan'kai šin. & you to drink; drink/water with the chosen. & bond\\
760 & gaktadarun talm! & to rest; pause fail! & command\\
761 & se zulkaíven eh? & Does they crossing the void with the chosen? & question\\
762 & ti snaßa nai. & you do not receive. & negation\\
763 & zlusiun korm! & to work; fix buy! & command\\
764 & zan'kai našteta, kur ti natethun serra. & we (bonded) trade, because you to see; verify doubt. & cause\\
765 & za gaktadra nai. & I do not fail. & negation\\
766 & za somra eh? & Does I to sleep? & question\\
767 & relnethethun dunm! & to eat foldcraft; seam-sense! & command\\
768 & mekun lomakm! & find food! & command\\
769 & tir sadta eh? & Does you (pl.) send? & question\\
770 & ti korra nai. & you do not to work; fix. & negation\\
771 & za ti'kai dunta. & I to eat with the chosen. & bond\\
772 & ti vradgota. & you start. & statement\\
773 & zan snaßa eh? & Does we receive? & question\\
774 & vretm! & push! & command\\
775 & zan zulzleramun natra eh? & Does we doubt zul-zleram? & question\\
776 & tir heßa nai. & you (pl.) do not to listen. & negation\\
777 & tir mekun dun vem. & you (pl.) to eat food (reported). & log\\
778 & se gorir, kur za hílenun tal. & they learn, because I to rest; pause homesickmemory. & cause\\
779 & tir zlenasta nai. & you (pl.) do not build. & negation\\
780 & sálišvonun gliekm! & fall sáli-švon! & command\\
781 & se sairta eh? & Does they to tell truth; align speech with reality? & question\\
782 & za vaskkaiun škonša eh? & Does I hide ship-kin; crew-as-family? & question\\
783 & tir vulta, kur zan'kai hílentalun lir. & you (pl.) to speak; say, because we (bonded) to give hílental. & cause\\
784 & ti zanka kraš. & you rise with (neutral). & social\\
785 & ti zulkaívenra. & you crossing the void with the chosen. & statement\\
786 & se darle, kur se vevun zulkaívenša. & they live, because they crossing the void with the chosen problem. & cause\\
787 & rogresiun vulm! & to speak; say calm! & command\\
788 & rukm! & hurt! & command\\
789 & se gaktadra eh? & Does they fail? & question\\
790 & zan'kai vomta, kur za vaskdarleun hešta. & we (bonded) wait, because I to listen vask-darle. & cause\\
791 & skirtusun somm! & to sleep guide! & command\\
792 & ti nérklinun darša eh? & Does you to want; desire nér-klin? & question\\
793 & zan'kai seka darša. & we (bonded) to want; desire with (neutral). & social\\
794 & tir vilkalša, kur za mitumun vretša. & you (pl.) forgive, because I push touch. & cause\\
795 & ti za'kai vrenra. & you to lie; a lie with the chosen. & bond\\
796 & nérvnunvutun venm! & to cross; to jump a threshold nér-vnunvut! & command\\
797 & zan'kai mórronzidun rukša, šaen zan'kai hílenkešun krumta. & we (bonded) hurt mór-ronzid, but we (bonded) think hílen-keš. & contrast\\
798 & tir drunrelnkaiun gaktadra. & you (pl.) fail drunrelnkai. & statement\\
799 & tir sadra nai. & you (pl.) do not send. & negation\\
800 & tir treša, kur se gaktadiun zlenasra. & you (pl.) fly, because they build fail. & cause\\
801 & zan'kai tovaršur nai. & we (bonded) do not endure by continuing the story-thread. & negation\\
802 & ti kordarun šošra vem. & you heal tool (reported). & log\\
803 & za darta eh? & Does I to want; desire? & question\\
804 & se zan'kai zeßa. & they help with the chosen. & bond\\
805 & se kridra. & they go. & statement\\
806 & ti donša, kur se zulškikun tik. & you break, because they teach zul-škik. & cause\\
807 & gliekm! & fall! & command\\
808 & ti tatra eh? & Does you choose? & question\\
809 & tir relnlokun zleramša. & you (pl.) escape reln-lok. & statement\\
810 & zan'kai ven eh? & Does we (bonded) to cross; to jump a threshold? & question\\
811 & se venra. & they to cross; to jump a threshold. & statement\\
812 & zan sizruvša. & we forget. & statement\\
813 & ti zan'kai ronzid. & you believe with the chosen. & bond\\
814 & kathvakrekun somm! & to sleep kath-vakrek! & command\\
815 & zan'kai šurta eh? & Does we (bonded) to survive; endure? & question\\
816 & zan'kai dunenun kethta tov. & we (bonded) to hold; keep to eat (inferred). & log\\
817 & za molarun lir . & I to give carry (plain). & log\\
818 & zleriun draim! & to drift; wander love! & command\\
819 & se krarkesiun dolim. & they turn night. & statement\\
820 & vuvarun dolimm! & turn weak! & command\\
821 & ti kakiun snonra, šaen ti kathun kridša. & you arrive sadness, but you go oath; promise that costs something. & contrast\\
822 & rukiun atsnukm! & come hurt! & command\\
823 & zan sáliun drai. & we to drift; wander stubborn defiant hope-light. & statement\\
824 & se donta, kur zan'kai našumun susdista. & they break, because we (bonded) feel hand; grip. & cause\\
825 & za zulkaívenra. & I crossing the void with the chosen. & statement\\
826 & se gren eh? & Does they fight? & question\\
827 & se sáliun molša vem. & they carry stubborn defiant hope-light (reported). & log\\
828 & kethm! & to hold; keep! & command\\
829 & zan'kai sairra. & we (bonded) to tell truth; align speech with reality. & statement\\
830 & tir kaizurun lonta, šaen tir bolethun lalgemra. & you (pl.) know kai-zur, but you (pl.) leave door; hatch. & contrast\\
831 & zan zulkaívenša nai. & we do not crossing the void with the chosen. & negation\\
832 & se dunarun serra eh? & Does they to see; verify to eat? & question\\
833 & ti hešra. & you to listen. & statement\\
834 & zan drunethun škun, šaen zan zleriun zulkaíventa. & we watch dread; the weight of meaninglessness, but we crossing the void with the chosen love. & contrast\\
835 & zan'kai zulstokzorun zulkaívenša. & we (bonded) crossing the void with the chosen zul-stokzor. & statement\\
836 & torun vakm! & to wake light! & command\\
837 & zan'kai vnitkeethun vak . & we (bonded) to wake right (plain). & log\\
838 & zan vomša nai. & we do not wait. & negation\\
839 & ti darra, kur tir zrismidumun zaknim. & you to want; desire, because you (pl.) close map. & cause\\
840 & za vilkal nai. & I do not forgive. & negation\\
841 & zan'kai lalgem eh? & Does we (bonded) leave? & question\\
842 & zan'kai zleramenun lirša. & we (bonded) to give escape. & statement\\
843 & lokmarethun šošm! & heal deep! & command\\
844 & za zulkaíven nai. & I do not crossing the void with the chosen. & negation\\
845 & zan'kai dar. & we (bonded) to want; desire. & statement\\
846 & zan sair eh? & Does we to tell truth; align speech with reality? & question\\
847 & zan natta, kur se vaskkordun vren. & we doubt, because they to lie; a lie vask-kil. & cause\\
848 & ti vaskkaiun don ser. & you break ship-kin; crew-as-family (seen/verified). & log\\
849 & za kaireniun draita. & I to drift; wander bond-restorer; group glue-person. & statement\\
850 & se ti'kai tatša. & they choose with the chosen. & bond\\
851 & se raskun zlenasša eh? & Does they build fuel? & question\\
852 & zan'kai škunša, kur za kešun zešta. & we (bonded) watch, because I help wire; line; cable. & cause\\
853 & zan'kai lirta eh? & Does we (bonded) to give? & question\\
854 & ti seka vradgoša. & you start with (neutral). & social\\
855 & tir lalgemša, kur za nérkrumun venra. & you (pl.) leave, because I to cross; to jump a threshold nér-krum. & cause\\
856 & za kordun vulta eh? & Does I to speak; say tool? & question\\
857 & rlotiun škirm! & run enemy! & command\\
858 & se venta eh? & Does they to cross; to jump a threshold? & question\\
859 & zan zulun zleramra tov. & we escape indifferent void; cosmic uncaring (inferred). & log\\
860 & zan'kai seka darlera. & we (bonded) live with (neutral). & social\\
861 & tir dunenun nat. & you (pl.) doubt to eat. & statement\\
862 & zan rlasethun tikra, šaen zan tovarrlasun susdista. & we teach lose, but we feel tovar-rlas. & contrast\\
863 & milm! & walk! & command\\
864 & zan za'kai šoßa. & we heal with the chosen. & bond\\
865 & tir snašta. & you (pl.) receive. & statement\\
866 & drunun naštem! & trade dread; the weight of meaninglessness! & command\\
867 & zan'kai kavgitun zlenas, šaen zan'kai krekun venša. & we (bonded) build air, but we (bonded) to cross; to jump a threshold anger. & contrast\\
868 & ti rlasta eh? & Does you lose? & question\\
869 & za korra. & I to work; fix. & statement\\
870 & zan'kai kešiun talra. & we (bonded) to rest; pause wire; line; cable. & statement\\
871 & ti venša. & you to cross; to jump a threshold. & statement\\
872 & za zulkaíventa nai. & I do not crossing the void with the chosen. & negation\\
873 & se molra nai. & they do not carry. & negation\\
874 & zan kairenun vomta, šaen zan lokmarethun molša. & we wait bond-restorer; group glue-person, but we carry deep. & contrast\\
875 & za dunša nai. & I do not to eat. & negation\\
876 & se kaiumun rukra ser. & they hurt the chosen (seen/verified). & log\\
877 & ti vakša. & you to wake. & statement\\
878 & ti lirša nai. & you do not to give. & negation\\
879 & se zradra eh? & Does they remember? & question\\
880 & tir za'kai lumlavša. & you (pl.) refuse with the chosen. & bond\\
881 & se heš, kur tir sáliarun ruk. & they to listen, because you (pl.) hurt stubborn defiant hope-light. & cause\\
882 & ti zan'kai tat. & you choose with the chosen. & bond\\
883 & za zulserun šinta eh? & Does I to drink; drink/water zulser? & question\\
884 & zan'kai šinša. & we (bonded) to drink; drink/water. & statement\\
885 & zan'kai šurša eh? & Does we (bonded) to survive; endure? & question\\
886 & za kraßa, kur zan'kai tovarniritun gaktadša. & I rise, because we (bonded) fail tovar-nirit. & cause\\
887 & se naštera eh? & Does they trade? & question\\
888 & ti mórun lonta eh? & Does you know laughter in the dark? & question\\
889 & zan kaiiun šurta, šaen zan rogresarun ronzid. & we to survive; endure chosen companion; bonded company, but we believe calm. & contrast\\
890 & zan vrunethun šur tov. & we to survive; endure fear (inferred). & log\\
891 & ti grenša, kur zan rlasethun tovaršurša. & you fight, because we endure by continuing the story-thread lose. & cause\\
892 & zan kraš. & we rise. & statement\\
893 & se šinra, kur tir šošumun snašra. & they to drink; drink/water, because you (pl.) receive heal. & cause\\
894 & vaskkaiun serm! & to see; verify ship-kin; crew-as-family! & command\\
895 & za vul eh? & Does I to speak; say? & question\\
896 & ronzidm! & believe! & command\\
897 & tir donša, kur zan'kai hílenkordun vrenša. & you (pl.) break, because we (bonded) to lie; a lie hílen-kil. & cause\\
898 & ti lesša nai. & you do not hate. & negation\\
899 & zan treenun lirša, šaen zan kaizulun lumlavta. & we to give fly, but we refuse kai-zul. & contrast\\
900 & ti mil. & you walk. & statement\\
901 & zan'kai druniun vakrekra. & we (bonded) promise dread. & statement\\
902 & molm! & carry! & command\\
903 & ti mit eh? & Does you touch? & question\\
904 & kathzešun draim! & to drift; wander kath-zeš! & command\\
905 & zan za'kai milša. & we walk with the chosen. & bond\\
906 & zan'kai serra eh? & Does we (bonded) to see; verify? & question\\
907 & zan'kai relnarun zlusša, šaen zan'kai vaskun mista. & we (bonded) buy fold-gate/place, but we (bonded) open ship; vessel. & contrast\\
908 & se vulša. & they to speak; say. & statement\\
909 & ti ti'kai snon. & you arrive with the chosen. & bond\\
910 & se zulun molša vem. & they carry indifferent void; cosmic uncaring (reported). & log\\
911 & za dunra nai. & I do not to eat. & negation\\
912 & hílenrelnkaiun vrenm! & to lie; a lie hílenrelnkai! & command\\
913 & zan mórsizruvun korta eh? & Does we to work; fix mór-sizruv? & question\\
914 & za somarun lumlav. & I refuse to sleep. & statement\\
915 & za zaknimenun lalgemra. & I leave close. & statement\\
916 & zan'kai susdisra. & we (bonded) feel. & statement\\
917 & za somša. & I to sleep. & statement\\
918 & zan rlas nai. & we do not lose. & negation\\
919 & ti somta eh? & Does you to sleep? & question\\
920 & tir ven. & you (pl.) to cross; to jump a threshold. & statement\\
921 & se zrosenun vak eh? & Does they to wake false? & question\\
922 & zan'kai sok nai. & we (bonded) do not to take. & negation\\
923 & za heš, kur zan'kai liriun vak. & I to listen, because we (bonded) to wake to give. & cause\\
924 & tovarrelnarun korm! & to work; fix tovarrelnar! & command\\
925 & zan serša nai. & we do not to see; verify. & negation\\
926 & za vaskdarleun zaknimra, šaen za drisun zulkaíventa. & I close vask-darle, but I crossing the void with the chosen home. & contrast\\
927 & za kraßa, kur se kathvakrekun somta. & I rise, because they to sleep kath-vakrek. & cause\\
928 & zan nérvaskkaiun trera. & we fly nér-vaskkai. & statement\\
929 & gusumun venm! & to cross; to jump a threshold far! & command\\
930 & mórdrisun milm! & walk mór-dris! & command\\
931 & ti vaskrorun dunra eh? & Does you to eat vask-ror? & question\\
932 & zan'kai krum eh? & Does we (bonded) think? & question\\
933 & zan'kai ruk nai. & we (bonded) do not hurt. & negation\\
934 & ti se'kai grenra. & you fight with the chosen. & bond\\
935 & kaiumun donm! & break the chosen! & command\\
936 & nérklinun ronzidm! & believe nér-klin! & command\\
937 & se zimša nai. & they do not pay. & negation\\
938 & zan kailedzrasun škonša, šaen zan drunun vretra. & we hide kai-ledzras, but we push dread; the weight of meaninglessness. & contrast\\
939 & zan'kai seka lonta. & we (bonded) know with (neutral). & social\\
940 & ti gliekra nai. & you do not fall. & negation\\
941 & za zlerethun darle, šaen za vaskšvulun lesta. & I live love, but I hate vask-švul. & contrast\\
942 & zan'kai kathvakrekun dunta tov. & we (bonded) to eat kath-vakrek (inferred). & log\\
943 & se našun vretša. & they push hand; grip. & statement\\
944 & tir vakrek eh? & Does you (pl.) promise? & question\\
945 & luvaenun korm! & to work; fix name! & command\\
946 & zan'kai vnunvutethun naštera eh? & Does we (bonded) trade captain? & question\\
947 & zan'kai atsnukta. & we (bonded) come. & statement\\
948 & ti gaktadarun vrenta eh? & Does you to lie; a lie fail? & question\\
949 & tir trelenun dolimra, šaen tir golenun škunta. & you (pl.) turn tired, but you (pl.) watch door. & contrast\\
950 & ti senun mitša. & you touch skin; surface; hull. & statement\\
951 & zan'kai hílenun zrad vem. & we (bonded) remember homesickmemory (reported). & log\\
952 & zan sivethun gorirra eh? & Does we learn lock; seal? & question\\
953 & za vilkal, kur zan kaiumun trera. & I forgive, because we fly the chosen. & cause\\
954 & ti snonra eh? & Does you arrive? & question\\
955 & se hílenzrismidun ronzidša tov. & they believe hílen-zrismid (inferred). & log\\
956 & ti rervaumun tovaršurša. & you endure by continuing the story-thread metal. & statement\\
957 & se mis nai. & they do not open. & negation\\
958 & zan zan'kai lonra. & we know with the chosen. & bond\\
959 & torun talm! & to rest; pause light! & command\\
960 & ti vilkal nai. & you do not forgive. & negation\\
961 & zan'kai vaviun dun, šaen zan'kai gimkredun dolim. & we (bonded) to eat day, but we (bonded) turn doctor. & contrast\\
962 & za seka našteta. & I trade with (neutral). & social\\
963 & za škinenun draira. & I to drift; wander dangerous. & statement\\
964 & se tarnun vradgoša eh? & Does they start money; credits; tokens? & question\\
965 & tir vakša. & you (pl.) to wake. & statement\\
966 & tir rorumun tovaršurša. & you (pl.) endure by continuing the story-thread station. & statement\\
967 & se klitrinarun darta tov. & they to want; desire port (inferred). & log\\
968 & ti grimun zlerra. & you love room. & statement\\
969 & ti susdisarun gliekra. & you fall feel. & statement\\
970 & zan'kai vren, kur tir tovarenun tikša. & we (bonded) to lie; a lie, because you (pl.) teach story-thread; chosen route. & cause\\
971 & za škun nai. & I do not watch. & negation\\
972 & zan'kai zanka škon. & we (bonded) hide with (neutral). & social\\
973 & za rlasša nai. & I do not lose. & negation\\
974 & zan'kai zaka zulkaívenša. & we (bonded) crossing the void with the chosen with (neutral). & social\\
975 & tir šinta eh? & Does you (pl.) to drink; drink/water? & question\\
976 & tir škirta eh? & Does you (pl.) run? & question\\
977 & za zaka venša. & I to cross; to jump a threshold with (neutral). & social\\
978 & se somethun dar eh? & Does they to want; desire to sleep? & question\\
979 & snustatiun hešm! & to listen medicine! & command\\
980 & za škon eh? & Does I hide? & question\\
981 & zan milta nai. & we do not walk. & negation\\
982 & ti vakrekta eh? & Does you promise? & question\\
983 & tir grenta nai. & you (pl.) do not fight. & negation\\
984 & ti lumlav. & you refuse. & statement\\
985 & se zan'kai zlenasša. & they build with the chosen. & bond\\
986 & tir škonra nai. & you (pl.) do not hide. & negation\\
987 & tir kailedzrasun mista. & you (pl.) open kai-ledzras. & statement\\
988 & se ven. & they to cross; to jump a threshold. & statement\\
989 & za kridenun mis, šaen za vekun atsnuk. & I open go, but I come dark. & contrast\\
990 & za seka tovaršurra. & I endure by continuing the story-thread with (neutral). & social\\
991 & zan se'kai venra. & we to cross; to jump a threshold with the chosen. & bond\\
992 & relnlokun talm! & to rest; pause reln-lok! & command\\
993 & za kathzrosun sizruvša, šaen za tovarrukun gliekša. & I forget kath-zros, but I fall tovar-ruk. & contrast\\
994 & za dunša, kur se sureniun gliekra. & I to eat, because they fall caretaker-survivor; the one who gets you through. & cause\\
995 & ti darša nai. & you do not to want; desire. & negation\\
996 & lomakm! & find! & command\\
997 & za za'kai sairta. & I to tell truth; align speech with reality with the chosen. & bond\\
998 & zan'kai kevralenun zeßa, šaen zan'kai vakarun lesra. & we (bonded) help thirsty, but we (bonded) hate to wake. & contrast\\
999 & se treta. & they fly. & statement\\
1000 & zradm! & remember! & command\\

\end{longtable}
\end{document}
