\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{xparse}

% ---------- Global script feel ----------
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\raggedright

% ---------- Counters ----------
\newcounter{gnpage}
\newcounter{gnpanel}[gnpage]
\renewcommand{\thegnpage}{\arabic{gnpage}}
\renewcommand{\thegnpanel}{\arabic{gnpanel}}

% ---------- Tunable indents ----------
\newlength{\GNDialogueLeft}
\newlength{\GNDialogueRight}
\setlength{\GNDialogueLeft}{0.22\textwidth}
\setlength{\GNDialogueRight}{0.22\textwidth}

% ---------- Title block ----------
\NewDocumentCommand{\GNTitle}{ m m O{\today} }{%
  \thispagestyle{empty}
  \vspace*{0.18\textheight}
  {\Large\bfseries #1\par}
  \vspace{0.5em}
  {\large #2\par}
  \vspace{1em}
  {\itshape #3\par}
  \vfill
  \newpage
}

% ---------- Page ----------
% Usage: \begin{GNPage}[optional page subtitle] ... \end{GNPage}
\NewDocumentEnvironment{GNPage}{ O{} }{%
  \refstepcounter{gnpage}
  \setcounter{gnpanel}{0}
  \vspace{1em}
  {\bfseries PAGE \thegnpage%
    \IfValueT{#1}{ - #1}%
  }\par
  \vspace{0.5em}
}{%
  \vspace{1em}
}

% ---------- Scene / Environment ----------
\NewDocumentCommand{\Env}{ +m }{%
  {\bfseries\MakeUppercase{#1}}\par
}

% ---------- Panel ----------
% Starred version doesn't increment (useful for splash / non-numbered beats)
% Usage: \Panel[optional label]{description}
\NewDocumentCommand{\Panel}{ s O{} +m }{%
  \par\vspace{0.5em}
  \IfBooleanTF{#1}{%
    {\bfseries PANEL%
      \IfValueT{#2}{\ \textnormal{(\MakeUppercase{#2})}}%
    }\par
  }{%
    \refstepcounter{gnpanel}
    {\bfseries PANEL \thegnpanel%
      \IfValueT{#2}{\ \textnormal{(\MakeUppercase{#2})}}%
    }\par
  }%
  #3\par
}

% ---------- Action ----------
\NewDocumentCommand{\Action}{ +m }{%
  #1\par
}

% ---------- Caption / Narration ----------
\NewDocumentCommand{\Caption}{ +m }{%
  {\itshape CAPTION: #1}\par
}
\NewDocumentCommand{\Narration}{ +m }{%
  {\itshape NARRATION: #1}\par
}

% ---------- SFX ----------
\NewDocumentCommand{\SFX}{ +m }{%
  {\bfseries SFX: \MakeUppercase{#1}}\par
}

% ---------- Transition ----------
\NewDocumentCommand{\Transition}{ +m }{%
  \hfill{\bfseries \MakeUppercase{#1}}\par
}

% ---------- Note to self (won't look like a panel element) ----------
\NewDocumentCommand{\Note}{ +m }{%
  {\small\itshape [NOTE: #1]}\par
}

% ---------- Dialogue ----------
% Usage: \Dialogue[optional parenthetical]{CHARACTER}{line(s)}
\NewDocumentCommand{\Dialogue}{ O{} m +m }{%
  {\par\noindent\centering\textsc{#2}\par}
  \begin{list}{}{%
    \setlength{\leftmargin}{\GNDialogueLeft}%
    \setlength{\rightmargin}{\GNDialogueRight}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parsep}{0pt}%
    \setlength{\topsep}{2pt}%
  }%
    \item
      \IfValueT{#1}{\emph{(#1)}\par}
      #3
  \end{list}%
}

% Optional quick beats
\NewDocumentCommand{\Beat}{}{%
  {\itshape (beat)}\par
}

\begin{document}

\GNTitle{KA\'{I}VEN}{noumena ashitaka hundim\"{a}gi-mei}[Draft - \today]

\begin{GNPage}[The End]
\Env{EXT. SPACE CLOSE TO BLACKHOLE}

\Panel[WIDE, ESTABLISHING]{\textbf{Exterior}. The ship hangs like a wounded insect in the dark. Plates peeled back. Scars of plasma burns. One engine dead, another coughing blue.\\
	Ahead: a \textbf{gargantuan black hole}, not just darkness but a presence an eye without eyelids. The accretion disk smears light into a bright, violent halo. Distant debris arcs inward like prayer beads being pulled from a snapped string.}

\Caption{We have run so long that even the stars feel tired of watching us.}

\Caption{The ship is ruined, its bones cracked, its skin torn, and there is no kindly harbor in the void.}

\Panel[INTERIOR, WIDE]{\textbf{Bridge.} The crew of six. Exhaustion is everywhere: slumped shoulders, trembling hands, bruises under eyes, dried blood at someone's hairline. Consoles flicker with dying light. A warning klaxon is either silent now or too constant to notice.\\

Only the protagonist is upright, facing forward, hands on the console like he's holding the world in place.}

\Caption{They trusted me. That is the sharpest thing in the room.}

\Caption{Not the alarms. Not the heat bleeding from the vents. Not even the black mouth waiting outside.}

\Panel[CLOSE-UP: DAMAGE / DETAILS]{A cracked screen shows: \textbf{HULL INTEGRITY: 12\%}.
A manual patch welded over a breach. A leaking cable sparking. A tiny photo or token taped beside a gauge, something human and stubborn.}

\Caption{How many times can a man make the same choice and still call it bravery?}

\Caption{How many times can he lead others into fire and name it \textbf{the only way}?}

\Panel[CLOSE-UP: CREW FACES, ONE BY ONE]{
Quick cuts:

\begin{itemize}
  \item One crew member asleep upright, jaw clenched like resisting a nightmare.
  \item Another quietly crying without sound.
  \item Someone's hand hovering over an eject lever, then falling away.
  \item Someone's staring at the black hole on the forward display as if looking at an old god.
\end{itemize}
}

\Caption{I wanted a simpler life once. I remember that version of me like a half-forgotten song.}

\Caption{A house. A door that locks. Mornings that mean bread and not battle.}

\Caption{But I stepped onto this road, again, and again, and each time I told myself it was the last.}

\Panel[INTERIOR, LOW ANGLE ON PROTAGONIST / FORWARD WINDOW]{The forward window dominates. The black hole swells larger now, a slow certainty. Light bends wrong. The ship is being \textit{drawn}, not by engines, but by fate wearing physics as a mask. The protagonist's face is lit by the warped glow-half saint, half criminal.}

\Caption{Why do I keep doing this?}

\Caption{Why does it feel... not like living... but like repeating?}

\Caption{As if some unseen hand has set the pieces upon the board before I ever touch them.}

\Caption{As if the world has grooves worn into it, and my feet find them no matter how I fight.}

\Panel[FULL-BLEED, COSMIC / SURREAL]{The ship tiny against the black hole's majesty. The accretion disk looks like a burning crown. The stars behind smear into thin lines, as though reality itself is being combed toward the dark. In the chaos, there is a strange calm like a theater right before the curtain falls.}

\Caption{I tell myself I am choosing.}

\Caption{I tell myself there is meaning in the struggle, and honor in the scars.}

\Caption{But some nights, like this one, I swear I can hear the machinery behind the sky.}

\Caption{A faint clicking. A silent counting.}

\Caption{And the cruelest thought of all is this:}

\Caption{If this ends the way it always ends... then what, in the name of all the lights that ever were, was I meant to learn?}

\end{GNPage}

\end{document}
