\documentclass[10pt,a4paper]{article}

\usepackage[a4paper,margin=0.55in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{hyperref}

\hypersetup{hidelinks}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2pt}
\setlist[enumerate]{leftmargin=*, itemsep=1.5pt, topsep=2pt}

\newcommand{\ordic}[1]{\textbf{#1}}
\newcommand{\gloss}[1]{\hfill{\small\textsf{#1}}}

\begin{document}

{\LARGE \textbf{Low Ordic: Common Phrases}}\\
{\small Glossed corpus. Low Ordic register: spoken, emotional, irregular, human.}

\vspace{6pt}\hrule\vspace{8pt}

\begin{multicols}{2}

\begin{enumerate}

% --- Greetings / presence
\item \ordic{nuu.} \gloss{Now / here.}
\item \ordic{ya nuu.} \gloss{I’m here.}
\item \ordic{tu nuu?} \gloss{You here?}
\item \ordic{we nuu.} \gloss{We’re here.}
\item \ordic{nuu we.} \gloss{Here together.}
\item \ordic{an nuu.} \gloss{Glad you’re here.}
\item \ordic{nuu an.} \gloss{It feels good now.}
\item \ordic{nuu shae l.} \gloss{It’s calm now.}
\item \ordic{nuu qe.} \gloss{Something feels uncertain.}
\item \ordic{nuu jo.} \gloss{This moment is joyful.}

% --- Check-ins / care
\item \ordic{tu es?} \gloss{Are you okay?}
\item \ordic{ya es.} \gloss{I’m okay / I exist.}
\item \ordic{ya es-na.} \gloss{I’m not okay.}
\item \ordic{tu mu?} \gloss{Do you remember?}
\item \ordic{ya mu.} \gloss{I remember.}
\item \ordic{ya mu-na.} \gloss{I forgot.}
\item \ordic{tu shae l?} \gloss{Are you calm?}
\item \ordic{ya shae l.} \gloss{I’m calm.}
\item \ordic{ya shae l-na.} \gloss{I’m not calm.}
\item \ordic{an tu.} \gloss{I care about you.}

% --- Feelings (imperfect, leaky)
\item \ordic{ya an.} \gloss{I love / I care.}
\item \ordic{an we.} \gloss{We care about each other.}
\item \ordic{ya jo.} \gloss{I’m happy.}
\item \ordic{ya sae.} \gloss{I’m sad.}
\item \ordic{ya ir.} \gloss{I’m angry.}
\item \ordic{ya ve.} \gloss{I’m scared / in awe.}
\item \ordic{ya qe.} \gloss{I’m unsure.}
\item \ordic{ya ru.} \gloss{I want / I desire.}
\item \ordic{ya ny.} \gloss{I’m ashamed.}
\item \ordic{ya shae l.} \gloss{I feel calm.}

% --- Simple actions
\item \ordic{ya lin.} \gloss{I’m going.}
\item \ordic{ya ret.} \gloss{I’m back / I return.}
\item \ordic{ya dur.} \gloss{I’m staying.}
\item \ordic{ya pun.} \gloss{I did it / it’s done.}
\item \ordic{ya vek.} \gloss{I changed.}
\item \ordic{ya ki.} \gloss{I’m doing it.}
\item \ordic{ya se.} \gloss{I allow it / okay.}
\item \ordic{ya vu.} \gloss{I resist / no.}
\item \ordic{ya so.} \gloss{I said it.}
\item \ordic{ya vi.} \gloss{I see.}

% --- Requests / softness
\item \ordic{tu se?} \gloss{Is that okay with you?}
\item \ordic{tu lin?} \gloss{Will you come?}
\item \ordic{tu dur?} \gloss{Will you stay?}
\item \ordic{tu ret?} \gloss{Will you come back?}
\item \ordic{tu mu-te?} \gloss{Do you remember this clearly?}
\item \ordic{tu vi-te?} \gloss{Do you see the truth of it?}
\item \ordic{fre ya?} \gloss{May I?}
\item \ordic{se we?} \gloss{Can we do this?}
\item \ordic{na tu.} \gloss{Please don’t.}
\item \ordic{gra tu.} \gloss{Please / show mercy.}

% --- Apology / repair
\item \ordic{ya qe-na.} \gloss{I was wrong.}
\item \ordic{ya so-na.} \gloss{I shouldn’t have said that.}
\item \ordic{ya ki-na.} \gloss{I shouldn’t have done that.}
\item \ordic{gra ya.} \gloss{Please forgive me.}
\item \ordic{gra we.} \gloss{Let’s forgive each other.}
\item \ordic{ya ret gra.} \gloss{I want to make it right.}
\item \ordic{sha xo-na.} \gloss{Let’s not break the bond.}
\item \ordic{xo ret.} \gloss{Let’s reconnect.}
\item \ordic{an xo we.} \gloss{Love binds us.}
\item \ordic{we dur.} \gloss{We’ll stay.}

% --- Conflict (softened)
\item \ordic{ya ir-na.} \gloss{I’m trying not to be angry.}
\item \ordic{ya qe ir.} \gloss{I’m angry and unsure.}
\item \ordic{tu ir.} \gloss{You’re angry.}
\item \ordic{tu ir-na.} \gloss{You’re not angry (anymore).}
\item \ordic{na sha.} \gloss{Don’t break this.}
\item \ordic{na dar.} \gloss{Don’t force it.}
\item \ordic{na jur.} \gloss{Don’t make this a rule.}
\item \ordic{na ord.} \gloss{Let’s not over-order this.}
\item \ordic{qe we.} \gloss{We’re not sure.}
\item \ordic{qe tu.} \gloss{I’m not sure about you.}

% --- Hope / future
\item \ordic{pos we.} \gloss{We will.}
\item \ordic{pos an.} \gloss{Things will feel better.}
\item \ordic{pos shae l.} \gloss{It will calm down.}
\item \ordic{pos ret.} \gloss{We’ll come back.}
\item \ordic{pos xo.} \gloss{We’ll stay connected.}
\item \ordic{pos mu.} \gloss{We’ll remember.}
\item \ordic{pos ord-na.} \gloss{It won’t be perfect.}
\item \ordic{pos qe.} \gloss{We’ll figure it out.}
\item \ordic{pos jo.} \gloss{There will be joy.}
\item \ordic{pos an we.} \gloss{We’ll still care.}

% --- Small closings
\item \ordic{nuu dur.} \gloss{Let’s stay here for now.}
\item \ordic{nuu ret-na.} \gloss{Don’t go yet.}
\item \ordic{nuu shae l we.} \gloss{Let’s be calm together.}
\item \ordic{we nuu shae l.} \gloss{We’re okay right now.}
\item \ordic{ya nuu shae l.} \gloss{I’m okay right now.}
\item \ordic{tu nuu shae l.} \gloss{You’re okay right now.}
\item \ordic{an nuu we.} \gloss{I’m glad we’re here.}
\item \ordic{we an.} \gloss{We care.}
\item \ordic{dur\ldots} \gloss{Stay\ldots (soft pause).}
\item \ordic{nuu\ldots} \gloss{Now\ldots (soft pause).}

% --- Everyday needs
\item \ordic{ya ru dur.} \gloss{I still want this.}
\item \ordic{ya ru-na.} \gloss{I don’t want it.}
\item \ordic{ru tu?} \gloss{Do you want it?}
\item \ordic{ru we.} \gloss{We want this.}
\item \ordic{ya dur-na.} \gloss{I can’t stay.}
\item \ordic{ya lin nuu.} \gloss{I’m leaving now.}
\item \ordic{ya ret pos.} \gloss{I’ll come back later.}
\item \ordic{ya pun-na.} \gloss{I didn’t finish.}
\item \ordic{ya pun jo.} \gloss{I did it—yay.}
\item \ordic{ya pun ir.} \gloss{I did it angrily.}

% --- Tiredness / overwhelm
\item \ordic{ya dur far.} \gloss{I’m exhausted.}
\item \ordic{ya mu far.} \gloss{My memory’s gone.}
\item \ordic{ya qe dur.} \gloss{I don’t know how long I can keep going.}
\item \ordic{ya shae l-na.} \gloss{I’m not calm anymore.}
\item \ordic{nuu far.} \gloss{This is too much right now.}
\item \ordic{ord far.} \gloss{This is too complicated.}
\item \ordic{na ord, nuu.} \gloss{Not rules—now.}
\item \ordic{qe qe.} \gloss{I’m really unsure.}
\item \ordic{ya qe qe.} \gloss{I’m doubting myself.}
\item \ordic{we qe dur.} \gloss{We’re struggling.}

% --- Affection / closeness
\item \ordic{ya an tu.} \gloss{I care about you.}
\item \ordic{an ya tu.} \gloss{I really care about you.}
\item \ordic{an nuu tu.} \gloss{I’m glad you’re here now.}
\item \ordic{an dur we.} \gloss{We still care.}
\item \ordic{xo we dur.} \gloss{Our bond still holds.}
\item \ordic{xo tu ya.} \gloss{I’m connected to you.}
\item \ordic{xo far-na.} \gloss{We’re not drifting apart.}
\item \ordic{an ret.} \gloss{The feeling came back.}
\item \ordic{an jo.} \gloss{It feels good.}
\item \ordic{an shae l.} \gloss{It feels safe.}

% --- Small jokes / lightness
\item \ordic{nuu jo, ord-na.} \gloss{This is funny, not orderly.}
\item \ordic{qe jo.} \gloss{Is this a joke?}
\item \ordic{jo jo.} \gloss{Haha.}
\item \ordic{ya so jo.} \gloss{I was joking.}
\item \ordic{so ir-na.} \gloss{I didn’t mean it angrily.}
\item \ordic{ord jo-na.} \gloss{Rules aren’t funny.}
\item \ordic{nuu jo we.} \gloss{We’re laughing right now.}
\item \ordic{ya pun jo-na.} \gloss{That didn’t go as planned.}
\item \ordic{ret jo.} \gloss{That’s funny again.}
\item \ordic{jo dur.} \gloss{The laughter continues.}

% --- Disagreement (gentle)
\item \ordic{ya qe tu.} \gloss{I’m not sure about that.}
\item \ordic{ya qe-na tu.} \gloss{I don’t think so.}
\item \ordic{ya vi dif.} \gloss{I see it differently.}
\item \ordic{tu vi-na.} \gloss{You don’t see it.}
\item \ordic{qe we ord.} \gloss{We don’t agree on the rules.}
\item \ordic{na xo-na.} \gloss{Let’s not break apart.}
\item \ordic{gra, qe we.} \gloss{Please—let’s be gentle.}
\item \ordic{nuu shae l, tu.} \gloss{Stay calm, okay?}
\item \ordic{ya ir, an-na.} \gloss{I’m angry, but I still care.}
\item \ordic{ir ret-na.} \gloss{Let’s not go back to anger.}

% --- Trust / reassurance
\item \ordic{ya vi tu.} \gloss{I see you.}
\item \ordic{ya lu tu.} \gloss{I understand you.}
\item \ordic{tu vi ya?} \gloss{Do you see me?}
\item \ordic{we vi we.} \gloss{We see each other.}
\item \ordic{ya mu tu.} \gloss{I remember you.}
\item \ordic{mu xo.} \gloss{The bond is remembered.}
\item \ordic{xo te.} \gloss{The bond is real.}
\item \ordic{xo dur-na?} \gloss{Is this still holding?}
\item \ordic{xo dur.} \gloss{Yes, it is.}
\item \ordic{gra xo.} \gloss{Please don’t let this break.}

% --- Waiting / uncertainty
\item \ordic{dur qe.} \gloss{Let’s wait and see.}
\item \ordic{nuu dur qe.} \gloss{For now, we wait.}
\item \ordic{pos qe.} \gloss{We’ll see later.}
\item \ordic{ret qe.} \gloss{The doubt came back.}
\item \ordic{qe mu.} \gloss{I’m not sure I remember right.}
\item \ordic{qe vi.} \gloss{I’m not sure what I see.}
\item \ordic{qe so.} \gloss{I don’t know what to say.}
\item \ordic{qe ki.} \gloss{I don’t know what to do.}
\item \ordic{qe nuu.} \gloss{Right now, I don’t know.}
\item \ordic{qe\ldots} \gloss{I’m thinking\ldots}

% --- Gentle endings
\item \ordic{nuu dur we.} \gloss{Let’s stay like this for now.}
\item \ordic{ya dur tu.} \gloss{I’ll stay with you.}
\item \ordic{tu dur ya?} \gloss{Will you stay with me?}
\item \ordic{we dur an.} \gloss{We’ll stay caring.}
\item \ordic{nuu shae l.} \gloss{It’s calm right now.}
\item \ordic{shae l ret.} \gloss{The calm is coming back.}
\item \ordic{an dur.} \gloss{The care remains.}
\item \ordic{xo dur.} \gloss{The bond remains.}
\item \ordic{nuu\ldots} \gloss{Now\ldots}
\item \ordic{dur\ldots} \gloss{Stay\ldots}

% --- Frustration / release
\item \ordic{ya ir dur.} \gloss{I’m still angry.}
\item \ordic{ir nuu.} \gloss{This hurts right now.}
\item \ordic{ir jo-na.} \gloss{This isn’t funny.}
\item \ordic{ir ord-na.} \gloss{This makes no sense.}
\item \ordic{ya ir, qe.} \gloss{I’m angry and confused.}
\item \ordic{ir ret.} \gloss{The anger came back.}
\item \ordic{ir far-na.} \gloss{I’m not over it yet.}
\item \ordic{ir pun.} \gloss{I snapped.}
\item \ordic{ir so.} \gloss{I said something in anger.}
\item \ordic{ir so-na.} \gloss{I shouldn’t have said that.}

% --- Vulnerability
\item \ordic{ya ve nuu.} \gloss{I’m scared right now.}
\item \ordic{ve qe.} \gloss{This is frightening.}
\item \ordic{ya ve-na tu.} \gloss{I’m not scared of you.}
\item \ordic{ve dur.} \gloss{The fear lingers.}
\item \ordic{ve far-na.} \gloss{The fear hasn’t gone away.}
\item \ordic{ya ny.} \gloss{I’m ashamed.}
\item \ordic{ny qe.} \gloss{This feels wrong.}
\item \ordic{ny far.} \gloss{I want to hide.}
\item \ordic{ya ny-na tu.} \gloss{I’m not ashamed with you.}
\item \ordic{ny shae l.} \gloss{The shame eases.}

% --- Intimacy / trust
\item \ordic{ya vi tu, nuu.} \gloss{I see you, right now.}
\item \ordic{tu vi ya, nuu?} \gloss{Do you see me right now?}
\item \ordic{ya lu tu.} \gloss{I understand you.}
\item \ordic{ya lu-na tu.} \gloss{I don’t understand you yet.}
\item \ordic{lu we.} \gloss{We understand each other.}
\item \ordic{lu far-na.} \gloss{We’re not lost.}
\item \ordic{xo nuu.} \gloss{We’re connected right now.}
\item \ordic{xo nuu dur.} \gloss{The connection holds.}
\item \ordic{xo qe-na.} \gloss{The bond feels steady.}
\item \ordic{xo qe.} \gloss{The bond feels uncertain.}

% --- Confession
\item \ordic{ya so te.} \gloss{I need to say something true.}
\item \ordic{ya so qe.} \gloss{I don’t know how to say this.}
\item \ordic{so qe-na.} \gloss{That didn’t come out right.}
\item \ordic{so mu.} \gloss{I’m saying this from memory.}
\item \ordic{so ir.} \gloss{I spoke from anger.}
\item \ordic{so an.} \gloss{I spoke from care.}
\item \ordic{so ve.} \gloss{I spoke from fear.}
\item \ordic{so shae l.} \gloss{I spoke calmly.}
\item \ordic{so far-na.} \gloss{I’m not finished saying this.}
\item \ordic{so\ldots} \gloss{I’m trying to speak\ldots}

% --- Boundaries
\item \ordic{na lin.} \gloss{Don’t come closer.}
\item \ordic{lin qe.} \gloss{Come closer, carefully.}
\item \ordic{dur far.} \gloss{I need space.}
\item \ordic{dur tu.} \gloss{Stay with me.}
\item \ordic{na so nuu.} \gloss{Not right now.}
\item \ordic{na qe tu.} \gloss{Please don’t ask that.}
\item \ordic{ya se-na.} \gloss{I can’t allow this.}
\item \ordic{ya se.} \gloss{Okay, I allow it.}
\item \ordic{ya vu-na.} \gloss{I won’t resist.}
\item \ordic{ya vu.} \gloss{I can’t do this.}

% --- Repair / forgiveness (human scale)
\item \ordic{gra nuu.} \gloss{Please, now.}
\item \ordic{gra ya tu.} \gloss{Please forgive me.}
\item \ordic{gra tu ya.} \gloss{I forgive you.}
\item \ordic{gra we nuu.} \gloss{Let’s forgive each other now.}
\item \ordic{gra ret.} \gloss{Forgiveness came back.}
\item \ordic{gra dur.} \gloss{Forgiveness holds.}
\item \ordic{gra qe.} \gloss{I’m not ready to forgive yet.}
\item \ordic{gra far-na.} \gloss{I don’t want to lose forgiveness.}
\item \ordic{gra shae l.} \gloss{Forgiveness feels calm.}
\item \ordic{gra xo.} \gloss{Let the bond heal.}

% --- Shared silence
\item \ordic{nuu shae l.} \gloss{It’s calm right now.}
\item \ordic{nuu so-na.} \gloss{No words right now.}
\item \ordic{dur so-na.} \gloss{Let’s not speak.}
\item \ordic{dur nuu.} \gloss{Let’s stay in this moment.}
\item \ordic{we nuu.} \gloss{We’re here together.}
\item \ordic{xo nuu shae l.} \gloss{The bond is calm now.}
\item \ordic{an nuu shae l.} \gloss{This feels safe now.}
\item \ordic{nuu\ldots} \gloss{Now\ldots}
\item \ordic{shae l\ldots} \gloss{Calm\ldots}
\item \ordic{dur\ldots} \gloss{Stay\ldots}

% --- Exhaustion / endurance
\item \ordic{ya dur far-na.} \gloss{I’m tired, but not done.}
\item \ordic{ya dur qe.} \gloss{I’m not sure I can keep going.}
\item \ordic{dur far, nuu.} \gloss{This is heavy right now.}
\item \ordic{dur jo-na.} \gloss{I can’t laugh at this.}
\item \ordic{ya ret dur.} \gloss{I came back anyway.}
\item \ordic{ya dur pos.} \gloss{I’ll stay for now.}
\item \ordic{we dur far.} \gloss{We’re all tired.}
\item \ordic{dur far we.} \gloss{This is hard for us.}
\item \ordic{dur far-na we.} \gloss{We’re hurting, but still here.}
\item \ordic{dur qe we.} \gloss{We don’t know how long we can hold on.}

% --- Humor as survival
\item \ordic{jo qe.} \gloss{That’s awkwardly funny.}
\item \ordic{jo ir-na.} \gloss{I’m laughing instead of being mad.}
\item \ordic{jo nuu.} \gloss{This is funny right now.}
\item \ordic{jo dur-na.} \gloss{The joke didn’t land.}
\item \ordic{ya so jo-na.} \gloss{Okay, that was a bad joke.}
\item \ordic{jo ret.} \gloss{It’s funny again.}
\item \ordic{jo far-na.} \gloss{It’s not that bad.}
\item \ordic{jo we.} \gloss{We’re laughing together.}
\item \ordic{jo xo.} \gloss{Laughter keeps us connected.}
\item \ordic{jo shae l.} \gloss{Laughter calms things down.}

% --- Stubborn hope
\item \ordic{pos dur.} \gloss{It’ll keep going.}
\item \ordic{pos jo.} \gloss{There’ll be laughter.}
\item \ordic{pos shae l.} \gloss{It’ll calm down.}
\item \ordic{pos ret we.} \gloss{We’ll come back together.}
\item \ordic{pos xo.} \gloss{The bond will hold.}
\item \ordic{pos an.} \gloss{The care will remain.}
\item \ordic{pos qe-na.} \gloss{We’ll figure it out.}
\item \ordic{pos mu.} \gloss{We’ll remember.}
\item \ordic{pos vi.} \gloss{We’ll see.}
\item \ordic{pos dur we.} \gloss{We’ll stay.}

% --- Love under strain
\item \ordic{ya an dur.} \gloss{I still care.}
\item \ordic{an far-na.} \gloss{The care hasn’t faded.}
\item \ordic{an dur we.} \gloss{We still care.}
\item \ordic{an qe.} \gloss{I’m unsure how I feel.}
\item \ordic{an qe-na.} \gloss{The feeling is real.}
\item \ordic{an ir-na.} \gloss{I care without anger.}
\item \ordic{an shae l.} \gloss{The care feels safe.}
\item \ordic{an dur-na.} \gloss{I’m scared it won’t last.}
\item \ordic{an ret.} \gloss{The feeling came back.}
\item \ordic{an xo dur.} \gloss{The care keeps us bound.}

% --- Presence / choosing to stay
\item \ordic{ya nuu dur.} \gloss{I’m staying here.}
\item \ordic{ya dur tu.} \gloss{I’m staying with you.}
\item \ordic{tu dur ya?} \gloss{Will you stay with me?}
\item \ordic{we dur nuu.} \gloss{We’re staying here.}
\item \ordic{nuu dur we.} \gloss{Let’s stay like this.}
\item \ordic{dur far-na nuu.} \gloss{It’s hard, but we’re here.}
\item \ordic{dur shae l nuu.} \gloss{Let’s calm down here.}
\item \ordic{dur an nuu.} \gloss{Let’s stay kind right now.}
\item \ordic{dur xo nuu.} \gloss{Let’s keep the bond.}
\item \ordic{dur\ldots} \gloss{Stay\ldots}

% --- Honest endings
\item \ordic{nuu qe-na.} \gloss{I don’t know right now.}
\item \ordic{nuu mu-na.} \gloss{I can’t remember right now.}
\item \ordic{nuu so-na.} \gloss{I don’t have words right now.}
\item \ordic{nuu ir-na.} \gloss{I’m not angry right now.}
\item \ordic{nuu shae l.} \gloss{It’s calm right now.}
\item \ordic{nuu an.} \gloss{There’s care right now.}
\item \ordic{nuu xo.} \gloss{There’s connection right now.}
\item \ordic{nuu we.} \gloss{We’re here together.}
\item \ordic{nuu\ldots} \gloss{Now\ldots}
\item \ordic{shae l\ldots} \gloss{Calm\ldots}

% --- Settling / aftermath
\item \ordic{shae l nuu.} \gloss{It’s calm now.}
\item \ordic{ir far.} \gloss{The anger faded.}
\item \ordic{ve far.} \gloss{The fear faded.}
\item \ordic{ny far.} \gloss{The shame eased.}
\item \ordic{qe far.} \gloss{The doubt eased.}
\item \ordic{an dur.} \gloss{The care remains.}
\item \ordic{xo dur.} \gloss{The bond remains.}
\item \ordic{we shae l.} \gloss{We’re calm together.}
\item \ordic{nuu shae l we.} \gloss{Right now, we’re okay.}
\item \ordic{dur shae l.} \gloss{Let’s stay calm.}

% --- Acceptance
\item \ordic{es nuu.} \gloss{This is how it is now.}
\item \ordic{es an.} \gloss{It’s okay.}
\item \ordic{es qe-na.} \gloss{It makes sense now.}
\item \ordic{es mu.} \gloss{I remember this.}
\item \ordic{es vi.} \gloss{I see it clearly.}
\item \ordic{es te.} \gloss{This feels true.}
\item \ordic{es far-na.} \gloss{It’s not lost.}
\item \ordic{es dur.} \gloss{It continues.}
\item \ordic{es pos.} \gloss{It will be okay.}
\item \ordic{es nuu shae l.} \gloss{This moment is calm.}

% --- Gratitude / warmth
\item \ordic{gra tu.} \gloss{Thank you.}
\item \ordic{gra we.} \gloss{Thank you, all of you.}
\item \ordic{gra nuu.} \gloss{Thank you, now.}
\item \ordic{an gra.} \gloss{With care, thank you.}
\item \ordic{gra xo.} \gloss{Thank you for the bond.}
\item \ordic{gra dur.} \gloss{Thank you for staying.}
\item \ordic{gra ret.} \gloss{Thank you for coming back.}
\item \ordic{gra shae l.} \gloss{Thank you for the calm.}
\item \ordic{gra an.} \gloss{Thank you for caring.}
\item \ordic{gra\ldots} \gloss{Thanks\ldots}

% --- Gentle reflection
\item \ordic{mu ret.} \gloss{I remember again.}
\item \ordic{mu shae l.} \gloss{The memory feels gentle.}
\item \ordic{mu an.} \gloss{The memory holds care.}
\item \ordic{mu xo.} \gloss{The memory holds us.}
\item \ordic{mu dur.} \gloss{The memory stays.}
\item \ordic{mu far-na.} \gloss{The memory isn’t lost.}
\item \ordic{mu pos.} \gloss{We’ll remember later.}
\item \ordic{mu qe-na.} \gloss{I understand the memory now.}
\item \ordic{mu nuu.} \gloss{I remember this moment.}
\item \ordic{mu\ldots} \gloss{Remembering\ldots}

% --- Togetherness
\item \ordic{we dur.} \gloss{We stayed.}
\item \ordic{we ret.} \gloss{We came back together.}
\item \ordic{we xo.} \gloss{We’re connected.}
\item \ordic{we an.} \gloss{We care.}
\item \ordic{we shae l.} \gloss{We’re calm.}
\item \ordic{we pos.} \gloss{We’ll be okay.}
\item \ordic{we mu.} \gloss{We remember.}
\item \ordic{we vi.} \gloss{We see each other.}
\item \ordic{we es.} \gloss{We are.}
\item \ordic{we\ldots} \gloss{We\ldots}

% --- Soft departures
\item \ordic{ya lin nuu.} \gloss{I’m going now.}
\item \ordic{ya ret pos.} \gloss{I’ll come back later.}
\item \ordic{tu dur.} \gloss{You stay.}
\item \ordic{dur tu.} \gloss{Stay with me.}
\item \ordic{na lin far.} \gloss{Don’t go too far.}
\item \ordic{lin shae l.} \gloss{Go gently.}
\item \ordic{lin an.} \gloss{Go with care.}
\item \ordic{lin xo.} \gloss{Go, still connected.}
\item \ordic{lin pos.} \gloss{See you later.}
\item \ordic{lin\ldots} \gloss{Go\ldots}

% --- Closing silences
\item \ordic{nuu.} \gloss{Now.}
\item \ordic{dur.} \gloss{Stay.}
\item \ordic{shae l.} \gloss{Calm.}
\item \ordic{an.} \gloss{Care.}
\item \ordic{xo.} \gloss{Bond.}
\item \ordic{we.} \gloss{Together.}
\item \ordic{es.} \gloss{Being.}
\item \ordic{nuu\ldots} \gloss{Now\ldots}
\item \ordic{dur\ldots} \gloss{Stay\ldots}
\item \ordic{shae l\ldots} \gloss{Calm\ldots}

\end{enumerate}

\end{multicols}

\vspace{6pt}\hrule\vspace{4pt}
{\small Notes: Low Ordic allows particle drop, emotional leakage, ellipsis, and reassignment. Meaning is carried as much by tone and context as by structure.}

\end{document}
